#### Affordable Floor Area

If the availability of fossil fuel sources such as oil, natural gas, and coal is limited, energy prices may increase as demand outstrips supply. New single-family homeowners may respond to increased energy prices by decreasing their living space in order to save on energy costs. This could involve reducing the size of the home or making other changes to the layout or design of the home to make it more energy efficient.

::: about_strategy
***About this strategy*** Our assumption is that cities can control (or at least influence) the size of new single-family homes to reduce energy use and increasing affordability. For this strategy, the growth of the single-family home size is reduced from `15%` to `5%`to make the new single-family home smaller. We apply this impact to `50%` of new single-family homes. Cities can adjust the growth rate of single-family homes' floor area and the percentage of new single-family homes being controlled.
:::

```{r tab-new-single-family-home-area, message=FALSE, warning=FALSE, include=FALSE}
#| echo: false
#| message: false

tab_new_single_family_home_area <- calc_affordable_floor_area(
  res_tb = building_data$residential,
  .selected_ctu = params$ctu_selection,
  .single_family_floor_area_growth_pct = 0
) %>%
  filter(
    year == 2018,
    var %in% c(
      "multifamily_average_floor_area_sqft_county",
      "multifamily_units",
      "single_family_units",
      "single_family_average_floor_area_sqft_ctu"
    )
  ) %>%
  add_variable_names(., "2018 Baseline") %>%
  full_join(
    calc_affordable_floor_area(
      res_tb = building_data$residential,
      .selected_ctu = params$ctu_selection,
      .single_family_floor_area_growth_pct = 0.05
    ) %>%
      filter(
        year == 2040,
        var %in% c(
          "multifamily_average_floor_area_sqft_county",
          "multifamily_units",
          "single_family_units",
          "single_family_average_floor_area_sqft_ctu"
        )
      ) %>% 
      add_variable_names(., "2040 Scenario"),
    by = "Variables") %>%
  tableformat(.,
              .caption = paste("New Single-Family Home Floor Area for",
                               params$ctu_selection))
```

```{r echo=FALSE}
tab_new_single_family_home_area
```


<br>


$$sqft_{2040}^{sf} = (units_{2040}^{sf} -units_{2018}^{sf}) * 0.5 * \left(\bar{sqft}_{2040}^{sf}-(1+X) * \bar{sqft}_{2018}\right)$$
where

-   $(1+X)$: The rate of growth in single-family floor area in square feet per year.
-   $sqft_{2040}^{sf}$: The total floor area of single-family homes in the community in 2040. 
-   $units_{2040}^{sf}$: The number of single-family housing units in the community in 2040.
-   $units_{2018}^{sf}$: The number of single-family housing units in the community in 2018.
-   $0.5$: This factor represents the assumption that half of the population will reduce the living space of new single-family homes in response to doubling energy prices.
-   $\bar{sqft}_{2040}^{sf}$: The average floor area of single-family homes in the community in 2040.
-   $\bar{sqft}_{2018}$: The average floor area of single-family homes in the community in 2018.

```{r fig03a1, message=FALSE, warning=FALSE, include=FALSE}
#| echo: false
#| message: false
fig03a1 <- 
  run_scenario_building(
  res_tb = building_energy_bau_data$residential,
  res_tb_bau = building_energy_bau_data$residential,
  run_residential = TRUE,
  run_non_residential = FALSE,
  .selected_ctu = params$ctu_selection,
  # affordable floor area
  .single_family_floor_area_growth_pct = 0,
  .new_homes_affected_pct = 1,
  # other variables
  .electrified_buildings_pct = 0,
  .new_homes_to_multifamily_pct = 0,
  .existing_high_efficiency_buildings_pct = 0,
  .home_behavior_change_pct = 0,
  .new_homes_leed_gold_pct = 0,
  .existing_home_retrofit_pct = 0,
  .existing_home_ultra_retrofit_pct = 0,
  .additional_electrified_residential_buildings_pct = 0,
  .grid_decarbonization_pct = 0,
  .enviro_factors = enviro_factors
) %>%
  filter(
    var %in% c(
      "residential_mwh",
      "residential_electricity_emissions_kg_co",
      "residential_therms",
      "residential_natural_gas_emissions_kg_co"
    ),
    year == 2018,
    scen == "bau"
  ) %>%
  ungroup() %>%
  add_variable_names(., "2018 Baseline") %>%
  full_join(
    run_scenario_building(
      res_tb = building_energy_bau_data$residential,
      res_tb_bau = building_energy_bau_data$residential,
      run_residential = TRUE,
      run_non_residential = FALSE,
      .selected_ctu = params$ctu_selection,
      # affordable floor area
      .single_family_floor_area_growth_pct = 0.15,
      .new_homes_affected_pct = 1,
      # other variables
      .electrified_buildings_pct = 0,
      .new_homes_to_multifamily_pct = 0,
      .existing_high_efficiency_buildings_pct = 0,
      .home_behavior_change_pct = 0,
      .new_homes_leed_gold_pct = 0,
      .existing_home_retrofit_pct = 0,
      .existing_home_ultra_retrofit_pct = 0,
      .additional_electrified_residential_buildings_pct = 0,
      .grid_decarbonization_pct = 0,
      .enviro_factors = enviro_factors
    )  %>%
      filter(
        var %in% c(
          "residential_mwh",
          "residential_electricity_emissions_kg_co",
          "residential_therms",
          "residential_natural_gas_emissions_kg_co"
        ),
        year == 2040,
        scen == "scen"
      ) %>%
      ungroup() %>%
      add_variable_names(., "2040 Business-as-Usual"),
    by = "Variables"
  ) %>%
  full_join(
    run_scenario_building(
      res_tb = building_energy_bau_data$residential,
      res_tb_bau = building_energy_bau_data$residential,
      run_residential = TRUE,
      run_non_residential = FALSE,
      .selected_ctu = params$ctu_selection,
      # affordable floor area
      .single_family_floor_area_growth_pct = 0.05,
      .new_homes_affected_pct = 1,
      # other variables
      .electrified_buildings_pct = 0,
      .new_homes_to_multifamily_pct = 0,
      .existing_high_efficiency_buildings_pct = 0,
      .home_behavior_change_pct = 0,
      .new_homes_leed_gold_pct = 0,
      .existing_home_retrofit_pct = 0,
      .existing_home_ultra_retrofit_pct = 0,
      .additional_electrified_residential_buildings_pct = 0,
      .grid_decarbonization_pct = 0,
      .enviro_factors = enviro_factors
    )  %>%
      filter(
        var %in% c(
          "residential_mwh",
          "residential_electricity_emissions_kg_co",
          "residential_therms",
          "residential_natural_gas_emissions_kg_co"
        ),
        year == 2040,
        scen == "scen"
      ) %>%
      ungroup() %>%
      add_variable_names(., "2040 Scenario"),
    by = "Variables") %>%
  
  tableformat(.,
              .caption = paste(
                "New Single-Family Home Floor Area for",
                params$ctu_selection
              )
  )
```

```{r echo=FALSE}
fig03a1
```


