#### Behavior Change for Residential Energy Efficiency

implementing efficient messaging, in-home energy displays, and smart meters can be an effective way to reduce household energy use, but the actual impact may depend on the specific circumstances of each city and the participation rate of households in these actions.

::: about_strategy
***About this strategy*** Efficient messaging, in-home energy display, and smart meters together can reduce household energy use by `11%`. The default is `100%` of homes change behaviors due to these actions. Cities can adjust this participation rate accordingly.
:::

```{r tab-behavior-change}
#| echo: false
#| message: false

calc_floor_area_behavior_change(
  res_tb = building_data$residential,
  .home_behavior_change_pct = 1.00,
  .enviro_factors = enviro_factors
) %>%
  ungroup() %>%
  filter(
    year == 2040,
    var %in% c(
      "multifamily_average_floor_area_sqft_county",
      "multifamily_units",
      "single_family_units",
      "single_family_average_floor_area_sqft_ctu"
    )
  ) %>%
  filter(ctu_name == params$ctu_selection) %>%
  left_join(., variables, by = "var") %>%
  select(desc, value) %>%
  rename("Variables" = desc,
         "Value" = value) %>%
  tableformat(., .caption = paste("Behavior Change", params$ctu_selection, "2040"))
```

<br>

$$ sqft_{2040}^{sf+mf} = X * population_{2040} * sqft_{2040}^{sf + mf }/population_{2040} * 0.11 $$

-   $0.11$: The percentage reduction in household energy use as a result of smart nudges.
-   $X$: The percentage of homes that change their behavior for energy efficiency.
-   $population_{2040}$: The total population in 2040.
-   $sqft_{2040}^{sf+mf}/population_{2040}$: The average square footage per person in 2040 for residential buildings.

*Figure* \@ref(fig:fig-behavior-change) *could be a static or interactive figure. Same questions/issues as above; the main one being is that the larger `run_scenario_building()` function has errors and isn't running right now.*

```{r fig-behavior-change}
#| echo: false
#| message: false

#| fig.cap = "Comparing output values between years 2018 and 2040 for different percentages of existing homes retrofitted to be more energy efficient.", 
#| fig.align = 'center',
#| fig.width = 8, fig.height = 6

seq(0.2, 1, .2) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_floor_area_behavior_change(
  res_tb = building_data$residential,
  .home_behavior_change_pct = .x,
  .enviro_factors = enviro_factors
)) %>%
  bind_rows(.id = "level")%>%
  filter(
    ctu_name == params$ctu_selection,
    var %in% c(
      "multifamily_average_floor_area_sqft_county",
      "multifamily_units",
      "single_family_units",
      "single_family_average_floor_area_sqft_ctu",
      "residential_mwh",
      "residential_therms"
    ))  %>%
  left_join(., variables, by = "var") %>%
  mutate(desc = str_replace_all(desc, "ft<sup>2</sup>", "ft2")) %>%
      rename("Variables" = desc) %>%#,
             # "2040" = value)
  mutate(level = paste0(as.numeric(level) * 100, "%")) %>%
  ggplot(aes(x = year, y = value, color = level)) +
  geom_line() +
  facet_wrap(~Variables, scales = "free", labeller = label_wrap_gen(width = 25)) +
  scale_x_continuous(breaks = c(2018, 2040)) +
  scale_color_manual(values = RColorBrewer::brewer.pal(6, "Blues")[2:6],
                     breaks = c("20%", "40%", "60%", "80%", "100%")) +
  labs(x = "", y = "", color = "Percentage of\nhomes reducing\nenergy use\nvia behavior\nchange") +
      theme(
      panel.background = element_rect(fill = "white"),
      panel.grid = element_blank(),
      axis.title = element_text(size = 14),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(size = 10)
    )

```