#### Decarbonization of Residential Energy Use

The energy grid could become cleaner by increasing the use of renewable energy sources such as solar, wind, and hydroelectric power, which produce significantly fewer greenhouse gas emissions than fossil fuels

::: about_strategy
***About this strategy***: Due to decarbonizing the grid, the emission factor of electricity decreases from `568` kg<sub>2</sub>e/MWh (2018 Minnesota state level) to `113` kgCO<sub>2</sub>e/MWh. This change is fixed by the grid.
:::

```{r tab-decarbonize-residential-energy}
#| echo: false
#| message: false

calc_ghg_residential(
  res_tb = building_energy_bau_data$residential,
  res_tb_bau = building_energy_bau_data$residential,
  .grid_decarbonization_pct = 0.8,
  .enviro_factors = enviro_factors
) %>%
  ungroup() %>%
  filter(
    ctu_name == params$ctu_selection,
    year == 2040,
    scen == "scen",
    var %in% c(
      "residential_mwh",
      "residential_electricity_emissions_kg_co",
      "residential_therms",
      "residential_natural_gas_emissions_kg_co"
    )
  ) %>%
  left_join(., variables, by = "var") %>%
  select(desc, value) %>%
  rename("Variables" = desc,
         "Value" = value) %>%
  tableformat(., .caption = "Decarbonize Residential Energy Use Minneapolis 2040")
```

<br>

$$
kgCO_2e = (population_{2040} * sqft/person_{2040}^{residential} * kWh/sqft_{residential}^{2040}) 
$$

*Figure* \@ref(fig:fig-decarbonize-residential-energy) *could be a static or interactive figure. Same questions/issues as above; the main one being is that the larger `run_scenario_building()` function has errors and isn't running right now.*

```{r fig-decarbonize-residential-energy}
#| echo: false
#| message: false

#| fig.cap = "Comparing output values between years 2018 and 2040 for different percentages of decarbonizing the grid.", 
#| fig.align = 'center',
#| fig.width = 8, fig.height = 6

seq(0.2, 1, .2) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_ghg_residential(
  res_tb = building_energy_bau_data$residential,
  res_tb_bau = building_energy_bau_data$residential,
  .grid_decarbonization_pct = .x,
  .enviro_factors = enviro_factors
)) %>%
  bind_rows(.id = "level")%>%
  filter(
    ctu_name == params$ctu_selection,
    scen == "scen",
    var %in% c(
      "residential_mwh",
      "residential_electricity_emissions_kg_co",
      "residential_therms",
      "residential_natural_gas_emissions_kg_co",
      "total_residential_emissions"
    ))  %>%
  left_join(., variables, by = "var") %>%
  mutate(desc = str_replace_all(desc, "<sub>2</sub>", "2")) %>%
      rename("Variables" = desc) %>%#,
             # "2040" = value)
  mutate(level = paste0(as.numeric(level) * 100, "%")) %>%
  ggplot(aes(x = year, y = value, color = level)) +
  geom_line() +
  facet_wrap(~Variables, scales = "free", labeller = label_wrap_gen(width = 25)) +
  scale_x_continuous(breaks = c(2018, 2040)) +
  scale_color_manual(values = RColorBrewer::brewer.pal(6, "Blues")[2:6],
                     breaks = c("20%", "40%", "60%", "80%", "100%")) +
  labs(x = "", y = "", color = "Percentage of\ngrid which is\ndecarbonized") +
      theme(
      panel.background = element_rect(fill = "white"),
      panel.grid = element_blank(),
      axis.title = element_text(size = 14),
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(size = 10)
    )
```