#### Retofit Existing Commercial Building 

Existing commercial buildings electrification.

::: about_strategy
***About this strategy*** Our assumption is that cities can control (or at least influence) the size of new single-family homes to reduce energy use and increasing affordability. For this strategy, the growth of the single-family home size is reduced from `15%` to `5%`to make the new single-family home smaller. We apply this impact to `50%` of new single-family homes. Cities can adjust the growth rate of single-family homes' floor area and the percentage of new single-family homes being controlled.
:::

```{r tab-existing-comm-building-efficiiency, include=FALSE}
tab_existing_comm_building_efficiency <-
  calc_existing_comm_building_efficiency(
    non_res_tb = building_energy_bau_data$non_residential,
    .selected_ctu = params$ctu_selection,
    .existing_high_efficiency_buildings_pct = 0.80
  )  %>%
  # filter out irrelevant rows
  filter(
    year == 2018,
    var %in% c(
      "commercial_mwh_per_worker",
      "industrial_mwh_per_worker",
      "commercial_mwh",
      "industrial_mwh",
      "commercial_therms",
      "industrial_therms",
      "total_industrial_commercial_emissions"
    )
  ) %>%
  # join variable names
  add_variable_names(., "2018") %>%
  full_join(
    calc_existing_comm_building_efficiency(
      non_res_tb = building_energy_bau_data$non_residential,
      .selected_ctu = params$ctu_selection,
      .existing_high_efficiency_buildings_pct = 0.80
    )  %>%
      # filter out irrelevant rows
      filter(
        year == 2040,
        var %in% c(
          "commercial_mwh_per_worker",
          "industrial_mwh_per_worker",
          "commercial_mwh",
          "industrial_mwh",
          "commercial_therms",
          "industrial_therms",
          "total_industrial_commercial_emissions"
        )
      ) %>%
      # join variable names
      add_variable_names(., "2040")
  ) %>%
  # add table format
  tableformat(.,
              .caption = paste("Retrofit Commercial Buildings",
                               params$ctu_selection,
                               "2040"))
```

```{r echo=FALSE}
tab_existing_comm_building_efficiency
```

```{r}
run_scenario_building(
  
)

```


