#### Decarbonize the Electricity Grid

Grid decarbonization 

::: about_strategy
***About this strategy*** Our assumption is that cities can control (or at least influence) the size of new single-family homes to reduce energy use and increasing affordability. For this strategy, the growth of the single-family home size is reduced from `15%` to `5%`to make the new single-family home smaller. We apply this impact to `50%` of new single-family homes. Cities can adjust the growth rate of single-family homes' floor area and the percentage of new single-family homes being controlled.
:::

```{r fig03b4, include=FALSE}
# 2018
fig03b4_2018 <- run_scenario_building(
   run_residential = FALSE,
   run_non_residential = TRUE,
   .selected_ctu = params$ctu_selection,
   non_res_tb = building_data$non_residential,
   non_res_tb_bau = building_data$non_residential,
   .renewable_ng_nonres = FALSE,
   .electrified_buildings_pct = 0,
   .non_res_natural_gas_for_water_heating_pct = 0.20,
   .non_res_natural_gas_for_space_heating_pct = 0.69,
   .commercial_smart_grid_pct = 1,
   .industrial_smart_grid_pct = 1,
   .smart_grid_energy_reduction_pct = 1,
   .grid_decarbonization_pct = 0,
   .existing_high_efficiency_buildings_pct = 0,
   .enviro_factors = enviro_factors
) %>%
   ungroup() %>%
   dplyr::filter(
      year == 2018,
      scen == "bau",
      var %in% c(
         "commercial_mwh",
         "industrial_mwh",
         "commercial_therms",
         "industrial_therms",
         "total_industrial_commercial_emissions"
      )
   ) %>%
   add_variable_names(., "2018 Baseline")

# 2040 BAU
fig03b4_2040bau <- run_scenario_building(
   run_residential = FALSE,
   run_non_residential = TRUE,
   .selected_ctu = params$ctu_selection,
   non_res_tb = building_data$non_residential,
   non_res_tb_bau = building_data$non_residential,
   .renewable_ng_nonres = FALSE,
   .electrified_buildings_pct = 0,
   .non_res_natural_gas_for_water_heating_pct = 0.20,
   .non_res_natural_gas_for_space_heating_pct = 0.69,
   .commercial_smart_grid_pct = 1,
   .industrial_smart_grid_pct = 1,
   .smart_grid_energy_reduction_pct = 1,
   .grid_decarbonization_pct = 0,
   .existing_high_efficiency_buildings_pct = 0,
   .enviro_factors = enviro_factors
) %>%
   ungroup() %>%
   dplyr::filter(
      year == 2040,
      scen == "bau",
      var %in% c(
         "commercial_mwh",
         "industrial_mwh",
         "commercial_therms",
         "industrial_therms",
         "total_industrial_commercial_emissions"
      )
   ) %>%
   add_variable_names(., "2040 Busines-As-Usual")

# 2040 Scenario
fig03b4_2040scen <- run_scenario_building(
   run_residential = FALSE,
   run_non_residential = TRUE,
   .selected_ctu = params$ctu_selection,
   non_res_tb = building_data$non_residential,
   non_res_tb_bau = building_data$non_residential,
   .renewable_ng_nonres = FALSE,
   .electrified_buildings_pct = 0,
   .non_res_natural_gas_for_water_heating_pct = 0.20,
   .non_res_natural_gas_for_space_heating_pct = 0.69,
   .commercial_smart_grid_pct = 1,
   .industrial_smart_grid_pct = 1,
   .smart_grid_energy_reduction_pct = 1,
   .grid_decarbonization_pct = 0,
   .existing_high_efficiency_buildings_pct = 0,
   .enviro_factors = enviro_factors
) %>%
   ungroup() %>%
   dplyr::filter(
      year == 2040,
      scen == "scen",
      var %in% c(
         "commercial_mwh",
         "industrial_mwh",
         "commercial_therms",
         "industrial_therms",
         "total_industrial_commercial_emissions"
      )
   ) %>%
   add_variable_names(., "2040 Scenario")

fig03b4 <- full_join(fig03b4_2018,
                     fig03b4_2040bau,
                     by = "Variables") %>%
   full_join(.,
             fig03b4_2040scen,
             by = "Variables")
```


```{r show_fig03b4, echo=FALSE}
fig03b4 %>%
  # add table format
  tableformat(.,
              .caption = paste("Decarbonise Grid",
                               params$ctu_selection,
                               "2040"))
```