# Building Energy - Non-Residential

Non-residential energy use refers to the energy consumption of commercial and industrial buildings, including both electricity and natural gas. The baseline data for energy use was obtained from Xcel Energy. The forecast of "business-as-usual" energy use is estimated by considering the number of people who are expected to work in these buildings.

In addition to electricity and natural gas, non-residential energy use may also include other energy sources such as propane, fuel oil, and renewable energy sources, some of which are not captured in this methodology. Accurate estimates of non-residential energy use can help `r params$ctu_selection` make informed decisions about energy efficiency and conservation efforts.

**Table** \@ref(tab:tab-non-residential-energy) shows the baseline and forecast assumptions for non-residential electricity and natural gas use for `r params$ctu_selection`.

```{r tab-non-residential-energy, include=FALSE}
#| echo: false
#| message: false
# city baseline
tab_non_residential_energy <- get_non_residential_energy_baseline(
  tb = building_energy_data,
  .selected_ctu = params$ctu_selection
) %>%
  add_variable_names(., "2018") %>%
  # city forecast
  full_join(
    calc_non_residential_energy_forecast(
      tb = building_energy_data,
      .selected_ctu = params$ctu_selection
    ) %>%
      add_variable_names(., "2040")
  ) %>%
  tableformat(.,
    .caption = paste(
      "Non-Residential Energy for ",
      params$ctu_selection
    )
  )
```

```{r echo=FALSE}
tab_non_residential_energy
```
As shown in \@ref(tab:tab-non-residential-energy) for making estimate about the business-as-usual energy use of the City of `r params$ctu_selection`, we assue that energy use intensity per worker for the commercial and industrial sectors remain constant. 

<br>

### Strategies

One of the most effective strategies for commercial and industrial buildings is to enhance their energy efficiency, which can be achieved through the adoption of new technologies. In Minnesota, there are several programs aimed at supporting non-residential consumers.

The **Conservation Improvement Program (CIP)** is a policy aimed at improving energy efficiency and conservation in households and businesses in Minnesota. Funded by ratepayers and administered by over 180 electric and natural gas utilities, with oversight from the Minnesota Department of Commerce's Division of Energy Resources, CIP promotes energy-efficient technologies and practices to residential, commercial, and industrial customers through various programs, such as energy audits, rebates on high-efficiency appliances, LED lighting, and air conditioning cycling programs.

The **Energy Conservation and Optimization Act of 2021 (ECO Act)** aims to strengthen the state's energy conservation programs, reduce greenhouse gas emissions, and create local jobs through energy efficiency projects in homes and businesses. The ECO Act expands the Conservation Improvement Program and makes energy efficiency improvements available to low-income households, public schools, and investor-owned utilities.

The **Minnesota SB2030** program focuses on reducing emissions in commercial, industrial, and institutional buildings. The program sets standards for energy efficiency and renewable energy for new and renovated buildings, based on the building's energy use per square foot. Over time, these standards will become more stringent, with the goal of reducing energy use and carbon dioxide emissions. The program also takes into account the impact of electric vehicle charging infrastructure on a building's energy use.

```{r child="03b_non_residential_strategies/03b1_retrofit_commercial_bldgns.Rmd", echo=FALSE}
```

```{r child="03b_non_residential_strategies/03b2_smart_grid_non_res.Rmd", echo=FALSE}
```

```{r child="03b_non_residential_strategies/03b3_non_res_renewable_ng.Rmd", echo=FALSE}
```

```{r child="03b_non_residential_strategies/03b4_non_res_decarbonize_grid.Rmd", echo=FALSE}
```