## Interventions/Strategies

The transportation module uses passenger miles traveled as a proxy for emissions. Miles traveled are calculated for each travel mode and then converted into emissions.

### Land Use

Transportation and land use policies are reliant on each other. Segregated and low-density land use patterns lead to a reliance on automobiles, increased vehicle miles traveled, and decreased efficacy of shared travel. However, altering land use policy is complex and only sometimes leads to changes in land use, as construction takes time to catch up with zoning. The scenario planning model uses the "five D's" (population and employment density, diversity of land uses, street design, distance to transit, and destination accessibility) as the land use factors that affect transportation.

The Metropolitan Council Transportation Policy Plan discusses the need for communities to consider the connections between land use and transportation as the region changes in the coming years. The plan mentions that it is not feasible to expand the highway system in a sustainable manner and that there is increased private investment in already developed parts of the region along transit lines. The Transportation Policy Plan also mentions that higher-density development and redevelopment occur in already-developed areas. That new growth is happening in suburban edge and emerging suburban edge communities, which stresses the regional highway system. The plan suggests that planners incorporate multimodal travel options, such as transit, walking, and biking, into land use and design to address issues like congestion, climate change, equity, and livability.

Increasing density decreases travel distances and encourages transit, cycling, and walking. We estimate a `0.4%` decrease in auto VMT for a `10%` increase in population density by 2040, a `7%` increase in walking/cycling VMT for a `10%` increase in population density by 2040, and a `7%` increase in transit VMT for a `10%` increase in population density by 2040.

```{r, tab-land-use-effects}
#| echo: false
#| message: false
ghg.sp::elast_5d %>%
  dplyr::filter(year == "2050") %>%
  mutate(type = stringr::str_to_title(type)) %>%
  select(
    `Mode` = type,
    `Population density` = population_density,
    `Employment density` = employment_density,
    `Land use diversity` = diversity,
    `Intersection design` = design,
    `Job accessability by transit` = job_access,
    `Distance to transit` = distance
  ) %>%
  tidyr::pivot_longer(
    names_to = "Land Use Adjustment",
    values_to = "Effect on miles traveled by 2050",
    cols = 2:7
  ) %>%
  mutate(
    `Multiplier relative to BAU` =
      ifelse(`Land Use Adjustment` == "Distance to transit", -0.1, 0.1)
  ) %>%
  select(
    Mode, `Land Use Adjustment`, `Multiplier relative to BAU`,
    `Effect on miles traveled by 2050`
  ) %>%
  rowwise() %>%
  mutate(across(c(3, 4), function(x) {
    ifelse(x != "NA",
           paste0(ifelse(x >= 0, "+", ""), x * 100, "%"),
           x
    )
  })) %>%
  gt::gt(
    id = "land-use-effects",
    caption = "Land use adjustment values",
    rowname_col = "`Land Use Adjustment`",
    groupname_col = "Mode"
  ) %>%
  gt::opt_row_striping() %>%
  gt::tab_options(row_group.font.weight = "bold")
```

```{r tab-land-use-change}
#| echo: false
#| message: false
# vmt_land_use_change()
```

```{r fig-landuse}
#| echo: false
#| message: false

#| fig.cap = "Comparing the direct and embodied greenhouse gas emissions between years 2018 and 2040 for different percentages of population density (aka land use) change.", 
#| fig.align = 'center',
#| fig.width = 8, fig.height = 3
# vmt_telework()
density_ctu %>%
  filter(type == "direct",
         year %in% c("2018", "2040")) %>% 
  mutate(level = paste0(as.numeric(param) * 100, "%")) %>%
  ggplot(aes(x = year, y = emissions, color = level, group = level)) +
  geom_line() + 
  facet_wrap(~type, scales = "free", 
             labeller = as_labeller(c(direct = str_wrap("Direct emissions", 10), 
                                      embodied = str_wrap("Embodied emissions", 10)
             ))) +
  scale_color_manual(values = RColorBrewer::brewer.pal(7, "Blues")[2:7],
                     breaks = c("-100%", "-50%", "0%", "50%", "100%")) +
  labs(x = "", 
       y = "", 
       color = paste0(params$ctu_selection,
                      "\npopulation density\nadjustment (%)")) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(
    panel.background = element_rect(fill = "white"),
    # panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 10)
  )
```

### Vehicle Electrification

The Greenhouse Gas Scenario Planning Tool aims to assess the potential impact of electric vehicles on greenhouse gas emissions in the Twin Cities metropolitan region.

The Tool uses data from the Market Acceptance of Advanced Automotive Technologies model to forecast electric vehicle market penetration in the region, and considers the GHG intensity of the region's electricity grid in calculating the GHG impact of electric vehicles.

The Tool also accounts for the fact that plug-in hybrid electric vehicles may operate in both electric and gasoline modes, and estimates the fraction of time that these vehicles operate in each mode based on the average distance that they can travel on electric power and data on household vehicle usage. The tool aims to provide insight into the potential for electric vehicles to reduce GHG emissions in the region and inform policy decisions related to electric vehicle adoption.

```{r fig-bev}
#| echo: false
#| message: false

#| fig.cap = "Comparing the direct and embodied greenhouse gas emissions between years 2018 and 2040 for different percentages of vehicles sales which are battery electric.", 
#| fig.align = 'center',
#| fig.width = 8, fig.height = 3
# vmt_annual_energy_outlook()
bev_ctu %>%
  filter(year %in% c("2018", "2040")) %>% 
  mutate(level = paste0(as.numeric(param) * 100, "%")) %>%
  ggplot(aes(x = year, y = emissions, color = level, group = level)) +
  geom_line() + 
  facet_wrap(~type, scales = "free", 
             labeller = as_labeller(c(direct = str_wrap("Direct emissions", 10), 
                                      embodied = str_wrap("Embodied emissions", 10)
             ))) +
  scale_color_manual(values = RColorBrewer::brewer.pal(7, "Blues")[2:7],
                     breaks = c("20%", "40%", "60%", "80%", "100%")) +
  labs(x = "", y = "", color = paste0(params$ctu_selection, "\nsales that are\nbattery electric (%)")) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(
    panel.background = element_rect(fill = "white"),
    # panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 10)
  )
```

### Public Transit

According to the [Minneapolis Transportation Action Plan (TAP)](https://go.minneapolismn.gov/), transit is essential for the city's multimodal transportation system. Transit contributes to economic competitiveness by attracting businesses, private investment, and top talent. Over 30,000 (16.5%) households in the city do not have access to or choose not to own a personal car, with the highest concentration of car-free individuals living in neighborhoods around downtown Minneapolis. The TAP aims to improve the coverage, speed, and reliability of transit service through a partnership with the Metropolitan Council and by defining clear priorities, goals, and actions for the city. The goal is to increase the number of transit trips from 13% in 2019 to 25% by 2030. Transit must be convenient, reliable, and frequent and effectively reduce trips made by single-occupancy vehicles to achieve these goals.

```{r tab-transit-service}
#| echo: false
#| message: false
# vmt_transit_service()
```

```{r fig-transitservice}
#| echo: false
#| message: false

#| fig.cap = "Comparing the direct and embodied greenhouse gas emissions between years 2018 and 2040 for different percentages of transit service.", 
#| fig.align = 'center',
#| fig.width = 8, fig.height = 3
# vmt_telework()
telework_ctu %>%
  filter(year %in% c("2018", "2050"),
         type == "direct") %>% 
  mutate(level = paste0(as.numeric(param) * 100, "%")) %>%
  ggplot(aes(x = year, y = emissions, color = level, group = level)) +
  geom_line() + 
  facet_wrap(~type, scales = "free", 
             labeller = as_labeller(c(direct = str_wrap("Direct emissions", 10), 
                                      embodied = str_wrap("Embodied emissions", 10)
             ))) +
  scale_color_manual(values = RColorBrewer::brewer.pal(7, "Blues")[2:7],
                     breaks = c("20%", "40%", "60%", "80%", "100%")) +
  labs(x = "", y = "", color = paste0(params$ctu_selection, "\ntransit service\nadjustment (%)")) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(
    panel.background = element_rect(fill = "white"),
    # panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 10)
  )
```

### Telework

Telework is working from home, reducing or eliminating the need to commute. In 2020, an estimated 35% of jobs in the United States could be performed from home, varying by metropolitan area. In Minneapolis and St. Paul, approximately 38-52% of jobs can be completed through telework.1 In a first approximation, telework eliminates commute vehicle miles traveled, which can reduce greenhouse gas emissions. However, the net effect on vehicle miles traveled, and household greenhouse gas production must also be considered. Recent studies that include energy associated with working from home find an effective reduction of about 2.5-3% in household CO2 emissions per telework day.

```{r}
#| echo: false
#| message: false
#help(vmt_telework, package = ghg.sp)
```


```{r fig-telework}
#| echo: false
#| message: false

#| fig.cap = "Comparing the direct and embodied greenhouse gas emissions between years 2018 and 2040 for different percentages of telworking.", 
#| fig.align = 'center',
#| fig.width = 8, fig.height = 3


# vmt_telework()
telework_ctu %>%
    filter(type == "direct",
         year %in% c("2018", "2040")) %>% 
  mutate(level = paste0(as.numeric(param) * 100, "%")) %>%
  ggplot(aes(x = year, y = emissions, color = level, group = level)) +
  geom_line() + 
  facet_wrap(~type, scales = "free", 
             labeller = as_labeller(c(direct = str_wrap("Direct emissions", 10), 
                                      embodied = str_wrap("Embodied emissions", 10)
             ))) +
  scale_color_manual(values = RColorBrewer::brewer.pal(7, "Blues")[2:7],
                     breaks = c("20%", "40%", "60%", "80%", "100%")) +
  labs(x = "", y = "", color = paste0(params$ctu_selection, "\nteleworking (%)")) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(
    panel.background = element_rect(fill = "white"),
    # panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 10)
  )
```

### Road Pricing

In the future, road pricing will likely be an essential component of transportation policy. Road pricing is key for automated vehicles and dynamic ride-sharing policies because it helps address the increased vehicle miles traveled projected to occur with adopting these technologies. Furthermore, road pricing provides an alternative revenue stream to the gas tax as vehicle efficiency increases and electric vehicle adoption grows.

The scenario planning model includes three alternatives for road pricing. The first is a vehicle miles traveled (VMT) fee that charges for each mile of travel. The second is a pay-as-you-drive insurance plan. Finally, the model includes a representation of congestion pricing. Ideally, congestion price should vary between the time of day and road link based on the current congestion level, but this is infeasible to model. Instead, an elevated VMT fee is applied to a portion of travel, approximating the percent of congested VMT. Credit-based congestion pricing is recommended as a revenue-neutral strategy to address equity in road pricing. It provides credits to travelers who shift their travel model or departure time.

```{r tab-road-policy}
#| echo: false
#| message: false
# vmt_road_policy()
```

*Figure* \@ref(fig:fig-roadprice) *could be a static or interactive figure. However, nothing seems to matter...*

```{r fig-roadprice}
#| echo: false
#| message: false

#| fig.cap = "Comparing the direct and embodied greenhouse gas emissions between years 2018 and 2040 for different levels of road pricing.", 
#| fig.align = 'center',
#| fig.width = 8, fig.height = 3

roadprice_ctu %>%
    filter(type == "direct",
         year %in% c("2018", "2040")) %>% 
  mutate(level = paste0("$", as.numeric(param))) %>%
  ggplot(aes(x = year, y = emissions, color = level, group = level)) +
  geom_line() + 
  facet_wrap(~type, scales = "free", 
             labeller = as_labeller(c(direct = str_wrap("Direct emissions", 10), 
                                      embodied = str_wrap("Embodied emissions", 10)
             ))) +
  scale_color_manual(values = RColorBrewer::brewer.pal(7, "Blues")[2:7]
                     # ,breaks = c("20%", "40%", "60%", "80%", "100%")
  ) +
  labs(x = "", y = "", color = paste0(params$ctu_selection, "\ncongestion price\n($ per mile)")) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(
    panel.background = element_rect(fill = "white"),
    # panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 10)
  )
```

### Parking Policy

Parking is an important component of the transportation system. A lack of parking can result in additional VMT from vehicles searching for a parking spot (up to 30% of vehicles in downtown traffic [@shoupCruisingParking2006]). Surface lots and parkades take up valuable space that could be put to more useful purposes.

The most common strategies for controlling parking demand are increasing its price and restricting the time that parking is available. These strategies have been reviewed in detail and price elasticities are available for the price of parking on VMT (see Table \@ref(tab:tab-parking-price-effects)).

The long-run elasticity of VMT to parking cost is estimated to be in the range of -0.18 to -0.45, with a mean of 0.07 based on a meta-analysis of values reported in the literature [@lehnerPriceElasticityParking2019].

```{=tex}
\begin{equation}
adj_{park} = 1 + \frac{price_{y}}{price_{y = 2018}} \times elast_{park_{m}} 
(\#eq:parking_price_adj)
\end{equation}
```
$price_y$ is the price of parking on dollars per hour in year $y$\
$price_{y = 2018}$ is the price of parking in dollars per hour in 2018\
$elast_{park}$ is the elasticity or effect size of parking price increase on travel mode $m$.

Values derived from @TRACE_1999 and referenced in @litmanUnderstandingTransportDemands2021 and @lehnerPriceElasticityParking2019.

```{r, tab-parking-price-effects}
#| echo: false
#| message: false
ghg.sp::elast %>%
  dplyr::filter(year == "2050") %>% 
  select(year, contains("park")) %>% 
  tidyr::pivot_longer(
    # names_to = "Increase in parking price",
    values_to = "Effect on miles traveled by 2050",
    cols = 2:4
  ) %>% 
  mutate(`Mode` = case_when(name == "park_elast" ~ "Automobile",
                            name == "park_active" ~ "Walking and cycling",
                            name == "park_transit" ~ "Transit")) %>% 
  mutate(
    `Increase in parking price relative to baseline` = 0.1,
    "Increase in parking price" = name
  ) %>%
  select(
    Mode,
    `Increase in parking price relative to baseline`,
    `Effect on miles traveled by 2050`
  ) %>% 
  rowwise() %>%
  mutate(across(c(2:3), function(x) {
    ifelse(x != "NA",
           paste0(ifelse(x >= 0, "+", ""), x * 100, "%"),
           x
    )
  })) %>%
  gt::gt(
    id = "parking-price-effects",
    caption = "Parking price adjustment values"
  ) %>%
  gt::tab_source_note(gt::md("TRACE 1999")) %>% 
  gt::opt_row_striping() %>%
  gt::tab_options(row_group.font.weight = "bold")
```


```{r fig-parkingprice}
#| echo: false
#| message: false

#| fig.cap = "Comparing the direct and embodied greenhouse gas emissions between years 2018 and 2040 for levels of parking pricing.", 
#| fig.align = 'center',
#| fig.width = 8, fig.height = 3

parking_ctu %>%
    filter(type == "direct",
         year %in% c("2018", "2040")) %>% 
  mutate(level = paste0("$", as.numeric(param))) %>%
  ggplot(aes(x = year, y = emissions, color = level, group = level)) +
  geom_line() + 
  facet_wrap(~type, scales = "free", 
             labeller = as_labeller(c(direct = str_wrap("Direct emissions", 10), 
                                      embodied = str_wrap("Embodied emissions", 10)
             ))) +
  scale_color_manual(values = RColorBrewer::brewer.pal(7, "Blues")[2:7]
                     # ,breaks = c("20%", "40%", "60%", "80%", "100%")
  ) +
  labs(x = "", y = "", color = paste0(params$ctu_selection, "\nparking price\n($ per hour)")) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(
    panel.background = element_rect(fill = "white"),
    # panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 10)
  )
```
