---
title: "Building Energy Inputs - Non-residential"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{building_energy_module_inputs_non_residential}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Techical Walkthrough

All the functions for the Building Energy Module are contained within the `ghg.sp` package.
```{r setup}
library(ghg.sp)
library(tidyr)
library(dplyr)
```

## Source Datasets

The `building_data.rda` is a preloaded dataset that includes cleaned datasets to 
run the building energy module of the `ghg.sp` package. 

This vignette covers non-residential energy inputs. 

## Running the module 

The `ghg.sp::run_scenario_building()` function compiles all of the strategies for the Building Energy Module and produces an output table. 

`ghg.sp::run_scenario_building()` call two separate functions:

* `ghg.sp::scen_building_residential()`
* `ghg.sp::scen_building_non_residential()`


Each of these functions, in turn call subsequent functions to obtain the estimate of greenhouse gas emissions under the scenarios specified by the end-user.

### Non-Residential Energy 

The following chunk prints the key input variables for the commercial/industrial strategies under the building energy module. 

```{r building}
building_data$non_residential %>%
  dplyr::filter(ctu_name == "Minneapolis") %>%
  tidyr::pivot_wider(names_from = year, values_from = value) %>%
  dplyr::select(ctu_name, var, `2018`, `2040`) %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.numeric, round, 2)
```

#### Non-Residential Variables

```{r table-non-residential-variables}
non_residential_variables
```
