---
title: "Building Energy Outputs - Residential"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{building_energy_module_outputs_residential}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ghg.sp)
```

### Residential Strategies 

#### Strategy B.R1: More Compact Residential Buildings 

This strategy goes beyond the Metropolitan Council’s 2040 Planned Land Use, by assuming that *50%* of all new single-family construction will now become multifamily, with the multifamily homes having smaller square footage. 

The main function related to this strategy is `ghg.sp::adj_unit_counts()` which allows to specify the percentage of single family homes to be converted to multifamily. 

```{r}
ghg.sp::adj_unit_counts(res_tb = building_data$residential,
                        .new_homes_to_multifamily_pct = 0.50) %>%
  dplyr::filter(
    ctu_name == "Lake Elmo",
    var %in% c(
      "multifamily_units",
      "single_family_units",
      "single_family_average_floor_area_sqft_ctu",
      "multifamily_average_floor_area_sqft_county"
    )
  ) %>%
  tidyr::pivot_wider(names_from = year, values_from = value) %>%
  dplyr::select(ctu_name, var, `2018`, `2040`) %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.numeric, round, 2)
```

#### Strategy B.R2: Increase Affordable Floor Area

This strategy assumes that about 50% of the single-family homes will respond to increased energy prices by decreasing their living space.

```{r}
ghg.sp::calc_affordable_floor_area(res_tb = building_data$residential,
                                   .single_family_floor_area_growth_pct = 0.05) %>%
  dplyr::filter(
    ctu_name == "Lake Elmo",
    var %in% c(
      "multifamily_units",
      "single_family_units",
      "single_family_average_floor_area_sqft_ctu",
      "multifamily_average_floor_area_sqft_county"
    )
  ) %>%
  tidyr::pivot_wider(names_from = year, values_from = value) %>%
  dplyr::select(ctu_name, var, `2018`, `2040`) %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.numeric, round, 2)
```

#### Strategy B.R3: Make New Homes LEED-Gold

This strategy calculates the effect of 50% of new homes being constructed to [LEED Gold](https://www.usgbc.org/leed) standards.

The main function involved in this strategy is `ghg.sp::calc_floor_area_leed()` which allows to specify the number of new homes to turn into LEED Gold. 

```{r}
ghg.sp::calc_floor_area_leed(
  res_tb = building_data$residential,
  .new_homes_leed_gold_pct = 0.5,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(
    ctu_name == "Lake Elmo",
    var %in% c(
      "multifamily_units",
      "single_family_units",
      "single_family_average_floor_area_sqft_ctu",
      "multifamily_average_floor_area_sqft_county"
    )
  ) %>%
  tidyr::pivot_wider(names_from = year, values_from = value) %>%
  dplyr::select(ctu_name, var, `2018`, `2040`) %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.numeric, round, 2)
```

#### Strategy B.R4: Retrofit Existing Homes to High Energy Efficiency

This strategy focuses on existing homes and assumes *80%* of them are retrofitted to performance-based high energy efficiency standards, which reduce energy use by *33%*, which is mid-way between the trends in EUI already seen, and a passive-house standard. Most of these cases have on-site energy use for renewable electricity, so they conflate renewable electricity with efficiency. 

The main function involved in this strategy is `ghg.sp::calc_floor_area_retrofit()` which allows to specify the number of existing homes to be retrofit. 

**Note**: Strategy **B.R5** & **B.R6** are carried simultaneously by function `ghg.sp::calc_floor_area_retrofit()`.

```{r}
ghg.sp::calc_floor_area_retrofit(
  res_tb = building_data$residential,
  .existing_home_retrofit_pct = 0.80,
  .existing_home_ultra_retrofit_pct = 0.20,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(
    ctu_name == "Lake Elmo",
    var %in% c(
      "multifamily_units",
      "single_family_units",
      "single_family_average_floor_area_sqft_ctu",
      "multifamily_average_floor_area_sqft_county"
    )
  ) %>%
  tidyr::pivot_wider(names_from = year, values_from = value) %>%
  dplyr::select(ctu_name, var, `2018`, `2040`) %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.numeric, round, 2)
```

#### Strategy B.R5: Retrofit Existing Homes to Passive-House

This strategy assumes that the remaining *20%* of homes from strategy B.R5 are retrofitted to high passive house standards, which results in *66%* reduction in energy use intensity.

The main function involved in this strategy is `ghg.sp::calc_floor_area_retrofit()` which allows to specify the number of existing homes to be retrofit.

**Note**: Strategy **B.R4** & **B.R5** are carried simultaneously by function `ghg.sp::calc_floor_area_retrofit()`. 

```{r}
ghg.sp::calc_floor_area_retrofit(
  res_tb = building_data$residential,
  .existing_home_retrofit_pct = 0.80,
  .existing_home_ultra_retrofit_pct = 0.20,
  .enviro_factors = enviro_factors
)  %>%
  dplyr::filter(
    ctu_name == "Lake Elmo",
    var %in% c(
      "multifamily_units",
      "single_family_units",
      "single_family_average_floor_area_sqft_ctu",
      "multifamily_average_floor_area_sqft_county"
    )
  ) %>%
  tidyr::pivot_wider(names_from = year, values_from = value) %>%
  dplyr::select(ctu_name, var, `2018`, `2040`) %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.numeric, round, 2)
```

#### Strategy B.R6: Implement Behavioral & Smart Nudges 

This strategy assumes that all homes receive effective messages which encourage them to reduce energy use via nudges such as in-home display and smart meters. Such nudges can reduce household energy use by *11%*, based on **Gilbert et al. (ACEEE 2010)**. We only apply this reduction to remaining electricity and gas after the building efficiency strategies. 

The main function involved in this strategy is `ghg.sp::calc_floor_area_behavior_change()`. 

**Note**: Strategy B.R7 is dependent on the outputs (if any) from strategies B.R5 and B.R6.  

```{r}
ghg.sp::calc_floor_area_behavior_change(
  res_tb = building_data$residential,
  .home_behavior_change_pct = 1.00,
  .enviro_factors = enviro_factors
)   %>%
  dplyr::filter(
    ctu_name == "Lake Elmo",
    var %in% c(
      "multifamily_units",
      "single_family_units",
      "single_family_average_floor_area_sqft_ctu",
      "multifamily_average_floor_area_sqft_county"
    )
  ) %>%
  tidyr::pivot_wider(names_from = year, values_from = value) %>%
  dplyr::select(ctu_name, var, `2018`, `2040`) %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.numeric, round, 2)
```

#### Strategy B.R7: Electrify Heating on Residential Buildings 

This strategy calculates reduction in energy use and GHG emissions due to electrifying heating in households (ex. switching to ground source heat pumps and tankless water heaters), assuming that by 2040, *59%* of households will use electric heating, based on the Net-Zero America report. Currently, *14%* of households already use electricity for heating, according to American Community Survey estimates (in Washington County). Therefore, a transition of *45%* of homes to electric heating (*59%*-*14%*) will have the modeled effects on GHG emissions. Note: In regions of Minnesota, *71%* of annual household gas is used for space heating, and *24%* of gas is used for water heating, according to the [Residential Energy Consumption Survey (RECS)](https://www.eia.gov/consumption/residential/data/2015/)

The main function involved in this strategy is `calc_electrify_residential_heating()`

```{r}
ghg.sp::calc_electrify_residential_heating(
  res_tb = calc_ghg_residential(
    res_tb = building_data$residential,
    res_tb_bau = building_data$residential,
    .grid_decarbonization_pct = 0.80,
    .enviro_factors = enviro_factors
  ),
  .additional_electrified_residential_buildings_pct = 0.45,
  .res_natural_gas_for_space_heating_pct = 0.71,
  .res_natural_gas_for_water_heating_pct = 0.24,
  .grid_decarbonization_pct = 0.80,
  .enviro_factors = enviro_factors
)
```

#### Strategy B.R8: Increase Renewable Natural Gas on Residential Buildings

For this strategy, the remaining homes that do not transition to electric space and water heating are assumed to switch to renewable natural gas, based on the State of Minnesota’s policy to aggregate all waste-to-energy project towards renewable natural gas. This includes municipal solid waste.

The main function involved in this strategy is `calc_residential_renewable_ng()`

*Note*: Strategy B.R8 is based on the outputs (if any) from strategy B.R7.
