---
title: "Building Energy Module"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{residential_building_energy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ghg.sp)
library(ggplot2)
```

# Source Data

All source tables are stored in the `Emissions` Microsoft SQL Database at the Metropolitan Council.
```{r}
names(building_energy_data)
```

# Business As Usual Data

The following functions don't run every time that the `run_scenario_building()` function is called by the end-user. Instead, they only need to be re-run if there is a change in the input tables that come from the database or there when there are changes to the code of such functions. The reason for this is to speed up the run-time. 

`run_scenario_building()` allows the user to specify the building energy scenario that they want to explore. 

Here is an example about what each table generates:

`get_demographic_baseline()` returns a table with the basic demographic characteristics of the cities/townships of the seven-county metropolitan region. 
The `$ctu` is used to show results at the city/township scales only. 
```{r}
get_demographic_baseline(tb = building_energy_data)$ctu %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

`get_demographic_baseline()` also can return a table with the basic demographic characteristics of counties in the metropolitan region. 
```{r}
get_demographic_baseline(tb = building_energy_data)$county %>%
  dplyr::filter(co_name == "Carver")
```

`calc_demographic_forecast()` returns a table with a forecast of the demographic characteristics of the cities/townships of the seven-county metropolitan region for the 2040. 
```{r}
calc_demographic_forecast(tb = building_energy_data)$ctu %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

`get_residential_energy_baseline()` returns a table with the available energy consumption data for cities/townships of the seven-county metropolitan region.  
```{r}
get_residential_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

`calc_residential_energy_forecast()` returns a table with a forecast of energy consumption for cities/townships of the seven-county metropolitan region. 
```{r}
calc_residential_energy_forecast(tb = building_energy_data) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

`get_by_ctu_non_residential_xcel_energy_baseline()` returns a table with non residential energy consumption data available for some (but not all) cities/townships of the seven-county metropolitan region. 
```{r}
get_by_ctu_non_residential_xcel_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

`get_by_county_non_residential_energy_baseline()` returns a table with non residential energy data avaiable for counties of the seven-county metropolitan region.  
```{r}
get_by_county_non_residential_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(co_name == "Carver")
```

`get_statewide_non_residential_energy` returns a table with non residential energy consumption data at the statewide scale. 
```{r}
get_statewide_non_residential_energy(tb = building_energy_data)
```

`get_non_residential_energy_baseline()` returns a table with non residential energy consumption data available for cities/townships in the seven-county metropolitan region.
```{r}
get_non_residential_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

`calc_non_residential_energy_forecast` returns a table with a forecast of non residential energy consumption data available for cities/townships in the seven-county metropolitan region. 
```{r}
calc_non_residential_energy_forecast(tb = building_energy_data) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```



# Residential Strategies

The following functions are not called individually. Instead, they are called in batch when using the function `run_scenario_building()`

```{r}
compile_bau_building_energy(tb = building_energy_data, ctu_selection = "Lake Elmo")$residential
```

### Calculate Affordable Floor Area

Cities can control the size of new single-family homes to reduce energy use. The growth of the single-family home size is reduced from 15% to 5% to make the new single-family home smaller. We apply this impact to 50% of new single-family homes. Cities can adjust the growth rate of single-family homes’ floor area and the percentage of new single-family homes being controlled.

`x` is the percent growth in single family floor area 

$$
S_1sqft = (units_{2040}^{sf} - units_{2018}^{sf}) \cdot x \cdot (\bar{sqft}_{2040}^{sf}- ( 1 + \frac{\Delta sqft}{year}) \cdot \bar{sqft}_{2018})
$$

```{r}
calc_affordable_floor_area(
  res_tb = building_energy_bau_data$residential,
  .single_family_floor_area_growth_pct = 0.05
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

### Calculate LEED floor area

Cities can adopt strict building codes requiring new homes to achieve LEED- Gold standards. We assumed 50% of new homes (single- and multi-family homes) to be LEED-Gold with both electricity and gas EUI reduced 64%. Cities can adjust the percentage of new homes achieving this standard.

`x` is the percent of new single family homes that are LEED certified.

$$
S_2sqft= (units_{2040}^{sf} - units_{2018}^{sf}) \cdot x \cdot (S_1sqft \cdot 0.64) \cdot \bar{sqft}_{2018})
$$

```{r}
calc_floor_area_leed(
  res_tb = building_energy_bau_data$residential,
  .new_homes_leed_gold_pct = 0.5,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

### Calculate Residential Floor Area Retrofit


```{r}
calc_floor_area_retrofit(
  res_tb = building_energy_bau_data$residential,
  .existing_home_retrofit_pct = 0.80,
  .existing_home_ultra_retrofit_pct = 0.20,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
calc_floor_area_behavior_change(
  res_tb = building_energy_bau_data$residential,
  .home_behavior_change_pct = 1.00,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

needs review
```{r}
calc_ghg_residential(
  res_tb = building_energy_bau_data$residential,
  res_tb_bau = building_energy_bau_data$residential,
  .grid_decarbonization_pct = 1,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")

building_energy_bau_data$residential %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

needs review
```{r}
calc_electrify_residential_heating(
  res_tb = calc_ghg_residential(
    res_tb = building_energy_bau_data$residential,
    res_tb_bau = building_energy_bau_data$residential,
    .grid_decarbonization_pct = 0.80,
    .enviro_factors = enviro_factors
  ),
  .additional_electrified_residential_buildings_pct = 0.45,
  .res_natural_gas_for_space_heating_pct = 0.71,
  .res_natural_gas_for_water_heating_pct = 0.24,
  .grid_decarbonization_pct = 0.80,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

not working
```{r}
calc_residential_renewable_ng(
  res_tb = calc_ghg_residential(
    res_tb = building_energy_bau_data$residential,
    res_tb_bau = building_energy_bau_data$residential,
    .grid_decarbonization_pct = 0.80,
    .enviro_factors = enviro_factors
  ),
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

# Non Residential Strategies

```{r}
compile_bau_building_energy(tb = building_energy_data)$non_residential %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
calc_existing_comm_building_efficiency(
  non_res_tb = building_energy_bau_data$non_residential,
  .existing_high_efficiency_buildings_pct = 0.80
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
calc_ghg_non_residential(
  non_res_tb = calc_existing_comm_building_efficiency(
    non_res_tb = building_energy_bau_data$non_residential,
    .existing_high_efficiency_buildings_pct = 0.80
  ),
  non_res_tb_bau = building_energy_bau_data$non_residential,
  .industrial_smart_grid_pct = 1,
  .commercial_smart_grid_pct = 1,
  .grid_decarbonization_pct = 0.8,
  .smart_grid_energy_reduction_pct = 1,
  .enviro_factors = enviro_factors,
  .existing_high_efficiency_buildings_pct = 0.8
)
```

```{r}
calc_electrify_commercial_heating(
  non_res_tb = calc_ghg_non_residential(
    non_res_tb = building_energy_bau_data$non_residential,
    non_res_tb_bau = building_energy_bau_data$non_residential,
    .industrial_smart_grid_pct = 1,
    .commercial_smart_grid_pct = 1,
    .grid_decarbonization_pct = 0.8,
    .smart_grid_energy_reduction_pct = 1,
    .enviro_factors = enviro_factors,
    .existing_high_efficiency_buildings_pct = 0.8
  ),
  .grid_decarbonization_pct = 0.8,
  .electrified_buildings_pct = 0.40,
  .non_res_natural_gas_for_water_heating_pct = 0.20,
  .non_res_natural_gas_for_space_heating_pct = 0.69,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
calc_non_res_renewable_ng(
  non_res_tb = calc_ghg_non_residential(
    non_res_tb = building_energy_bau_data$non_residential,
    non_res_tb_bau = building_energy_bau_data$non_residential,
    .industrial_smart_grid_pct = 1,
    .commercial_smart_grid_pct = 1,
    .grid_decarbonization_pct = 0.8,
    .smart_grid_energy_reduction_pct = 1,
    .enviro_factors = enviro_factors,
    .existing_high_efficiency_buildings_pct = 0.8
  ),
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```


```{r}
scen_building_non_residential(
  non_res_tb = ,
  non_res_tb_bau = building_energy_bau_data$non_residential,
  .electrified_buildings_pct = 0.40,
  .non_res_natural_gas_for_water_heating_pct = 0.20,
  .non_res_natural_gas_for_space_heating_pct = 0.69,
  .commercial_smart_grid_pct = 1.00,
  .industrial_smart_grid_pct = 1.00,
  .smart_grid_energy_reduction_pct = 1.00,
  .grid_decarbonization_pct = 0.80,
  .existing_high_efficiency_buildings_pct = 0.80,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

