---
title: "Building Energy Outputs - Non-residential"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{building_energy_module_outputs_non_residential}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ghg.sp)
library(tidyr)
```

------------------------------------------------------------------------

### Non-Residential Strategies

#### Strategy B.C1: Retrofit Existing Commercial Buildings

This strategy models the reduction in energy use intensity (EUI) that
will result if *80%* of existing commercial buildings use
performance-based energy efficiency strategies (many recent studies have
shown that performance-based assessment is better than LEED
certification), which is assumed to reduce EUI by *25%*-*50%*, based on
case studies from the Minnesota SB program.

The main function involved in this strategy is
`ghg.sp::calc_existing_comm_building_efficiency()`

$x$ = % of LEED Gold Buildings $q$ = number of commercial workers $r$ =
% assumed reduction in energy intensity $BY$ = Baseline Year $$
SB.C1 = (x_{BY} * q_{BY} * r)  \dfrac{MWh_{BY}}{q_{BY}}
$$

```{r}
ghg.sp::calc_existing_comm_building_efficiency(
  non_res_tb = building_data$non_residential,
  .existing_high_efficiency_buildings_pct = 0.80
) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  tidyr::pivot_wider(names_from = year, values_from = value) 
```

#### Strategy B.C2: Smart Grid Efficiency for Commercial Buildings

Achieving a net-zero grid requires transitioning to a smart grid. While
the smart grid often helps in managing the load, studies have also shown
that it can reduce energy use by approximately *11%*, based on PNNL's
"The Smart Grid: An Estimation of the Energy and CO<sub>2</sub>
Benefits." This *11%* will apply across the board to anyone connected to
the electricity grid. In this strategy and the next, we applied this
percentage reduction to *100%* of the electricity use. This strategy
applies to commercial buildings because the residential reductions are
already incorporated in the previous strategy.

The main function associated with this strategy
`ghg.sp::calc_ghg_non_residential()` The argument
'.industrial_smart_grid_pct', '.commercial_smart_grid_pct'

**Note**: All commercial and industrial energy use is computed on the
basis of employees, not square footage.

SB.c2 = CO2e in Tonnes x = % of commercial buildings in the smart grid q
= number of commercial workers r = % assumed reduction in energy
intensity

$$
SB.C2  = (x * q_{BY} * r) * (\dfrac{MWh_{BY}}{q_{BY}} - SB.C1) * EF_{electricity FY}
$$

```{r}
calc_ghg_non_residential(
  non_res_tb = building_data$non_residential,
  non_res_tb_bau = building_data$non_residential,
 .industrial_smart_grid_pct = 1,
 .commercial_smart_grid_pct = 1,
 .grid_decarbonization_pct = 1,
 .smart_grid_energy_reduction_pct = 1,
 .enviro_factors = enviro_factors,
 .existing_high_efficiency_buildings_pct = 0.8
) %>% 
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  tidyr::pivot_longer(cols = !ctu_name, names_to = "var")
```

#### Strategy B.C3: Smart Grid Efficiency for Industrial Buildings

Smart grid and technologies are projected to reduce industrial
electricity use by *11%*.

The main function associated with this strategy
`ghg.sp::calc_ghg_non_residential()`

SB.C3 = resulting electricity use from implementing the strategy q =
number of commercial workers r = % of industries in the smart grid x = %
reduction in energy due to smart grid $FY$ = forecast year

$$
SB.C3_{FY} = r_{FY} * q * \dfrac{MWh}{q_{FY}} * x * EF_{electricityFY}
$$

```{r}
calc_ghg_non_residential(
  non_res_tb = building_data$non_residential,
  non_res_tb_bau = building_data$non_residential,
  .industrial_smart_grid_pct = 1,
  .commercial_smart_grid_pct = 1,
  .grid_decarbonization_pct = 1,
  .smart_grid_energy_reduction_pct = 1,
  .enviro_factors = enviro_factors,
  .existing_high_efficiency_buildings_pct = 0.8
) %>%
  tidyr::pivot_longer(cols = !ctu_name) %>%
  tidyr::separate(name, sep = "\\.", into = c("var", "scen", "year") )  %>%
  tidyr::pivot_wider(names_from = year, values_from = value) 
```

#### Strategy B.C4: Electrifying Heating for Commercial Buildings

This equation calculates reduction in energy use and GHG emissions due
to electrifying heating in commercial buildings which is currently being
done with natural gas and propane (ex. switching to electricity-based
ground source heat pumps and tankless water heaters). Using CBECS data,
in the 2018 baseline, *70%* of commercial buildings in the Midwest are
using either gas or propane; approximately *19%* are primarily on
electricity for heating. We model that the *19%* will increase to *59%*,
similar to the residential assumption, in the Net-Zero America report.
Therefore, a transition of *40%* of commercial buildings to electric
heating (*59%*-*19%*) will have the modeled effects on GHG emissions. In
Minnesota, approximately *69%* of gas is used for space heating,
according to a Navigant Consulting report (2009).

The main function involved in this strategy is
`calc_electrify_commercial_heating()`

```{r}
calc_electrify_commercial_heating(
  non_res_tb = calc_ghg_non_residential(
    non_res_tb = building_data$non_residential,
    non_res_tb_bau = building_data$non_residential,
    .industrial_smart_grid_pct = 1,
    .commercial_smart_grid_pct = 1,
    .grid_decarbonization_pct = 0.8,
    .smart_grid_energy_reduction_pct = 1,
    .enviro_factors = enviro_factors,
    .existing_high_efficiency_buildings_pct = 0.8
  ),
  .grid_decarbonization_pct = 0.8,
  .electrified_buildings_pct = 0.40,
  .non_res_natural_gas_for_water_heating_pct = 0.20,
  .non_res_natural_gas_for_space_heating_pct = 0.69,
  .enviro_factors = enviro_factors
)
```

#### Strategy B.C5 Renewable Natual Gas

```{r}
calc_non_res_renewable_ng(
  non_res_tb = calc_ghg_non_residential(
    non_res_tb = building_data$non_residential,
    non_res_tb_bau = building_data$non_residential,
    .industrial_smart_grid_pct = 1,
    .commercial_smart_grid_pct = 1,
    .grid_decarbonization_pct = 0.8,
    .smart_grid_energy_reduction_pct = 1,
    .enviro_factors = enviro_factors,
    .existing_high_efficiency_buildings_pct = 0.8
  ),
  .enviro_factors = enviro_factors
) %>% 
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  tidyr::pivot_longer(cols = !ctu_name, names_to = "var")
```

### Running the Module

```{r}
ghg.sp::run_scenario_building(
  res_tb = building_data$residential,
  non_res_tb = building_data$non_residential,
  res_tb_bau = building_data$residential,
  non_res_tb_bau = building_data$non_residential,
  .enviro_factors = enviro_factors,
  .electrified_buildings_pct = 0.4,
  .non_res_natural_gas_for_water_heating_pct = 0.2,
  .non_res_natural_gas_for_space_heating_pct = 0.69,
  .commercial_smart_grid_pct =  0.5,
  .industrial_smart_grid_pct = 0.5,
  .smart_grid_energy_reduction_pct = 0, 
  .new_homes_to_multifamily_pct = 0.5,
  .existing_high_efficiency_buildings_pct = 0.8,
  .home_behavior_change_pct = 1,
  .single_family_floor_area_growth_pct = 0.05,
  .new_homes_affected_pct = 0.3,
  .new_homes_leed_gold_pct = 0.5,
  .existing_home_retrofit_pct = 0.8,
  .existing_home_ultra_retrofit_pct = 0.2,
  .res_natural_gas_for_space_heating_pct = 0.71,
  .res_natural_gas_for_water_heating_pct = 0.24,
  .additional_electrified_residential_buildings_pct = 0.45,
  .grid_decarbonization_pct = 0.5
) %>%  
  dplyr::filter(ctu_name == "Lake Elmo")
```
