---
title: "building_energy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{residential_building_energy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ghg.sp)
library(ggplot2)
```

# Source Data
all tables are stored in `Emissions`
```{r}
names(building_energy_data)
```

# Business As Usual Data

The following function don't run everytime that the ghg.sp::run_scenario_building() function is called. Instead, they only need to be re-run if there is a change in the input tables that come from the database. The reason for this is to speed up the run-time. 

Here is an example about what each table generates:

```{r}
ghg.sp::get_demographic_baseline(tb = building_energy_data)$ctu %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
ghg.sp::get_demographic_baseline(tb = building_energy_data)$county %>%
  dplyr::filter(co_name == "Carver")
```

```{r}
ghg.sp::calc_demographic_forecast(tb = building_energy_data)$ctu %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
ghg.sp::get_residential_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
ghg.sp::calc_residential_energy_forecast(tb = building_energy_data) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
ghg.sp::get_by_ctu_non_residential_xcel_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
ghg.sp::get_by_county_non_residential_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(co_name == "Carver")
```

```{r}
ghg.sp::get_statewide_non_residential_energy(tb = building_energy_data)
```

```{r}
ghg.sp::get_non_residential_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
ghg.sp::calc_non_residential_energy_forecast(tb = building_energy_data) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```



# Residential Strategies

The following functions are not called individually. Instead, they are called in batch when using the function `ghg.sp::run_scenario_building()`

```{r}
ghg.sp::compile_bau_building_energy(tb = building_energy_data, ctu_selection = "Lake Elmo")$residential
```

```{r}
ghg.sp::calc_affordable_floor_area(
  res_tb = building_energy_bau_data$residential,
  .single_family_floor_area_growth_pct = 0.05
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
ghg.sp::calc_floor_area_leed(
  res_tb = building_energy_bau_data$residential,
  .new_homes_leed_gold_pct = 0.5,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```


```{r}
ghg.sp::calc_floor_area_retrofit(
  res_tb = building_energy_bau_data$residential,
  .existing_home_retrofit_pct = 0.80,
  .existing_home_ultra_retrofit_pct = 0.20,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
ghg.sp::calc_floor_area_behavior_change(
  res_tb = building_energy_bau_data$residential,
  .home_behavior_change_pct = 1.00,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

needs review
```{r}
ghg.sp::calc_ghg_residential(
  res_tb = building_energy_bau_data$residential,
  res_tb_bau = building_energy_bau_data$residential,
  .grid_decarbonization_pct = 1,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")

building_energy_bau_data$residential %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

needs review
```{r}
ghg.sp::calc_electrify_residential_heating(
  res_tb = calc_ghg_residential(
    res_tb = building_energy_bau_data$residential,
    res_tb_bau = building_energy_bau_data$residential,
    .grid_decarbonization_pct = 0.80,
    .enviro_factors = enviro_factors
  ),
  .additional_electrified_residential_buildings_pct = 0.45,
  .res_natural_gas_for_space_heating_pct = 0.71,
  .res_natural_gas_for_water_heating_pct = 0.24,
  .grid_decarbonization_pct = 0.80,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

not working
```{r}
ghg.sp::calc_residential_renewable_ng(
  res_tb = calc_ghg_residential(
    res_tb = building_energy_bau_data$residential,
    res_tb_bau = building_energy_bau_data$residential,
    .grid_decarbonization_pct = 0.80,
    .enviro_factors = enviro_factors
  ),
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

# Non Residential Strategies

```{r}
ghg.sp::compile_bau_building_energy(tb = building_energy_data)$non_residential %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
ghg.sp::calc_existing_comm_building_efficiency(
  non_res_tb = building_energy_bau_data$non_residential,
  .existing_high_efficiency_buildings_pct = 0.80
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
ghg.sp::calc_ghg_non_residential(
  non_res_tb = ghg.sp::calc_existing_comm_building_efficiency(
    non_res_tb = building_energy_bau_data$non_residential,
    .existing_high_efficiency_buildings_pct = 0.80
  ),
  non_res_tb_bau = building_energy_bau_data$non_residential,
  .industrial_smart_grid_pct = 1,
  .commercial_smart_grid_pct = 1,
  .grid_decarbonization_pct = 0.8,
  .smart_grid_energy_reduction_pct = 1,
  .enviro_factors = enviro_factors,
  .existing_high_efficiency_buildings_pct = 0.8
)
```

```{r}
ghg.sp::calc_electrify_commercial_heating(
  non_res_tb = calc_ghg_non_residential(
    non_res_tb = building_energy_bau_data$non_residential,
    non_res_tb_bau = building_energy_bau_data$non_residential,
    .industrial_smart_grid_pct = 1,
    .commercial_smart_grid_pct = 1,
    .grid_decarbonization_pct = 0.8,
    .smart_grid_energy_reduction_pct = 1,
    .enviro_factors = enviro_factors,
    .existing_high_efficiency_buildings_pct = 0.8
  ),
  .grid_decarbonization_pct = 0.8,
  .electrified_buildings_pct = 0.40,
  .non_res_natural_gas_for_water_heating_pct = 0.20,
  .non_res_natural_gas_for_space_heating_pct = 0.69,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

```{r}
ghg.sp::calc_non_res_renewable_ng(
  non_res_tb = calc_ghg_non_residential(
    non_res_tb = building_energy_bau_data$non_residential,
    non_res_tb_bau = building_energy_bau_data$non_residential,
    .industrial_smart_grid_pct = 1,
    .commercial_smart_grid_pct = 1,
    .grid_decarbonization_pct = 0.8,
    .smart_grid_energy_reduction_pct = 1,
    .enviro_factors = enviro_factors,
    .existing_high_efficiency_buildings_pct = 0.8
  ),
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```


```{r}
ghg.sp::scen_building_non_residential(
  non_res_tb = ,
  non_res_tb_bau = building_energy_bau_data$non_residential,
  .electrified_buildings_pct = 0.40,
  .non_res_natural_gas_for_water_heating_pct = 0.20,
  .non_res_natural_gas_for_space_heating_pct = 0.69,
  .commercial_smart_grid_pct = 1.00,
  .industrial_smart_grid_pct = 1.00,
  .smart_grid_energy_reduction_pct = 1.00,
  .grid_decarbonization_pct = 0.80,
  .existing_high_efficiency_buildings_pct = 0.80,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

