---
title: "Building Energy Module"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{residential_building_energy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ghg.gert)
library(ggplot2)
```

# Source Data

All source tables are stored in the `Emissions` Microsoft SQL Database at the Metropolitan Council.
```{r}
names(building_energy_data)
```

# Business As Usual Data

The following functions don't run every time that the `run_scenario_building()` function is called by the end-user. Instead, they only need to be re-run if there is a change in the input tables that come from the database or there when there are changes to the code of such functions. The reason for this is to speed up the run-time. 

`run_scenario_building()` allows the user to specify the building energy scenario that they want to explore. 

Here is an example about what each table generates:

### Demographic Baseline

#### City/Township Demographics

`get_demographic_baseline()` returns a table with the basic demographic characteristics of the cities/townships of the seven-county metropolitan region. 
The `$ctu` is used to show results at the city/township scales only. 
```{r}
get_demographic_baseline(tb = building_energy_data)$ctu %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

#### County Demographics

`get_demographic_baseline()` can also return a table with the basic demographic characteristics of counties in the metropolitan region. 
```{r}
get_demographic_baseline(tb = building_energy_data)$county %>%
  dplyr::filter(co_name == "Carver")
```


### Calculate Demographic Forecast

Calculates demographic forecast characteristics by CTU.

`calc_demographic_forecast()` returns a table with a forecast of the demographic characteristics of the cities/townships of the seven-county metropolitan region for the 2040. 
```{r}
calc_demographic_forecast(tb = building_energy_data)$ctu %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

### Get Residential Energy Baseline

`get_residential_energy_baseline()` returns a table with the available energy consumption data for cities/townships of the seven-county metropolitan region.  
```{r}
get_residential_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

### Calculate City/Township Residential Energy Forecast

`calc_residential_energy_forecast()` returns a table with a forecast of energy consumption for cities/townships of the seven-county metropolitan region. 
```{r}
calc_residential_energy_forecast(tb = building_energy_data) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

### Get City/Township Non Residential Xcel Energy Baseline

`get_by_ctu_non_residential_xcel_energy_baseline()` returns a table with non residential energy consumption data available for some (but not all) cities/townships of the seven-county metropolitan region. 
```{r}
get_by_ctu_non_residential_xcel_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

### Get County Non Residential Energy Baseline

`get_by_county_non_residential_energy_baseline()` returns a table with non residential energy data avaiable for counties of the seven-county metropolitan region.  
```{r}
get_by_county_non_residential_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(co_name == "Carver")
```

### Get Statewide Non Residential Energy 

`get_statewide_non_residential_energy` returns a table with non residential energy consumption data at the statewide scale. 
```{r}
get_statewide_non_residential_energy(tb = building_energy_data)
```

### Get City/Township Non Residential Energy Baseline

`get_non_residential_energy_baseline()` returns a table with non residential energy consumption data available for cities/townships in the seven-county metropolitan region.
```{r}
get_non_residential_energy_baseline(tb = building_energy_data) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

`calc_non_residential_energy_forecast` returns a table with a forecast of non residential energy consumption data available for cities/townships in the seven-county metropolitan region. 
```{r}
calc_non_residential_energy_forecast(tb = building_energy_data) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```



# Residential Strategies

The following functions are not called individually by the user, with the exception of special cases. 
Instead, they are called all at once in sequence when using the function `run_scenario_building()`.

```{r}
compile_bau_building_energy(tb = building_energy_data, ctu_selection = "Lake Elmo")$residential
```

### Calculate Affordable Floor Area

Cities can control the size of new single-family homes to reduce energy use. The growth of the single-family home size is reduced from 15% to 5% to make the new single-family home smaller. We apply this impact to 50% of new single-family homes. Cities can adjust the growth rate of single-family homes’ floor area and the percentage of new single-family homes being controlled.

`X_1` is the percent growth in single family floor area 

$$
S_1sqft_{2040}^{sf} = (units_{2040}^{sf} - units_{2018}^{sf}) \cdot X_1 \cdot (\bar{sqft}_{2040}^{sf}- ( 1 + \frac{\Delta sqft}{year}) \cdot \bar{sqft}_{2018})
$$
```{r}
calc_affordable_floor_area(
  res_tb = building_energy_bau_data$residential,
  .single_family_floor_area_growth_pct = 0.05
) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

### Calculate LEED floor area

Cities can adopt strict building codes requiring new homes to achieve LEED- Gold standards. We assumed `50%` of new homes (single- and multi-family homes) to be LEED-Gold with both electricity and gas EUI reduced `64%`. Cities can adjust the percentage of new homes achieving this standard.

`x` is the percent of new single family homes that are LEED certified.

$$
S_2sqft_{2040}^{sf} = (units_{2040}^{sf} - units_{2018}^{sf}) \cdot x \cdot (S_1sqft \cdot 0.64 \cdot \bar{sqft}_{2018}^{sf})
$$

```{r}
calc_floor_area_leed(
  res_tb = building_energy_bau_data$residential,
  .new_homes_leed_gold_pct = 0.5,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

### Calculate Residential Floor Area Retrofit

Cities can retrofit existing homes to be energy efficient. We assume 80% of existing homes with energy use intensity reduced by 33%. Cities can adjust the percentage of existing homes to be retrofitted.

`X_1` is ...

`X_2` is ...

$$
S_3sqft = (units_{2018}^{sf} \cdot X_1 \cdot 0.33 \cdot \bar{sqft}_{2018}^{sf} ) + (units_{2018}^{mf} \cdot X_2 \cdot 0.33 \cdot \bar{sqft}_{2018}^{mf})
$$

```{r}
calc_floor_area_retrofit(
  res_tb = building_energy_bau_data$residential,
  .existing_home_retrofit_pct = 0.80,
  .existing_home_ultra_retrofit_pct = 0.20,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

### Calculate Floor Area Behavior Change

Efficient messaging, in-home energy display, and smart meters together can reduce household energy use by 11%. The default is 100% of homes change behaviors due to these actions. Cities can adjust this participation rate accordingly.

`X_1` is the percent of homes that change behavior

$$
S_4sqft = X_1 \cdot population_{2040} \cdot S_3sqft/person_{2040}^{residential} \cdot 0.11
$$
```{r}
calc_floor_area_behavior_change(
  res_tb = building_energy_bau_data$residential,
  .home_behavior_change_pct = 1.00,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

### Calculate GHG Residential

Due to decarbonizing the grid, the emission factor of electricity decreases from 568 kgCO2e/MWh (2018 Minnesota state level) to 113 kgCO2e/MWh. This change is fixed by the grid.

`X_1` is the percent reduction in the grid emissions factor.

$$
S_5kgCO_2e = (population_{2040} \cdot S_4sqft/person_{2040}^{residential} \cdot kWh/sqft_{residential}^{2040}) \cdot [(kg CO_2e/MWh_{2040}^{electricity} \cdot 1 \times 10^{-3} kWh) \cdot X_1]
$$
```{r}
calc_ghg_residential(
  res_tb = building_energy_bau_data$residential,
  res_tb_bau = building_energy_bau_data$residential,
  .grid_decarbonization_pct = 0.8,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(
    geog_name == "Lake Elmo",
    year == 2040,
    var %in% c(
      "residential_electricity_emissions_kg_co",
      "residential_natural_gas_emissions_kg_co"
    )
  )
```

### Calculate Electrify Residential Heating

59% of homes are using air-source heat pumps and electric water heaters, instead of using gas for heating. 
Cities can adjust the percentage of homes participated in electrifying heating.

`X_1` is the percent of additional households with heating electrified
`0.71` is the percent of homes using gas for space heating

$$
S_6kgCO_2e = (population_{2040} \cdot S_4sqft/person_{2040}^{residential} \cdot thems/sqft_{2040}^{residential}) \cdot X_1 \cdot kgCo_2e/therm \cdot 0.71 
$$
```{r}
calc_electrify_residential_heating(
  res_tb = calc_ghg_residential(
    res_tb = building_energy_bau_data$residential,
    res_tb_bau = building_energy_bau_data$residential,
    .grid_decarbonization_pct = 0.80,
    .enviro_factors = enviro_factors
  ),
  .additional_electrified_residential_buildings_pct = 0.45,
  .grid_decarbonization_pct = 0.80,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(
    geog_name == "Lake Elmo",
    year == 2040,
    var %in% c(
      "residential_electricity_emissions_kg_co",
      "residential_natural_gas_emissions_kg_co"
    )
  )
```

### Calculate Residential Renewable Natural Gas

The remaining homes that did not transition to electric space and water heating are assumed to have switched to renewable natural gas, based on the State of Minnesota’s policy to aggregate all waste-to-energy project towards renewable natural gas. This includes municipal solid waste, agricultural waste, etc., converted to biogas, which is assumed to be carbon neutral.

$$ S_{7}kgCo_2e = (population_{2040} \cdot S_4sqft/person_{2040}^{residential} \cdot thems/sqft_{2040}^{residential}) \cdot 0.317 \cdot kgCo_2e/therm \cdot 0.71$$ 

```{r}
calc_residential_renewable_ng(
  res_tb = calc_ghg_residential(
    res_tb = building_energy_bau_data$residential,
    res_tb_bau = building_energy_bau_data$residential,
    .grid_decarbonization_pct = 0.80,
    .enviro_factors = enviro_factors
  ),
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

# Non Residential Strategies

```{r}
compile_bau_building_energy(tb = building_energy_data)$non_residential %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

### Calculate Existing Commercial Buildings Efficiency


```{r}
calc_existing_comm_building_efficiency(
  non_res_tb = building_energy_bau_data$non_residential,
  .existing_high_efficiency_buildings_pct = 0.80
) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

### Calculate GHG Non Residential



```{r}
calc_ghg_non_residential(
  non_res_tb = calc_existing_comm_building_efficiency(
    non_res_tb = building_energy_bau_data$non_residential,
    .existing_high_efficiency_buildings_pct = 0.80
  ),
  non_res_tb_bau = building_energy_bau_data$non_residential,
  .grid_decarbonization_pct = 0.8,
  .smart_grid_energy_reduction_pct = 1,
  .enviro_factors = enviro_factors,
  .existing_high_efficiency_buildings_pct = 0.8
)
```

```{r}
calc_electrify_commercial_heating(
  non_res_tb = calc_ghg_non_residential(
    non_res_tb = building_energy_bau_data$non_residential,
    non_res_tb_bau = building_energy_bau_data$non_residential,
    .grid_decarbonization_pct = 0.8,
    .smart_grid_energy_reduction_pct = 1,
    .enviro_factors = enviro_factors,
    .existing_high_efficiency_buildings_pct = 0.8
  ),
  .grid_decarbonization_pct = 0.8,
  .electrified_buildings_pct = 0.40,
  .non_res_natural_gas_for_water_heating_pct = 0.20,
  .non_res_natural_gas_for_space_heating_pct = 0.69,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

```{r}
calc_non_res_renewable_ng(
  non_res_tb = calc_ghg_non_residential(
    non_res_tb = building_energy_bau_data$non_residential,
    non_res_tb_bau = building_energy_bau_data$non_residential,
    .grid_decarbonization_pct = 0.8,
    .smart_grid_energy_reduction_pct = 1,
    .enviro_factors = enviro_factors,
    .existing_high_efficiency_buildings_pct = 0.8
  ),
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```


```{r}
scen_building_non_residential(
  non_res_tb = ,
  non_res_tb_bau = building_energy_bau_data$non_residential,
  .electrified_buildings_pct = 0.40,
  .non_res_natural_gas_for_water_heating_pct = 0.20,
  .non_res_natural_gas_for_space_heating_pct = 0.69,
  .smart_grid_energy_reduction_pct = 1.00,
  .grid_decarbonization_pct = 0.80,
  .existing_high_efficiency_buildings_pct = 0.80,
  .enviro_factors = enviro_factors
) %>%
  dplyr::filter(geog_name == "Lake Elmo")
```

