#### Increased Multifamily

Compact community development can reduce building energy use by increasing the proportion of multifamily homes in the housing stock. Multifamily homes have a smaller floor area than single-family homes[@moura2015120], which can translate to energy use reductions [@goldstein2020carbon]. 

The Metropolitan Council’s Thrive MSP plan projects an increase in the percentage of multifamily housing by 2040. This is consistent with the Minneapolis 2040 plan move to eliminate single-family-only zoning areas.

::: about_strategy
***About this strategy*** Here, we model that `50%` of all new single-family construction will become multifamily, with the multifamily homes having smaller square footage.  
:::

Please note that the Metropolitan Council’s projection to 2040 already assumed a significant compact development.

```{r tab-increased-multifamily, message=FALSE, warning=FALSE, include=FALSE}
tab_increased_multifamily <- calc_affordable_floor_area(
  res_tb = building_data$residential,
  .selected_ctu = params$ctu_selection,
  .single_family_floor_area_growth_pct = 0
) %>%
  filter(
    year == 2018,
    var %in% c(
      "multifamily_average_floor_area_sqft_county",
      "multifamily_units",
      "single_family_units",
      "single_family_average_floor_area_sqft_ctu"
    )
  ) %>%
  add_variable_names(., "2018 Baseline") %>%
  full_join(
    calc_affordable_floor_area(
      res_tb = building_data$residential,
      .selected_ctu = params$ctu_selection,
      .single_family_floor_area_growth_pct = 0.15
    ) %>%
      filter(
        year == 2040,
        var %in% c(
          "multifamily_average_floor_area_sqft_county",
          "multifamily_units",
          "single_family_units",
          "single_family_average_floor_area_sqft_ctu"
        )
      ) %>%
      add_variable_names(., "2040 Scenario"),
    by = "Variables"
  ) %>%
  tableformat(.,
    .caption = paste(
      "New Single-Family Home Floor Area for",
      params$ctu_selection
    )
  )
```

```{r echo=FALSE}
tab_increased_multifamily
```


<br>

<!-- **GHG Impact Calculation** -->

<!-- Calculate floor area changes: -->

<!-- $$ \text{new }sqft_{2040}^{sf} = units_{2040}^{sf} - units_{2018}^{sf} * x * (\bar{sqft}_{2040}^{sf} - \bar{sqft}_{2040}^{mf})$$ -->
<!-- *Where:* -->

<!-- -  $x\text{, where } 0 \le x \le 1$:  The percentage of new single-family homes that are converted to multifamily homes -->
<!-- -  $units_{2018}^{sf}$: The number of single-family housing units in the community in 2018 -->
<!-- -  $units_{2040}^{sf}$: The projected number of single-family housing units in the community in 2040 -->
<!-- -  $sqft_{2040}^{sf}$: The projected total floor area of single-family homes in the community in 2040 -->
<!-- -  $\bar{sqft}_{2040}^{sf}$: The projected average floor area of single-family homes in the community in 2040 -->
<!-- -  $\bar{sqft}_{2040}^{mf}$: The projected  average floor area of multifamily homes in the community in 2040 -->

<!-- Calculate new emissions: -->

<!-- $$\text{new total }kgCO_2e^{sf}_{2040} = \text{new }sqft_{2040}^{sf} * \left( \frac{MWh_{2040}^{residential}}{sqft_{2040}^{residential}} * \frac{kgCO2e}{MWh_{2040}} + \frac{therm_{2040}^{residential}}{sqft_{2040}^{residential}} * \frac{kgCO2e}{therm_{2040}} \right)$$ -->

<!-- *Where:* -->

<!-- -  $MWh_{2040}^{residential}/sqft_{2040}^{residential}$: Projected residential electricity use intensity per floor area in 2040 -->
<!-- -  $therm_{2040}^{residential}/sqft_{2040}^{residential}$: Projected residential natural gas use intensity per floor area in 2040 -->
<!-- -  $kgCO2e/MWh$ = The projected electricity emission factor in 2040 -->
<!-- -  $kgCO_2e/therms_{2040}$: The projected emissions factor of natural gas in 2040 -->

<!-- <br> -->

```{r fig03a1, message=FALSE, warning=FALSE, include=FALSE}
fig03a1_data <-
  res_bau_2018 %>%
  full_join(
    res_bau_2040,
    by = "Variables"
  ) %>%
  full_join(
    run_scenario_building(
      res_tb = building_energy_bau_data$residential,
      res_tb_bau = building_energy_bau_data$residential,
      run_residential = TRUE,
      run_non_residential = FALSE,
      .selected_ctu = params$ctu_selection,
      # affordable floor area
      .single_family_floor_area_growth_pct = 0.15,
      .new_homes_affected_pct = 1,
      # other variables
      .electrified_buildings_pct = 0,
      .new_homes_to_multifamily_pct = 0.50,
      .existing_high_efficiency_buildings_pct = 0,
      .home_behavior_change_pct = 0,
      .new_homes_leed_gold_pct = 0,
      .existing_home_retrofit_pct = 0,
      .existing_home_ultra_retrofit_pct = 0,
      .additional_electrified_residential_buildings_pct = 0,
      .grid_decarbonization_pct = 0,
      .enviro_factors = enviro_factors
    ) %>%
      filter(
        var %in% c(
          "residential_mwh",
          "residential_electricity_emissions_kg_co",
          "residential_therms",
          "residential_natural_gas_emissions_kg_co"
        ),
        year == 2040,
        scen == "scen"
      ) %>%
      ungroup() %>%
      add_variable_names(., "2040 Scenario"),
    by = "Variables"
  )

fig03a1_table <- fig03a1_data %>% tableformat(.,
  .caption = paste(
    "New Single-Family Home Floor Area for",
    params$ctu_selection
  )
)
```

```{r echo=FALSE}
fig03a1_table
```

```{r echo=FALSE}
chartformat(fig03a1_data)
```

**Affordability co-benefit**: With electricity costs projected to nearly double by 2040, smaller floor area is likely to be more affordable for families in all types of housing. However, single-family floor areas in the Twin Cities are increasing. Therefore, the next section also demonstrates the effect of more affordable, smaller floor area, even in new single-family construction, going forward. 
