#### New Single Family Homes Energy Efficiency

*LEED* (Leadership in Energy and Environmental Design) is a green building rating system that is used to evaluate the sustainability of buildings. Buildings that meet *LEED* standards are designed and constructed to be energy efficient and to use fewer natural resources.

The amount of energy savings that can be achieved through *LEED* certifications will depend on the specific design and construction practices that are used, as well as the location and climate of the building. According to the [U.S. Green Building Council](https://www.usgbc.org/leed/rating-systems/residential), buildings that are designed to meet *LEED* Gold standards typically achieve energy savings of around 25% compared to conventional buildings.

The Greenhouse Gas Scenario Planning Tool uses *LEED Gold* for residential buildings as an example of how building standard could reduce energy use and therefore carbon emissions.

::: about_strategy
***About this strategy*** Our main assumption is that cities can adjust the percentage of new single family homes achieving a standard like LEED. Cities can adopt strict building codes requiring new homes to achieve LEED Gold (or similar) standards. We assumed `50%` of new single-family homes to be LEED-Gold with both electricity and gas energy use intensity reduced `64%`.
:::

```{r tab-new-home-efficiency, message=FALSE, warning=FALSE, include=FALSE}
#| echo: false
#| message: false

fig_tab_new_home_efficiency <- calc_floor_area_leed(
  res_tb = building_data$residential,
  .selected_ctu = params$ctu_selection,
  .new_homes_leed_gold_pct = 0,
  .enviro_factors = enviro_factors
) %>%
  ungroup() %>%
  filter(
    year == 2018,
    var %in% c(
      "multifamily_average_floor_area_sqft_county",
      "multifamily_units",
      "single_family_units",
      "single_family_average_floor_area_sqft_ctu"
    )
  ) %>%
  add_variable_names(., "2018 Baseline") %>%
  full_join(
    calc_floor_area_leed(
      res_tb = building_data$residential,
      .selected_ctu = params$ctu_selection,
      .new_homes_leed_gold_pct = 0,
      .enviro_factors = enviro_factors
    ) %>%
      ungroup() %>%
      filter(
        year == 2040,
        var %in% c(
          "multifamily_average_floor_area_sqft_county",
          "multifamily_units",
          "single_family_units",
          "single_family_average_floor_area_sqft_ctu"
        )
      ) %>%
      add_variable_names(., "2040 Scenario"),
    by = "Variables"
  ) %>%
  tableformat(.,
    .caption = paste("New Home Efficiency", params$ctu_selection, "2040 Assumptions")
  )
```

```{r echo=FALSE}
fig_tab_new_home_efficiency
```

<br>

**GHG Impact Calculation**

Calculate new floor area:

$$\text{new } sqft_{2040}^{sf} = x * (units_{2040}^{sf}-units_{2018}^{sf}) * \bar{sqft}_{2040}^{sf}) * 0.64$$

*Where:*
 
-   $x \text{, where } 0 \le x \le 1$: The percentage of new single family homes that are LEED Gold
-   $sqft_{2040}^{sf}$: The total floor area of single-family homes in the community in 2040
-   $units_{2040}^{sf}$: The number of single-family housing units in the community in 2040
-   $units_{2018}^{sf}$: The number of single-family housing units in the community in 2018
-   $\bar{sqft}_{2040}^{sf}$: The average floor area of single-family homes in the community in 2040
-   $0.64$: This factor represents the reduction in energy use intensity due to highly-efficient LEED Gold construction

Calculate new emissions:

$$\text{new total }kgCO_2e^{sf}_{2040} = \text{new }sqft_{2040}^{sf} * (MWh_{2040}^{residential}/sqft_{2040}^{residential} * kgCO2e/MWh_{2040} + therm_{2040}^{residential}/sqft_{2040}^{residential} * kgCO2e/therm_{2040})$$

*Where:*

-  $MWh_{2040}^{residential}/sqft_{2040}^{residential}$: The projected residential electricity use intensity per floor area in 2040
-  $therm_{2040}^{residential}/sqft_{2040}^{residential}$: The projected residential natural gas use intensity per floor area in 2040
-  $kgCO2e/MWh$ = The projected electricity emission factor in 2040
-  $kgCO_2e/therms_{2040}$: The projected emissions factor of natural gas in 2040

```{r message=FALSE, warning=FALSE, include=FALSE}
fig03a3_data <-
  res_bau_2018 %>%
  full_join(
    res_bau_2040,
    by = "Variables"
  ) %>%
  full_join(
    run_scenario_building(
      res_tb = building_energy_bau_data$residential,
      res_tb_bau = building_energy_bau_data$residential,
      run_residential = TRUE,
      run_non_residential = FALSE,
      .selected_ctu = params$ctu_selection,
      # affordable floor area
      .single_family_floor_area_growth_pct = 0.15,
      .new_homes_affected_pct = 1,
      # other variables
      .electrified_buildings_pct = 0,
      .new_homes_to_multifamily_pct = 0,
      .existing_high_efficiency_buildings_pct = 0,
      .home_behavior_change_pct = 0,
      .new_homes_leed_gold_pct = 0.5,
      .existing_home_retrofit_pct = 0,
      .existing_home_ultra_retrofit_pct = 0,
      .additional_electrified_residential_buildings_pct = 0,
      .grid_decarbonization_pct = 0,
      .enviro_factors = enviro_factors
    ) %>%
      filter(
        var %in% c(
          "residential_mwh",
          "residential_electricity_emissions_kg_co",
          "residential_therms",
          "residential_natural_gas_emissions_kg_co"
        ),
        year == 2040,
        scen == "scen"
      ) %>%
      ungroup() %>%
      add_variable_names(., "2040 Scenario"),
    by = "Variables"
  )

fig03a3_table <- fig03a3_data %>% tableformat(.,
  .caption = paste("New Home Efficiency", params$ctu_selection, "2040")
)
```


```{r echo=FALSE}
fig03a3_table
```

```{r echo=FALSE}
chartformat(fig03a3_data)
```
