#### Decarbonization of Residential Energy Use

The energy grid could become cleaner by increasing the use of renewable energy sources such as solar, wind, and hydroelectric power, which produce significantly fewer greenhouse gas emissions than fossil fuels

::: about_strategy
***About this strategy***: Due to decarbonizing the grid, the emission factor of electricity decreases from `568` kg<sub>2</sub>e/MWh (2018 Minnesota state level) to `113` kgCO<sub>2</sub>e/MWh. This change is fixed by the grid.
:::

<br>

**GHG Impact Calculation**

Calculate electricity use in 2040:

$$
\text{Busines as Usual }MWh_{2040}^{residential} = pop_{2040} * \frac{sqft_{2040}^{residential}}{pop_{2040}} * \frac{MWh}{sqft_{2040}^{residential}}
$$

*Where:*

-  $pop_{2040}$: The projected population in 2040
-  $sqft_{2040}^{residential}/pop_{2040}$: The projected residential floor area per capita in 2040
-  $MWh/sqft_{2040}^{residential}$: The projected residential electricity per floor area in 2040


Calculate new emissions:

$$\text{new electricity }kgCO_2e^{residential}_{2040} = \text{Busines as Usual }MWh_{2040}^{residential} *  kgCO2e/MWh_{2040} * (1-x) $$

*Where:*

-  $(1-x) \text{, where } 0 \le x \le 1$: The percentage reduction in the emissions of the electric grid
-  $kgCO2e/MWh$ = The projected electricity emission factor in 2040


```{r tab-decarbonize-residential-energy, message=FALSE, warning=FALSE, include=FALSE}

fig03a6_data <-   
  res_bau_2018 %>%
  full_join(
    res_bau_2040,
    by = "Variables"
  ) %>%
  full_join(
    run_scenario_building(
      res_tb = building_energy_bau_data$residential,
      res_tb_bau = building_energy_bau_data$residential,
      run_residential = TRUE,
      run_non_residential = FALSE,
      .selected_ctu = params$ctu_selection,
      # affordable floor area
      .single_family_floor_area_growth_pct = 0.15,
      .new_homes_affected_pct = 1,
      # other variables
      .electrified_buildings_pct = 0,
      .new_homes_to_multifamily_pct = 0,
      .existing_high_efficiency_buildings_pct = 0,
      .home_behavior_change_pct = 0,
      .new_homes_leed_gold_pct = 0,
      .existing_home_retrofit_pct = 0,
      .existing_home_ultra_retrofit_pct = 0,
      .additional_electrified_residential_buildings_pct = 0,
      .grid_decarbonization_pct = 0.8,
      .enviro_factors = enviro_factors
    )  %>%
      filter(
        var %in% c(
          "residential_mwh",
          "residential_electricity_emissions_kg_co",
          "residential_therms",
          "residential_natural_gas_emissions_kg_co"
        ),
        year == 2040,
        scen == "scen"
      ) %>%
      ungroup() %>%
      add_variable_names(., "2040 Scenario"),
    by = "Variables")

fig03a6_table <- 
  fig03a6_data %>%  
  tableformat(., .caption = "Decarbonize Residential Energy Use Minneapolis 2040")
```

```{r echo=FALSE}
fig03a6_table
```

```{r echo=FALSE}
chartformat(fig03a6_data)
```
