#### Electrification of Residential Heating

The Greenhouse Gas Scenario Planning Tool estimates that by 2040, Minnesota will decarbonize the overall grid to 80% relative to 2018 levels, consistent with Xcel Energy's plans to achieve a net-zero power grid by 2050.

Switching to more efficient electric heat pumps and tank-less water heaters with a net zero-carbon future electric grid can significantly reduce emissions.

Only around 14% of homes in MN have adopted these new technologies. The scenario planning tool projects that by 2050, 59% of residential and commercial buildings will switch to electrically driven heat pumps, which can heat and cool homes while offering long-term cost savings. Furthermore, the scenario planning model utilizes efficiency of the heat pump derived from pilot tests in Minnesota, so the results are more realistic for our region. National studies indicate 100% adoption of electric heat pumps may be unlikely in cold areas, due to which we also model renewable natural gas for the remaining buildings, described in the following strategy.

::: about_strategy
**About this strategy** For this strategy, we made the assumption that by 2040 `59%` of homes are using air-source heat pumps and electric water heaters, instead of using gas for heating. Cities can adjust the percentage of homes participated in electrifying heating.
:::

```{r tab-electrify-residential-heating, message=FALSE, warning=FALSE, include=FALSE}
fig03a7_data <- res_bau_2018 %>%
  full_join(res_bau_2040,
            by = "Variables") %>%
  full_join(
    run_scenario_building(
      res_tb = building_energy_bau_data$residential,
      res_tb_bau = building_energy_bau_data$residential,
      run_residential = TRUE,
      run_non_residential = FALSE,
      .selected_ctu = "Minneapolis",
      ## affordable floor area
      .single_family_floor_area_growth_pct = 0.15,
      .new_homes_affected_pct = 1,
      # other variables
      .electrified_buildings_pct = 0,
      .new_homes_to_multifamily_pct = 0,
      .existing_high_efficiency_buildings_pct = 0,
      .home_behavior_change_pct = 0,
      .new_homes_leed_gold_pct = 0,
      .existing_home_retrofit_pct = 0,
      .existing_home_ultra_retrofit_pct = 0,
      .additional_electrified_residential_buildings_pct = 0.45,
      .res_natural_gas_for_space_heating_pct = 0.71,
      .res_natural_gas_for_water_heating_pct = 0.24,
      .grid_decarbonization_pct = 0,
      .enviro_factors = enviro_factors
    )  %>%
      filter(
        var %in% c(
          "residential_mwh",
          "residential_electricity_emissions_kg_co",
          "residential_therms",
          "residential_natural_gas_emissions_kg_co"
        ),
        year == 2040,
        scen == "scen"
      ) %>%
      ungroup() %>%
      add_variable_names(., "2040 Scenario"),
    by = "Variables"
  )

fig03a7_table <-
  fig03a7_data %>%
  tableformat(., .caption = "Electrify Residential Heating 2040")
```

```{r echo=FALSE}
fig03a7_table
```

Note that this table shows the impact of electrification before considering the decarbonization of the electric grid. 
