
# Process census data

```{r setup}
library(tidyverse)
# library(fs)
library(sf)
library(tigris)
library(janitor)
library(leaflet)
`%not_in%` <- Negate(`%in%`)
```

## Setup

### Set parameters

```{r echo = T}
geo_year <- 2021
miles <- .5 #for buffer around transit 

```

### Make some functions

```{r}
coverage_fxn <- function(.geo_main, .geo_census){
  .geo_main %>%
    st_transform(26915) %>% # use a better crs projection for this
    st_intersection(.geo_census %>%
                      st_transform(26915) %>% # use better accuracy projection
                      st_buffer(-30) %>% # give some breathing room for boundaries
                      mutate(total_census_area = st_area(.))) %>%
    mutate(intersect_area = st_area(.)) %>%
    mutate(coverage = as.numeric(intersect_area / total_census_area),
           # if coverage is nearly 100, count it all - usually would go with .95 here, but eval shows only 2 dicey block groups w/ interstate issues
           coverage = case_when(coverage > .8 ~ 1, 
                                TRUE ~ coverage)) %>%
    filter(coverage > .2) # get rid of slivers - usually would go with .05, but eval shows only 2 dicey block group and they have interstate problem
}
```

## Process geometries

Get block groups and assign to CTUs.

```{r census-geography}

# block groups
MN_counties <- c("Anoka", "Carver", "Dakota", "Hennepin", "Ramsey", "Scott", "Washington"
                 # ,"Sherburne", "Isanti", "Chisago", "Goodhue", "Rice", "Le Sueur", "Sibley", "McLeod", "Wright"
)
# WI_counties <- c("St. Croix", "Polk", "Pierce")

MNblock_group <- tigris::block_groups(
  year = geo_year,
  state = "MN",
  county = MN_counties,
  class = "sf"
) %>%
  select(GEOID)

#link with ctu
temp <- tempfile()
temp2 <- tempfile()
download.file(
  "https://resources.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_metc/bdry_metro_counties_and_ctus/shp_bdry_metro_counties_and_ctus.zip",
  destfile = temp
)

unzip(zipfile = temp, exdir = temp2) # list.files(temp2)
ctu_boundary <- sf::read_sf(paste0(temp2, pattern = "/CTUs.shp")) %>%
  select(CTU_ID, CTU_NAME, Shape_Area) %>%
  st_transform(26915) %>%
  st_buffer(-100) #give some breathing room to intersect with block groups


bg_geo <- ctu_boundary %>%
  st_intersection(MNblock_group%>%
                    st_transform(26915)) #bind_rows(MNblock_group, WIblock_group)
# tract_geo <- bind_rows(MNtract, WItract)

# bg_geo %>% st_drop_geometry() %>% count(GEOID) %>% filter(n > 1) %>% arrange(-n)
# bg_geo %>% filter(GEOID == "271390813022") %>%
#   ggplot() + 
#   geom_sf(fill = "blue") +
#   geom_sf(data = filter(ctu_boundary, CTU_NAME %in% c("San Francisco Twp.",
#                                                       "St. Lawrence Twp.",
#                                                       "Sand Creek Twp.",
#                                                       "Belle Plaine Twp.",
#                                                       "Belle Plaine")), color = "red", fill = "transparent")

```

Get a half mile buffer around transportation stops. 

```{r feature-geo}

## transit stops ---------------
temp <- tempfile()
temp2 <- tempfile()
download.file(
  "ftp://ftp.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_metc/trans_transit_stops/shp_trans_transit_stops.zip",
  destfile = temp
)

unzip(zipfile = temp, exdir = temp2) # list.files(temp2)
trans_stops <- sf::read_sf(paste0(temp2, pattern = "/TransitStops.shp")) %>%
  filter(
    active_sta == "Active"
    ,busstop_yn == "Y",
    board_flag == 1
  ) 

trans_buffer <- trans_stops %>%
  # st_transform(4326) %>%
  st_transform(26915) %>% # use better accuracy projection
  st_buffer(dist = 1609.34 * miles) %>% # the 3857 projection uses meters as a distance, so 1.0 mi =
  group_by(active_sta) %>%
  summarise(geometry = st_union(geometry))

trans_stops %>%
  ggplot() +
  geom_sf(data = trans_buffer, fill = "blue", alpha = .4) +
  geom_sf(color = "red", size = .1) +
  geom_sf(data = ctu_boundary, fill = "transparent") +
  theme_void()
# 
# leaflet::leaflet() %>%
#   leaflet::addTiles() %>%
#   leaflet::addPolygons(data = trans_buffer %>% st_transform(4326))%>%
#   leaflet::addCircles(color = "red", data = trans_stops %>% st_transform(4326))
```

## Get census and feature geometry overlap

Overlap: 

```{r agency-coverage}
coverage_bg_transportation <- coverage_fxn(.geo_main = trans_buffer, .geo_census = bg_geo)

```

Population numbers:

```{r get-population-estimates}

# ###########
# # code if using small area estimates; we'll need this for 2021 numbers
# ##########
# temp <- tempfile()
# download.file("https://resources.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_metc/society_small_area_estimates/xlsx_society_small_area_estimates.zip",
#   destfile = temp
# )
# bg_pop_2019 <- readxl::read_xlsx(unzip(temp, "SmallAreaEstimatesBlockGroup.xlsx")) %>%
#   janitor::clean_names() %>%
#   filter(
#     est_year == 2019
#   ) %>%
#   select(pop_est, hh_est, bg10, pophh_est) %>%
#   rename(
#     pop2020 = pop_est,
#     pophh2020 = pophh_est,
#     hh2020 = hh_est
#   )
# fs::file_delete("./SmallAreaEstimatesBlockGroup.xlsx")

###############
# this is the workflow for decennial census numbers
##############
temp <- tempfile()
download.file("https://resources.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_metc/society_census2020population/xlsx_society_census2020population.zip",
              destfile = temp
)
bg_pop <- readxl::read_xlsx(unzip(temp, "Census2020PopulationBlockGroup.xlsx")) %>%
  janitor::clean_names() %>%
  transmute(
    geoid20 = geog_unit,
    pop2020 = poptotal
    # pophh2020 = popinhh,
    # hh2020 = hhtotal,
    # # popover18_census = popover18,
    # whitenh_census = whitenh / pop2020,
    # blacknh_census = blacknh / pop2020,
    # amindnh_census = amindnh / pop2020,
    # asiannh_census = asiannh / pop2020,
    # pacificnh_census = pacificnh / pop2020,
    # othernh_census = othernh / pop2020,
    # multracenh_census = multracenh / pop2020,
    # hisppop_census = hisppop / pop2020
  ) 

fs::file_delete("./Census2020PopulationBlockGroup.xlsx")
```


# Use coverage of geography + census to get demographics in area

create function to link at block group and tract levels

```{r link-census}

link_census_bg_coverage <- function(.df, .group_var) {
  .group_var <- enquos(.group_var)
  
  persons <- .df %>% 
    # coverage_bg_transportation %>% #
    st_drop_geometry() %>%
    # filter(GEOID == "270530265082") %>%
    left_join(bg_pop, by = c("GEOID" = "geoid20")) %>%
    # in some instances, the census has a population > 0 while acs has a population == 0 and also vice versa
    # filter(is.nan(male) | is.nan(female))
    filter(pop2020 > 0)  %>%
    # adjust the person count by the coverage
    mutate(
      adj_poptotal = coverage * pop2020
    ) %>%
    # summarise at the GROUP_VAR level
    # group_by(CTU_NAME) %>%
    group_by(!!!.group_var) %>%
    summarise(
      across(c(adj_poptotal), sum, na.rm = T),
    )  %>%
    mutate(measurement = "persons")
  
  census_agency_summary_1 <- persons %>%
    arrange(!!!.group_var) %>%
    pivot_longer(names_to = "code", values_to = "value", -c(!!!.group_var, measurement)) %>%
    pivot_wider(names_from = measurement, values_from = value) 
  
  
  return(census_agency_summary_1)
}

compile_census <- function(.bg_df, .group_var2){
  step1 <- link_census_bg_coverage(.df = .bg_df, .group_var = !!enquo(.group_var2)) %>%
    arrange(!!enquo(.group_var2))
  return(step1)
}
```  



## Summarize and add in census 2020 population from database

### Establish Database Connection

```{r message=FALSE}
library(odbc)
library(DBI)


dbConnector <- function(server, database,
                        uid = "mc\\eesch", 
                        pwd = getOption("councilR.pwd"),
                        local = TRUE, port = 1433, tds_version = 7.4) {
  if (local) {
    if (Sys.info()[["sysname"]] == "Windows") {
      DBI::dbConnect(odbc::odbc(),
                     Driver = "ODBC Driver 17 for SQL Server",
                     Server = "dbsqlcl11t.test.local,65414",
                     Database = "CD_Emissions",
                     Trusted_Connection = "yes"
      )
    } else {
      # user <- base::readline(prompt = "User: ")
      # pswd <- base::readline(prompt = "Password: ")
      
      DBI::dbConnect(
        odbc::odbc(),
        Driver = "FreeTDS",
        Database = "CD_Emissions",
        Uid = uid,
        Pwd = pwd,
        Server = "dbsqlcl11t.test.local,65414"
      )
    }
  } else {
    DBI::dbConnect(odbc::odbc(),
                   Driver = "FreeTDS",
                   Database = database,
                   Uid = uid,
                   Pwd = pwd,
                   Server = server,
                   Port = port,
                   TDS_Version = tds_version
    )
  }
}

conn <- dbConnector(server, database, counciR.uid, councilR.pwd, is_local)
rm(server, database, uid, pwd, is_local)

```

```{r link-census-transpercent}

census2020pop <-
  # dbGetQuery(conn,"SELECT * FROM metro_sp_mod_2.census_facts 
  #                           WHERE subject_area_desc = 'Population and Households' 
  #                           AND datasource = 'Census 2020'
  #                           AND subject_level_desc = 'Population Total'" ) %>%
  councilR::import_from_emissions("metro_sp_mod_2.census_facts") %>% 
  filter(
    subject_area_desc == "Population and Households",
    datasource == "Census 2020",
    subject_level_desc == 'Population Total') %>% 
  tibble::as_tibble() %>%
  group_by(ctu_id) %>% 
  summarise(persons_census2020 = sum(measure)) #right now this db is split across counties

ctu_pop2020_transit <- compile_census(.bg_df = coverage_bg_transportation,
                                      .group_var2 = CTU_ID) %>%
  select(-code) %>%
  rename(persons_0.5mi_transit = persons,
         ctu_id = CTU_ID) %>%
  full_join(census2020pop) %>%
  
  #if census 2020 pop is < transit buffer pop, adjst accordingly
  mutate(persons_0.5mi_transit = if_else(persons_0.5mi_transit > persons_census2020, 
                                         persons_census2020, persons_0.5mi_transit)) %>%
  
  mutate(percent_pop_0.5mi_transit = persons_0.5mi_transit / persons_census2020) %>%
  full_join(st_drop_geometry(ctu_boundary) %>% select(CTU_ID, CTU_NAME), 
            by = c("ctu_id" = "CTU_ID")) 

save(ctu_pop2020_transit, file = "../data-raw/ctu_pop2020_transit.rda")

```



# Make some figs

```{r}
ctu_pop2020_transit %>%
  ggplot(aes(x = percent_pop_0.5mi_transit)) +
  geom_histogram() +
  theme_minimal()


ctu_pop2020_transit %>%
  full_join((ctu_boundary) %>% select(CTU_ID, CTU_NAME), 
            by = c("ctu_id" = "CTU_ID")) %>%
  ggplot() +
  geom_sf(aes(fill = percent_pop_0.5mi_transit, geometry = geometry)) +
  theme_void()

```




