---
title: "Implications of Green Infrastructure & Land Use"
author: Ellen
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Implications of Green Infrastructure & Land Use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      echo = F)
options(rmarkdown.html_vignette.check_title = FALSE)

library(ghg.sp)
library(knitr)
library(councilR)
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)

`%not_in%` <- Negate(`%in%`)

```


## Implications of Land Use & Green Infrastructure Module

This document tests and synthesizes some outcomes of the land use and green infrastructure modules. A compliment to the introduction vignette; think of the introduction vignette as being the textbook "chapter," and this implications vignette is the "worked examples" at the end of the chapter.  

For simplicity, most implications will be aggregated at the community designation level which is a framework for organizing the communities across the region into broad groups. 


```{r commmunity-designations}
  temp <- tempfile()
  download.file(
    "https://resources.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_metc/society_thrive_msp2040_com_des/gpkg_society_thrive_msp2040_com_des.zip",
    destfile = temp
  )
  comdes <-
    sf::read_sf(unzip(temp, "society_thrive_msp2040_com_des.gpkg")) %>%
    group_by(CTU_NAME) %>%
    mutate(max_area = max(Shape_Area)) %>% #there are some ctus with multiple designations; assign the designation to the most land area
    mutate(flag = case_when(CTU_NAME %in% c("Empire Twp.", "Independence") &  #not the biggest land area
                              COMDESNAME == "Emerging Suburban Edge" ~ 1,
                            CTU_NAME %in% c("Empire Twp.", "Independence") & #not the biggest land area
                              COMDESNAME != "Emerging Suburban Edge" ~ NA_real_,
                            Shape_Area == max_area ~ 1)) %>%
    filter(flag == 1) %>%
    sf::st_drop_geometry() %>%
    ungroup() %>%
    transmute(ctu_name = CTU_NAME,
              comdes = COMDESNAME) %>%
    mutate(ctu_name = if_else(ctu_name == "Credit River", "Credit River Twp.", ctu_name))
  fs::file_delete("society_thrive_msp2040_com_des.gpkg")
  
```

## Land Use and Land Cover Data

### Table Descriptions
The following is a description of the content of all input tables of the land use module, which are
contained in the `land_use_data` dataset.

**$ctu_county**: 
Table shows the relationship between city/townships and county. 

*Test: Are there any city/townships which span counties?*

```{r echo=FALSE}
multiple_cos <- land_use_data$ctu_county %>%
  count(ctu_name) %>%
  filter(n > 1)

if (nrow(multiple_cos) < 1) {
  print("County assignment is okay!")
} else {
  print ("Figure out county assignment for these ctus: ")
  land_use_data$ctu_county %>%
    filter(ctu_name %in% c(multiple_cos$ctu_name)) %>%
    kable()
}
```

**$ctu_forecast**: 
Table shows the demographic forecasts for city/township from Community Profiles.

*Test: Does Lake Elmo population match 2020 census data?*
```{r echo=FALSE}
if (land_use_data$ctu_forecast %>%
  dplyr::filter(ctu_name == "Lake Elmo",
                year == 2020,
                metric == "population") %>% .$value == 11335) {
  print("Yes, looks like the table is pulling the most recent census data. Do check on the forecast, which is (unfortuantely) only accessible via a website download right now; https://metrocouncil.org/Data-and-Maps/Research-and-Data/Thrive-2040-Forecasts.aspx")
} else {
  print("Database probably needs to be updated")
}
```

*Test: Does it look like people, households, and jobs are changing at a reasonable rate?*
```{r}
#| fig.cap = "Lines show ctu-level data and forecasts. Black points and error bars show the mean +- 1 standard error.",
#| fig.width = 7,
#| fig.height = 6

# ee tests
test <- land_use_data$ctu_forecast %>%
  full_join(comdes, by = "ctu_name")

land_use_data$ctu_forecast %>%
  full_join(comdes, by = "ctu_name") %>% 
  filter(!is.na(metric)) %>% #these nas are coming from some ctus missing data (hilltop, rogers, ft snelling), keeping in to check progress
  group_split(metric) %>%
  purrr::map(
    ~ggplot(aes(x = year, 
                y = value), 
            data = .) + 
      geom_line(aes(group = ctu_name),
                col = "grey80") +
      stat_summary(fun.data = mean_se, size = .25) +
      facet_wrap(~comdes, scales = "free_y") +
      theme_council_open() +
      labs(y = "", x = "", col = "", fill = "", shape = "",
           title = .$metric) +
      theme(
        legend.position = "bottom",
        legend.text = element_text(size = 6),
        legend.spacing.y = unit(0, 'cm')))

```

*Test: Which areas are losing people, jobs or households? Does anything fancy with respect to land annexations need to be considered?*
```{r}

#some areas are losing people/jobs
land_use_data$ctu_forecast %>%
  full_join(comdes, by = "ctu_name") %>%
  pivot_wider(names_from = year, values_from = value) %>%
  filter(`2010` > `2040`) %>%
  select(ctu_name, metric, `2010`, `2040`) %>%
  mutate(difference = `2040` - `2010`) %>%
  # split(.$metric) %>%
  kable()
```


**$scenario_parameters**: 

TODO: what checks should be done here?


```{r echo=FALSE}
land_use_data$scenario_parameters %>%
  kable()
```

**$land_cover_percentages_filled**: 


*Test: Does land cover sum to 100%? Does total land area look okay?*

```{r echo=FALSE}
if (nrow(
  land_use_data$land_cover_percentages_filled %>% 
  group_by(ctu_name, description_2) %>%
  summarise(sum = sum(land_cover_percent, na.rm = T)) %>%
  filter(round(sum,3) %not_in% c(0,1))
) > 0) {
  print("Looks like there are probably land cover issues")
} else {
  print("Land cover sums to either 0% or 100%.")
}

land_use_data$land_cover_percentages_filled %>% 
  group_by(ctu_name) %>%
  summarise(total_area = sum(hectares)) %>%
  arrange(total_area) %>%
  ggplot(aes(x = total_area)) +
  geom_histogram() +
  theme_council_open() +
  scale_x_continuous(labels = scales::comma) +
  labs(y = "", x = "total land area (hectares)", 
       title = "Does total land area in ctus look okay?")
```


*Test: What does Lake Elmo's land cover look like?*

TODO: Does this make sense as-is or is some post-processing in order?

1) Is it okay that some descriptions have NO data (like railway)? Should these be zeros instead of NA's?
2) Should agricultural land cover get set to zero in non-agricultural descriptions?
3) Should open_water description be filled with 100% water land cover? 

```{r echo=FALSE}
#| fig.cap = "Lake Elmo land cover.",
#| fig.width = 7,
#| fig.height = 6

land_use_data$land_cover_percentages_filled %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  ggplot(aes(x = description_2, y = land_cover_percent, fill = land_cover_description_2)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Paired") +
  theme_council_open() +
  coord_flip() +
  labs(x = "") +
  facet_wrap(~ctu_name)
```

*Test: What does land cover look like by community designation?*

TODO: Are we good here or need to do post processing?

1) Why does 'seasonal vacation', 'extractive' exist in "urban centers"? 
2) What about "farmstead" in urban? 
3) One agricultural community has 100% of single_family_attached land use being "agricultural" land cover? Should things like this get reclassified?
4) It looks like "water" land cover NEVER makes up any of the "open_water" land use? Should that get changed??
5) What else does this figure show that might be indicative of some needed post processing? 

```{r}
#| fig.cap = "Land cover by community designation.",
#| fig.width = 7,
#| fig.height = 8

land_use_data$land_cover_percentages_filled %>%
    full_join(comdes, by = "ctu_name") %>% #filter(comdes == "Urban Center") %>%
  ggplot(aes(x = description_2, y = land_cover_percent, col = land_cover_description_2, group = ctu_name)) +
  geom_point() + 
  scale_fill_brewer(palette = "Paired") +
  theme_council_open() +
  labs(x = "") +
  # theme(axis.text = element_text(angle = 40)) +
  facet_wrap(~comdes) +
  coord_flip() +
  scale_color_brewer(palette = "Paired")

```


**$land_composition_ctu**: 
Table provides the percentages of land use composition categorized in to three types of development: *exurban development*, *urban expansion*, and *urban infill*. 
*For example: What percentage or what number of hectares of exurban development land area corresponds to parking lot surfaces within agricultural in the city of Lake Elmo for the forecast year 2040?*
```{r echo=FALSE}
land_use_data$land_composition_ctu %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::filter(year == 2040) %>%
  head(., n = 60) %>%
  kable()
```

**$current_conservation_tillage_county**:
Tables shows the percent of agricultural land that is managed with conservation tillage by county. This is a necessary assumption for the conservation tillage strategy. 
*Data comes from the Census of Agriculture (2017).*
*For example: What percentage of agricultural land was using conservation tillage in the Washington County in 201?*
```{r echo=FALSE}
land_use_data$current_conservation_tillage_county %>%
  head(., n = 10) %>%
  kable()
```

**$land_by_development_type_sum_bau**:
Table shows the forecast total number of hectares by city/township for the business as usual scenario. 
*For example: What is the forecast number of hectare for the city of Lake Elmo in 2040?*
```{r echo=FALSE}
land_use_data$land_by_development_type_sum_bau %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 10) %>%
  kable()
```

**$multifamily_mixed_area**: 
Table show the estimate of the number of hectares of multifamily mixed area by city/township for three development types: *exurban development*, *urban expansion*, and *urban infill*.
*For example: What is the forecast hectares of multifamily or mixed use land use for the year 2040 for the city of Lake Elmo?*
```{r echo=FALSE}
land_use_data$multifamily_mixed_area %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 10) %>%
  kable()
```

## Functions

The impact of different land use and greenery strategies is estimated using a series of functions with arguments provided by the user.

The key principle is of this model is that different land cover types have different carbon stock and sequestration value. By altering the composition of a city/township, we could observe higher or lower overall carbon stocks or sequestration value per year. 

### Functions to Calculate Urban Form Strategy. 

The first function is `ghg.sp::calc_by_development_type()`
For help run the following command `?ghg.sp::calc_by_development_type()`
This function is intermediary and does not provide the final result. 

`ghg.sp::calc_by_development_type()` 
The function returns a table with the land composition of different urban form scenarios of land use by development type.

The tool will  perform a series of calculations in 
response to different values of argument `.urban_form_scenario`.
For example, if the selected input is `"bau"`, or the business as usual, 
the tool will select the table it needs. 

The possible scenarios to select on the `.urban_form_scenario` argument include:
* `"bau"`: the business as usual scenario.
* `"post_covid_sprawl"`: post-covid sprawl scenario.
* `"compact_development_beyond_bau"`: compact development beyond the business as usual scenario.
* `"compact_development_with_drs"`: compact development with dynamic ride sharing.

The following code prints the value of function `ghg.sp::calc_by_development_type()` filtered to show only the city of Lake Elmo. 
```{r}
ghg.sp::calc_land_by_development_type(tb = land_use_data,
                                      .urban_form_scenario = "bau") %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  kable()
```

The second function is `ghg.sp::calc_scen_land_use()`
The following code runs the `ghg.sp::calc_scen_land_use()` for the `"bau"` or business as usual scenario. 
```{r}
ghg.sp::calc_scen_land_use(tb =  land_use_data,
                           .urban_form_scenario = "bau") %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 70) %>%
  kable()
```

The third function is `ghg.sp::calc_land_cover_by_land_use()`
The following code runs the `ghg.sp::calc_land_cover_by_land_use()` for the `"bau"` or business as usual scenario.
```{r}
ghg.sp::calc_land_cover_by_land_use(tb =  land_use_data,
                                    .urban_form_scenario = "bau") %>%
  dplyr::filter(year == 2040) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::filter(land_cover_description_2 == 	"forest") %>%
  head(., n = 25)  %>%
  kable()
```

The fourth function is `ghg.sp::calc_land_cover_by_city()`
This function returns the urban form scenario land composition in hectares without adding additional strategies. This is the last function used to estimate the impact of urban form changes alone.  
```{r}
ghg.sp::calc_land_cover_by_city(tb = land_use_data,
                                .urban_form_scenario = "bau") %>%
  dplyr::filter(year == 2040) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 15) %>%
  kable()
```

### Function to Calculate Tree Planting Strategy

Having calculated the hectares of land cover under the selected urban form scenario, we now will add the change to the land composition  that is cause by implement a tree planting strategy. 

Function `ghg.sp::calc_tree_planting_land_cover()`.
```{r}
ghg.sp::calc_tree_planting_land_cover(
  tb = land_use_data,
  detail = FALSE,
  .urban_form_scenario = "bau",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247
)  %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::filter(year == 2040) %>%
  tidyr::pivot_longer(
    names_to = "var",
    values_to = "value",
    cols = -c("ctu_name", "year")
  ) %>%
  head(., n = 15) %>%
  kable()
```

### Function to Calcuate Parking Lot Reduction Strategy

Having calculate the hectares of land cover under the selected urban form and tree planting scenario, we now will add the change to the land composition that is cause by implementing a reduced parking lot strategy. 

Function `ghg.sp::calc_parking_lot_land_cover()`.
```{r}
ghg.sp::calc_parking_lot_land_cover(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .parking_lot_reduction_percentage = 0.8,
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  detail = FALSE
)  %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::filter(year == 2040) %>%
  tidyr::pivot_longer(
    names_to = "var",
    values_to = "value",
    cols = -c("ctu_name", "year")
  ) %>%
  head(., n = 15) %>%
  kable()
```

### Function to Calculate Conservation Tillage Strategy

```{r}
ghg.sp::calc_conservation_tillage(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .conservation_tillage_intervention = "current_conservation_tillage",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .w2w_diesel_emission_factor_kg_co2e_per_gal = 12.50,
  .avoided_emissions_tractor_use_mg_co2e_per_hectare = 0.0102,
  .agricultural_land_carbon_stock_mg_c_per_hectare = 3,
  .maximum_soc_accumation_under_reduced_or_no_till_ag = 1.54,
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  .parking_lot_reduction_percentage = 0.8,
  detail = FALSE
)  %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  tidyr::pivot_longer(
    names_to = "var",
    values_to = "value",
    cols = -c("ctu_name", "co_name")
  ) %>%
  head(., n = 15) %>%
  kable()
```

---

# STOP REVIEW HERE 


----


```{r}
ghg.sp::calc_carbon_stock_per_ctu(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  .parking_lot_reduction_percentage = 0.8,
  .conservation_tillage_intervention = "current_conservation_tillage",
  .w2w_diesel_emission_factor_kg_co2e_per_gal = 12.50,
  .agricultural_land_carbon_stock_mg_c_per_hectare = 3,
  .maximum_soc_accumation_under_reduced_or_no_till_ag = 1.54,
  .avoided_emissions_tractor_use_mg_co2e_per_hectare = 0.0102,
  detail = FALSE
) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 10) %>%
  kable()
```

