---
title: "Implications of Green Infrastructure & Land Use"
author: Ellen, Liz
output: rmarkdown::html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      echo = F)
# options(rmarkdown.html_vignette.check_title = FALSE)

library(ghg.sp)
library(knitr)
library(councilR)
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)

`%not_in%` <- Negate(`%in%`)

```


## Implications of Land Use & Green Infrastructure Module

This document tests and synthesizes some outcomes of the land use and green infrastructure modules. A compliment to the introduction vignette; think of the introduction vignette as being the textbook "chapter," and this implications vignette is the "worked examples" at the end of the chapter.  

For simplicity, most implications will be aggregated at the community designation level which is a framework for organizing the communities across the region into broad groups. 


```{r commmunity-designations}
temp <- tempfile()
download.file(
  "https://resources.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_metc/society_thrive_msp2040_com_des/gpkg_society_thrive_msp2040_com_des.zip",
  destfile = temp
)
comdes <-
  sf::read_sf(unzip(temp, "society_thrive_msp2040_com_des.gpkg")) %>%
  group_by(CTU_NAME) %>%
  mutate(max_area = max(Shape_Area)) %>% #there are some ctus with multiple designations; assign the designation to the most land area
  mutate(flag = case_when(CTU_NAME %in% c("Empire Twp.", "Independence") &  #not the biggest land area
                            COMDESNAME == "Emerging Suburban Edge" ~ 1,
                          CTU_NAME %in% c("Empire Twp.", "Independence") & #not the biggest land area
                            COMDESNAME != "Emerging Suburban Edge" ~ NA_real_,
                          Shape_Area == max_area ~ 1)) %>%
  filter(flag == 1) %>%
  sf::st_drop_geometry() %>%
  ungroup() %>%
  transmute(ctu_name = CTU_NAME,
            comdes = COMDESNAME) %>%
  mutate(ctu_name = if_else(ctu_name == "Credit River", "Credit River Twp.", ctu_name))
fs::file_delete("society_thrive_msp2040_com_des.gpkg")

```

## Land Use and Land Cover Data

### `land_use_data` Table Descriptions
The following is a description of the content of all input tables of the land use module, which are
contained in the `land_use_data` dataset.

**$ctu_county**: 
Table shows the relationship between city/townships and county. 

*Q: Are there any city/townships which span multiple counties?* 

1. GitHub [issue #17](https://github.com/Metropolitan-Council/ghg.sp/issues/17) gives suggestions as to how these CTUs should be handled. One option here would be to consider ADDING a column (percent_coverage?) so that it is clear how county assignment will happen (esp. when Matt suggests using a weighted mean when county-level averages are needed). 

```{r echo=FALSE}
multiple_cos <- land_use_data$ctu_county %>%
  count(ctu_name) %>%
  filter(n > 1)

if (nrow(multiple_cos) < 1) {
  print("County assignment is okay!")
} else {
  print ("Figure out county assignment for these ctus: ")
  land_use_data$ctu_county %>%
    filter(ctu_name %in% c(multiple_cos$ctu_name)) %>%
    kable()
}

```

2. Additionally! Please *REMOVE* county columns from all other databases associated with this project! For instance, the `db_tables$t_housing_stock_ctu` table has both ctu_name and co_name columns. If ONLY the `ctu_county` table contains the county information, that means that you will only have to consider this issue (county assignments) once instead of multiple times (yay!). Then this "global" information (ctu and county assignment) can just be joined into other tables and you will be confident in the accuracy of the assignment. 

```{r}

db_tables$t_housing_stock_ctu %>%
  filter(year == 2018,
         ctu_name == "St. Anthony") %>%
  select(ctu_name, co_name) %>%
  head()

```

**$ctu_forecast**: 
Table shows the demographic forecasts for city/township from Community Profiles.

*Q: Does Lake Elmo population match 2020 census data?*

```{r echo=FALSE}
if (land_use_data$ctu_forecast %>%
    dplyr::filter(ctu_name == "Lake Elmo",
                  year == 2020,
                  metric == "population") %>% .$value == 11335) {
  print("Yes, looks like the table is pulling the most recent census data. Do check on the forecast, which is (unfortuantely) only accessible via a website download right now; https://metrocouncil.org/Data-and-Maps/Research-and-Data/Thrive-2040-Forecasts.aspx")
} else {
  print("Database doesn't match 2020 census data on community profiles; database probably needs to be updated")
}
```

*Q: Does it look like people, households, and jobs are changing at a reasonable rate?*

ee thoughts --> yeah, this mostly looks okay. Check out a ctus that are losing people/jobs in table below.

```{r, message = F}
#| fig.cap = "Lines show ctu-level data and forecasts. Black points and error bars show the mean +- 1 standard error.",
#| fig.width = 7,
#| fig.height = 6

# ee tests
test <- land_use_data$ctu_forecast %>%
  full_join(comdes, by = "ctu_name")

land_use_data$ctu_forecast %>%
  full_join(comdes, by = "ctu_name") %>% 
  filter(!is.na(metric)) %>% #these nas are coming from some ctus missing data (hilltop, rogers, ft snelling), keeping in to check progress
  group_split(metric) %>%
  purrr::map(
    ~ggplot(aes(x = year, 
                y = value), 
            data = .) + 
      geom_line(aes(group = ctu_name),
                col = "grey80") +
      stat_summary(fun.data = mean_se, size = .25) +
      facet_wrap(~comdes, scales = "free_y") +
      theme_council_open() +
      labs(y = "", x = "", col = "", fill = "", shape = "",
           title = .$metric) +
      theme(
        legend.position = "bottom",
        legend.text = element_text(size = 6),
        legend.spacing.y = unit(0, 'cm')))

```

*Q: Which areas are losing people, jobs or households?*

1) Does this make sense? Or does something fancy with respect to land annexations need to be considered?

```{r}

#some areas are losing people/jobs
land_use_data$ctu_forecast %>%
  full_join(comdes, by = "ctu_name") %>%
  pivot_wider(names_from = year, values_from = value) %>%
  filter(`2010` > `2040`) %>%
  select(ctu_name, metric, `2010`, `2040`) %>%
  mutate(difference = `2040` - `2010`) %>%
  # split(.$metric) %>%
  kable()
```


**$scenario_parameters**: 

*Q: why do all scenario_description_1 have exurban_development ==0 ?*


```{r echo=FALSE}
land_use_data$scenario_parameters %>%
  kable()
```

**$land_cover_percentages_filled**: 

What is the most root data source? Milnar and Ramaswani, 2020?


*Q: Does land cover sum to 100%?* 

```{r echo=FALSE}
if (nrow(
  land_use_data$land_cover_percentages_filled %>% 
  group_by(ctu_name, description_2) %>%
  summarise(sum = sum(land_cover_percent, na.rm = T), .groups = "keep") %>%
  filter(round(sum,3) %not_in% c(0,1))
) > 0) {
  print("Looks like there are probably land cover issues")
} else {
  print("Land cover sums to either 0% or 100%.")
}
```

*Q: Does total land area look okay?*

ee thoughts --> yup, looks okay

```{r, message = F}
#| fig.width = 6,
#| fig.height = 4

land_use_data$land_cover_percentages_filled %>% 
  group_by(ctu_name) %>%
  summarise(total_area = sum(hectares), .groups = "keep") %>%
  arrange(total_area) %>%
  ggplot(aes(x = total_area)) +
  geom_histogram() +
  theme_council_open() +
  scale_x_continuous(labels = scales::comma) +
  labs(y = "", x = "total land area (hectares)", 
       title = "Does total land area in ctus look okay?")
```


*Q: What does Lake Elmo's land cover look like? Do these values make sense as-is or is some post-processing in order?*

1) Is it okay that some descriptions have NO data (like railway)? 
2) It seems like zero may be more appropriate than NA's - update those?
3) Should agricultural land cover get set to zero in non-agricultural descriptions (i.e., why is agricultural land cover in "retail and other commercial" land use)?
4) Should open_water description be filled with 100% water land cover? I do think that the wetland and woody_wetland land covers also seem appropriate, but am unsure why nearly 50% of "open water" is forest land cover, for instance. 
5) What is the difference between railway and major_railway? There is no highway equivalent (yes there is major_highway, but no plain "highway") - can the railroad stuff get condensed/collapsed/simplified like the highway stuff?

```{r echo=FALSE, message = F}
#| fig.cap = "Lake Elmo land cover.",
#| fig.width = 7,
#| fig.height = 8

land_use_data$land_cover_percentages_filled %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  ggplot(aes(x = description_2, y = land_cover_percent, fill = land_cover_description_2)) +
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Paired") +
  theme_council_open() +
  coord_flip() +
  labs(x = "",
       y = "Land Use") +
  facet_wrap(~ctu_name) +
  theme(legend.position = "bottom") +
  guides(
    col = guide_legend(nrow = 4, byrow = TRUE),
    fill = guide_legend(nrow = 4, byrow = TRUE),
    shape = guide_legend(nrow = 4, byrow = TRUE)
  )
```

*Q: What does land cover look like by community designation? Are we good here or need to do post processing?* There is a lot going on in this figure, but I'm using it for high-level red-flag checks. It might be worth unpacking this a bit more once these initial questions are answered and/or data updates made. 

1) Why does 'seasonal vacation', 'extractive' exist in "urban centers"? Can this be set to zero and/or re-calculated? 
2) Why does "farmstead" occur in urban communities? Again, can this be set to zero or recalculated? (Or simply just confirmed and/or checked via a map??)
3) One agricultural community has 100% of single_family_attached land use being "agricultural" land cover? Should things like this get reclassified?
4) It looks like "water" land cover *NEVER* makes up any of the "open_water" land use? Should that get changed??
5) What else does this figure show that might be indicative of some needed post processing? 
6) Why do agricultural communities have railway AND major_railway land uses? Can these two be combined?

```{r, message = F}
#| fig.cap = "Land cover by community designation.",
#| fig.width = 8,
#| fig.height = 12

land_use_data$land_cover_percentages_filled %>%
  full_join(comdes, by = "ctu_name") %>% #filter(comdes == "Urban Center") %>%
  ggplot(aes(x = description_2, 
             y = land_cover_percent, 
             col = land_cover_description_2, 
             group = ctu_name)) +
  geom_point(alpha = 0.5) + 
  scale_fill_brewer(palette = "Paired") +
  theme_council_open() +
  labs(x = "") +
  # theme(axis.text = element_text(angle = 40)) +
  facet_wrap(~comdes) +
  coord_flip() +
  scale_color_brewer(palette = "Paired") +
  theme(legend.position = "bottom") +
  guides(
    col = guide_legend(nrow = 4, byrow = TRUE),
    fill = guide_legend(nrow = 4, byrow = TRUE),
    shape = guide_legend(nrow = 4, byrow = TRUE)
  )

```


**$land_composition_ctu**: 
Table provides the percentages of land use composition categorized in to three types of development: *exurban development*, *urban expansion*, and *urban infill*. 

What is the root data source?

*Q: How does Lake Elmo's land use/cover change based on the type of development?*

1) If we look only at the land use/covers which change from 2016 to 2040, the "exurban development" type doesn't affect the percent of ANY land use/cover in Lake Elmo. Is there a multiplier / parameter missing somewhere? 
2) Does it make sense that "urban_infill" only impacts "railway" and "retail_and_other_commercial" while the "urban_expansion" doesn't affect those two land cover/uses?

```{r}
land_use_data$land_composition_ctu %>%
  dplyr::filter(ctu_name %in% c("Lake Elmo",
                                "St. Paul")) %>%
  select(-total_hectares, -hectares) %>%
  pivot_wider(names_from = year, values_from = percent) %>%
  mutate(change = `2040` - `2016`) %>%
  filter(change != 0) %>%
  kable()
```

*Q: Do the different development types change land use/cover in logical ways?* The first figure shows all land uses, the second zooms in on land cover/uses which make up 5% or more of the land base. 

1) Urban Expansion makes Lake Elmo go from less than 5% single family detached homes in 2016 to over 80% single family detached homes in 2040. It seems like either the 2016 number could be an underestimate, or the 2040 number is an over estimate - maybe check on that? 
2) Does it make sense that the urban expansion scenario would "consume" all "park_recreational_or_preserve"? I can see consumption of agricultural and undeveloped land, but based on what I know from Paul, the parks stay parks. 
3) Why is the 2016 land cover/use different across the different scenarios? I was under the impression that the starting point was the same for all scenarios (and that the land use/cover amounts came from observed, on-the-ground data in 2016), and that it was only *future* years where the scenarios really came into play. What am I missing, or does the 2016 land cover/use need to be updated? 

```{r echo=FALSE}
#| fig.cap = "Development changes Lake Elmo land use/cover.",
#| fig.width = 8,
#| fig.height = 6

land_use_data$land_composition_ctu %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  ggplot(aes(x = as.character(year), y = percent, fill = description_2)) +
  geom_bar(stat = "identity") +
  facet_wrap(~development_name) +
  theme_council_open() +
  labs(y = "", x = "",
       title = "Lake Elmo",
       subtitle = "All land cover/uses") 

land_use_data$land_composition_ctu %>%
  dplyr::filter(ctu_name == "Lake Elmo",
                percent >= .05) %>%
  ggplot(aes(x = as.character(year), y = percent, fill = description_2)) +
  geom_bar(stat = "identity") +
  facet_wrap(~development_name) +
  theme_council_open() +
  labs(y = "", x = "",
       title = "Lake Elmo",
       subtitle = "For land cover/uses which make up 5% or more") +
  scale_fill_brewer(palette = "Paired")

```

*Q: What about development scenarios on other communities?*

1) Brooklyn Center is showing up twice in this `land_use_data$land_composition_ctu` database. Please update so there's only one value for each year/development_name/description_2 for Brooklyn Center so that we know which value to refer to! As an FYI, Brooklyn Center *is* "falling out" of the analysis at later steps because of this issue - once there is only one entry for Brooklyn Center in this `land_consumption_ctu` table, this should be resolved. 

```{r echo=FALSE}

land_use_data$land_composition_ctu %>%
  select(-total_hectares, -percent) %>%
  dplyr::group_by(ctu_name, development_name, description_2, year) %>%
  dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
  dplyr::filter(n > 1L) %>%
  head() %>%
  kable()
```


So EXCLUDING Brooklyn Center and looking at all other communities (and only the land uses/covers which changed by 5% or more):

1) Urban infill only changes retail and railway in all communities? 
2) Should urban expansion change (decrease) airport land use/cover in urban center areas (bu over 50%)? It seems like this values should be pretty constant.
3) Under what conditions does urban expansion decrease single family attached, detached, and multifamily residential? In some instances these values are below zero indicating LOSS of this land cover/use.

```{r}
#| fig.cap = "Development changes land use/cover - entire region.",
#| fig.width = 8,
#| fig.height = 6

land_use_data$land_composition_ctu %>%
  select(-total_hectares, -hectares) %>%
  filter(ctu_name != "Brooklyn Center") %>% #filter(description_2 == "farmstead") %>%
  pivot_wider(names_from = year, values_from = percent) %>%
  mutate(change = `2040` - `2016`) %>%
  filter(abs(change) >= .05) %>%
  left_join(comdes, by = "ctu_name") %>%
  ggplot(aes(x = description_2, y = change, col = comdes, group = ctu_name)) +
  geom_point(alpha = 0.5)+
  facet_wrap(~development_name, scales = "free") +
  coord_flip() +
  theme_council_open() +
  scale_color_brewer(palette = "Paired") +
  theme(legend.position = "bottom") +
  geom_hline(yintercept = 0) +
  guides(
    col = guide_legend(nrow = 4, byrow = TRUE),
    fill = guide_legend(nrow = 4, byrow = TRUE),
    shape = guide_legend(nrow = 4, byrow = TRUE)
  ) +
  labs(title = "Change in land use over time across scenarios - all ctus.", 
       y = "Change in percent of land use/cover from 2016 to 2040\n(positive values indicate EXPANSION of land use type by 2040,\nnegative values indicate LOSS of land use type by 2040", x = "")


```

4) Why do communities gain farmstead with urban expansion? It seems like agricultural should be lost, especially if the land use/cover of farmstead was zero in 2016. The below table is showing the change in number of hectares of farmstead land cover/use.

```{r}
land_use_data$land_composition_ctu %>%
  select(-total_hectares, -percent) %>%
  filter(ctu_name != "Brooklyn Center") %>% filter(description_2 == "farmstead") %>%
  pivot_wider(names_from = year, values_from = hectares) %>%
  mutate(change = `2040` - `2016`) %>%
  filter(abs(change) >= .05, `2016` == 0) %>%
  left_join(comdes, by = "ctu_name") %>%
  rename(`2016 hectares` = `2016`,
         `2040 hectares` = `2040`,
         change_hectares = change) %>%
  kable()

```

**$current_conservation_tillage_county**:
Tables shows the percent of agricultural land that is managed with conservation tillage by county. This is a necessary assumption for the conservation tillage strategy. 

*Q: Just with my background, these numbers seemed pretty low to me. For instance, check out [this high level report from the USDA on conservation tillage numbers](https://www.ers.usda.gov/webdocs/publications/90201/eib197_summary.pdf?v=8320.9); you'll see that across the US, over 60% of corn is in conservation tillage and over 70% of beans are in conservation tillage (and these are the two main crops grown in MN). So I went directly to the USDA's data (via API), and got some pretty different numbers.*

1) Please update the GHG database values so that they match USDA Census values for conservation tillage (see figure below)! If the definition of conservation tillage varies between the USDA and this project, that should also be explained (and am happy to help confirm that the correct numbers are being used - the api thing can be a little tricky, feel free to ask for help!). 


To get raw USDA data
- Get an API key from USDA: https://quickstats.nass.usda.gov/api
- Enter in the console: `usethis::edit_r_environ()`
- When the `.Renviron` file comes up in the editor, type: `USDA_KEY="KEY GOES HERE, INSIDE QUOTES"`
- Save and close the `.Renviron` file.
- Restart R.


```{r echo=FALSE}
#| fig.cap = "Comparison of conservation tillage numbers from USDA versus from database.",
#| fig.width = 5,
#| fig.height = 3


get_usda <- FALSE

if(get_usda == TRUE) {
  mn_conservation_tillage_fxn <- function(.county){
    raw <- httr::GET(paste0(
      "http://quickstats.nass.usda.gov/api/api_GET/?key=",
      Sys.getenv("USDA_KEY"),
      "&source_desc=CENSUS",
      "&commodity_desc=PRACTICES",
      "&class_desc=LAND%20USE,%20CROPLAND",
      "&unit_desc=ACRES",
      "&agg_level_desc=COUNTY",#STATE",
      "&county_name=", .county, #WASHINGTON",
      "&domain_desc=TOTAL",
      "&year=2017",
      "&state_alpha=MN&format=JSON"))
    
    parsed <- jsonlite::fromJSON(httr::content(x = raw, as="text"))$data
    
    return(parsed)
  }
  
  mn_conservation_tillage_raw <-
    mn_conservation_tillage_fxn("ANOKA") %>%
    bind_rows(mn_conservation_tillage_fxn("CARVER")) %>%
    bind_rows(mn_conservation_tillage_fxn("DAKOTA")) %>%
    bind_rows(mn_conservation_tillage_fxn("HENNEPIN")) %>%
    bind_rows(mn_conservation_tillage_fxn("RAMSEY")) %>%
    bind_rows(mn_conservation_tillage_fxn("SCOTT")) %>%
    bind_rows(mn_conservation_tillage_fxn("WASHINGTON"))
  
  usda_data <- mn_conservation_tillage_raw %>%
    mutate(Value = as.numeric(gsub(",", "", Value))) %>%
    filter(prodn_practice_desc != "COVER CROP PLANTED, (EXCL CRP)") %>%
    select(county_name, prodn_practice_desc, Value) %>%
    group_by(county_name) %>%
    mutate(total_acres = sum(Value),
           percent = Value / total_acres) %>%
    mutate(tillage = if_else(prodn_practice_desc %in% 
                               c("CONSERVATION TILLAGE, (EXCL NO-TILL)",
                                 "CONSERVATION TILLAGE, NO-TILL"),
                             "conservation tillage", "conventional")) %>%
    group_by(county_name, tillage) %>%
    summarise(current_conservation_tillage_percent = sum(percent), .groups = "keep") %>%
    filter(tillage == "conservation tillage") %>%
    mutate(source = "USDA API",
           co_name = stringr::str_to_title(county_name)) 
  save(usda_data, file = "./usda_data.rda")
} else {
  load("./usda_data.rda")
}

usda_data %>%
  bind_rows(land_use_data$current_conservation_tillage_county %>%
              mutate(source = "GHG database")) %>%
  ggplot(aes(y = fct_rev(co_name), 
             x = current_conservation_tillage_percent,
             col = source)) +
  geom_point() +
  theme_council_open() +
  labs(y = "", x = "Conservation tillage percent")


```

**$land_by_development_type_sum_bau**:
Table shows the forecast total number of hectares by city/township for the business as usual scenario. 

*Q: Why does the BAU scenario have land areas that are in some cases a lot different than the sum of land area in the `land_use_data` database?*

1) Yay Hilltop land areas look to be the same!
2) Why would are area like Landfall have 201 total hectares in 2016 but only 16 total hectares in the 2040 business as usual scenario?
3) Would recommend trying to reduce redundancy in databases - if you can get at the numbers in `land_by_development_type_sum_bau` by summing disaggregated numbers in `land_cover_percentages_filled` you'll eventually have fewer places where inevitable typos or errors happen.

```{r echo=FALSE, message = F}

land_use_data$land_cover_percentages_filled %>% 
  group_by(ctu_name) %>%
  summarise(land_cover_percentages_filled = sum(hectares), .groups = "keep") %>%
  arrange(land_cover_percentages_filled) %>%
  full_join(land_use_data$land_by_development_type_sum_bau) %>%
  head() %>%
  kable()
```


<!-- # ELLEN ENDING CHECKS HERE FOR NOW -->


**$multifamily_mixed_area**: 
Table show the estimate of the number of hectares of multifamily mixed area by city/township for three development types: *exurban development*, *urban expansion*, and *urban infill*.

*Q: why do the values for multifamily differ between the `multifamily_mixed_area` and `land_composition_ctu` tables (where description_2 == "multifamily" and year == "2040" for the land_composition_ctu table)? It would be good to have a describe how/why these two things are different (I'm not sure I know), or maybe an update to the database is needed.*

```{r echo=FALSE, message = F}
land_use_data$multifamily_mixed_area  %>%
  rename(multifamily_mixed_area__hectares = hectares) %>%
  # left_join(comdes, by = "ctu_name")  %>%
  full_join(
    land_use_data$land_composition_ctu %>%
      filter(str_detect(description_2, "multi")) %>%
      filter(year == 2040) %>%
      rename(land_composition_ctu__multifamily2040__hectares = hectares)
  ) %>%
  select(-description_2, -total_hectares, -percent, -year) %>%
  filter(ctu_name == "Lake Elmo") %>%
  kable()
```

*Q: Does it make sense that exurban development scenario has nearly 6 times more multifamily mixed area than does urban infill? Does it make sense that urban expansion basically has no change in the total hectares in multifamily mixed area? I'm not sure, but it could be good to describe what these different scenarios mean so that people reading the vignette understand a bit more and can contextualize their understanding.*


```{r, message = F}
#| fig.cap = "Total hectares, multifamily mixed area.",
#| fig.width = 5,
#| fig.height = 4

land_use_data$multifamily_mixed_area  %>%
  rename(multifamily_mixed_area__hectares = hectares) %>%
  left_join(comdes, by = "ctu_name") %>%
  group_by(comdes, development_name) %>%
  summarise(hectares = sum(multifamily_mixed_area__hectares), .groups = "keep") %>%
  ggplot(aes(x = development_name, y = hectares, fill = comdes)) +
  geom_bar(stat = "identity") +
  # facet_wrap(~comdes) +
  theme_council_open() +
  scale_fill_brewer(palette = "Paired") +
  scale_y_continuous(labels = scales::comma)+
  labs(y = "", x = "", 
       title = "Total hectares, multifamily_mixed_area")
# coord_flip()

```

## Functions

The impact of different land use and greenery strategies is estimated using a series of functions with arguments provided by the user.

The key principle is of this model is that different land cover types have different carbon stock and sequestration value. By altering the composition of a city/township, we could observe higher or lower overall carbon stocks or sequestration value per year. 

### Functions to Calculate Urban Form Strategy. 

`ghg.sp::calc_by_development_type()` 

*TEST: this function is breaking for scenario `compact_dev_with_drs`*

1) Please ensure this function runs for all scenarios!

```{r, echo = T, include=F, eval =F}

# run for every scenario description
c(land_use_data$scenario_parameters$scenario_description_2) %>% 
  purrr::set_names() %>% 
  purrr::map(~ghg.sp::calc_land_by_development_type(tb = land_use_data,
                                                    .urban_form_scenario = .x) #%>%
             # mutate(scenario_description_2 = paste(.x))
  ) %>%
  bind_rows(.id = "scenario_description_2") #its a bit confusing, but .id must equal the column name in the list that distinguishes the different scenarios; so in this case it is from scenario_parameters$scenario_description_2; and called scenario_description_2

```

But if we do some checks without the breaking scenario...

*Q: Does it make sense that the difference between exurban_development, urban_expansion, and urban_infill are ALWAYS greater than the difference between "bau", "post_covid_sprawl", and "compact_dev_with_drs"?*

1) I'm feeling a little overwhelmed with the number of times I'm seeing "scenarios" - what's the difference between the c(exurban_development, urban_expansion, and urban_infill) scenarios and the c("bau", "post_covid_sprawl", and "compact_dev_with_drs") scenarios? Is there a synonym for scenarios that can be used for one of these options to add some clarity? 
2) Does it make sense that the "scenario_mixed_use_compact_zoning_park" and "scenario_mixed_use_mf_new" only appear in the "exurban_development" scenario? 
3) Does it make sense that the "bau_mixed_use_compact_zoning" scenario only appears in the "exurban_development" scenario?

```{r}
#| fig.width = 8,
#| fig.height = 4
output <- 
  c(land_use_data$scenario_parameters$scenario_description_2[c(1:2, 4)]) %>% 
  purrr::set_names() %>% 
  purrr::map(~ghg.sp::calc_land_by_development_type(tb = land_use_data,
                                                    .urban_form_scenario = .x) #%>%
             # mutate(scenario_description_2 = paste(.x))
  ) %>%
  bind_rows(.id = "scenario_description_2")

output %>%
  filter(ctu_name == "Lake Elmo") %>%
  ggplot(aes(x = development_name, y = hectares, fill = scenario)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~scenario_description_2) +
  theme_council_open() +
  scale_fill_brewer(palette = "Paired") +
  theme(axis.text.x = element_text(angle = 24)) +
  labs(title = "Lake Elmo")

```

The second function is `ghg.sp::calc_scen_land_use()`

*Test: again, this function breaks for the "compact_dev_with_drs" scenario.*

```{r}
# output <- 
#   c(land_use_data$scenario_parameters$scenario_description_2) %>% 
#     purrr::set_names() %>% 
#     purrr::map(~ghg.sp::calc_scen_land_use(tb = land_use_data,
#                                               .urban_form_scenario = .x) 
#                  ) %>%
#   bind_rows(.id = "scenario_description_2")
```

But moving past that...

1) Why is the output of this function (calc_scen_land_use) called "scenario_hectares" while the output of the previous function (calc_land_by_development_type) was simply "hectares"? Could they both become "hectares" for consistency's sake? 
2) Why does "compact_dev_with_drs" have zero hectares for the exurban_development and urban_expansion? 
3) Why are the land cover hectares different across the different scenarios here? Is the calc_scen_land_use function assuming that Lake Elmo's total amount of land will change based on the scenarios (i.e. is the function multiplying a scaling factor through the total hectares rather than through the PERCENT of hectares?)

```{r}
#| fig.width = 8,
#| fig.height = 6
#| 
output <- 
  c(land_use_data$scenario_parameters$scenario_description_2[c(1:2, 4)]) %>% 
  purrr::set_names() %>% 
  purrr::map(~ghg.sp::calc_scen_land_use(tb = land_use_data,
                                         .urban_form_scenario = .x) 
  ) %>%
  bind_rows(.id = "scenario_description_2")


output %>%
  filter(ctu_name == "Lake Elmo") %>%
  ggplot(aes(x = development_name, y = scenario_hectares, fill = description_2)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~scenario_description_2) +
  theme_council_open() +
  # scale_fill_brewer(palette = "Paired") +
  theme(axis.text.x = element_text(angle = 24)) +
  labs(title = "Lake Elmo")

```

The third function is `ghg.sp::calc_land_cover_by_land_use()`

*Test: yeeeep, breaking for the compact_dev_with_drs scenario....*


```{r}
# output <- 
#   c(land_use_data$scenario_parameters$scenario_description_2) %>% 
#     purrr::set_names() %>% 
#     purrr::map(~ghg.sp::calc_land_cover_by_land_use(tb = land_use_data,
#                                               .urban_form_scenario = .x) 
#                  ) %>%
#   bind_rows(.id = "scenario_description_2")
```

Moving on...

1) Why is the output of this function (calc_scen_land_use) called "land_cover_land_use_hectares"? **Recommendation: introduce a new column to all these exports! Consider "output_type" and "hectares" ...that way you can put "land cover land use" or whatever else might distinguish these, and know that the numeric value will be hectares in all cases (since the column will be simply labeled "hectares"!)**
2) Why does "compact_dev_with_drs" have way less land_cover_land_use_hectares in year 2040 as compared to year 2016? Is Lake Elmo loosing land? 
3) Why does "post_covid_sprawl" give Lake Elmo way more land in 2040 as compared to year 2016? Again it seems like a multiplier might be getting applied in the wrong place? Why would the total land base (total land_cover_land_use_hectares) change between years within one city? It does make sense that the ALLOCATION of land use/cover would change, but the TOTAL hectares seem like it should remain constant.

```{r}
#| fig.width = 8,
#| fig.height = 8
#| 
output <- 
  c(land_use_data$scenario_parameters$scenario_description_2[c(1:2, 4)]) %>% 
  purrr::set_names() %>% 
  purrr::map(~ghg.sp::calc_land_cover_by_land_use(tb = land_use_data,
                                                  .urban_form_scenario = .x) 
  ) %>%
  bind_rows(.id = "scenario_description_2")


output %>%
  filter(ctu_name == "Lake Elmo") %>%
  ggplot(aes(x = description_2, y = land_cover_land_use_hectares, fill = land_cover_description_2)) + 
  geom_bar(stat = "identity") +
  facet_wrap(year~scenario_description_2) +
  theme_council_open() +
  scale_fill_brewer(palette = "Paired") +
  theme(axis.text.x = element_text(angle = 24)) +
  labs(title = "Lake Elmo") +
  coord_flip() +
  labs(x = "")

```




The fourth function is `ghg.sp::calc_land_cover_by_city()`

*Test: still, breaking for the compact_dev_with_drs scenario....*


```{r}
# output <- 
#   c(land_use_data$scenario_parameters$scenario_description_2) %>% 
#     purrr::set_names() %>% 
#     purrr::map(~ghg.sp::calc_land_cover_by_city(tb = land_use_data,
#                                               .urban_form_scenario = .x) 
#                  ) %>%
#   bind_rows(.id = "scenario_description_2")
```

Moving on...

1) Again, consider the "land_cover_hectares" variable name...how/why is this different than the "land_cover_land_use_hectares" variable name in the previous function (calc_land_cover_by_city)?
2) Why is Lake Elmo getting more land under the post_covid_sprawl in year 2040? 
3) Why is Lake Elmo getting less land under the compact_dev_with_drs in year 2040?

```{r}
#| fig.width = 8,
#| fig.height = 5
#| 
output <- 
  c(land_use_data$scenario_parameters$scenario_description_2[c(1:2, 4)]) %>% 
  purrr::set_names() %>% 
  purrr::map(~ghg.sp::calc_land_cover_by_city(tb = land_use_data,
                                              .urban_form_scenario = .x) 
  ) %>%
  bind_rows(.id = "scenario_description_2")


output %>%
  filter(ctu_name == "Lake Elmo") %>%
  ggplot(aes(x = as.character(year), y = land_cover_hectares, fill = land_cover_description_2)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~scenario_description_2) +
  theme_council_open() +
  scale_fill_brewer(palette = "Paired") +
  # theme(axis.text.x = element_text(angle = 24)) +
  labs(title = "Lake Elmo") +
  # coord_flip() +
  labs(x = "")

```




### Function to Calculate Tree Planting Strategy

Function `ghg.sp::calc_tree_planting_land_cover()`.

*Test: meh, still the compact_dev_with_drs issue.*

1) It looks like the tree planting on all pervious surfaces increases tree cover - great! But the "double" and "match la million trees goal" don't seem different? Should they be? If they aren't different, can the "double" intervention get changed to "triple" or something else that WILL produce different results? I guess the two strategies just seem redundant right now (so simply deleting one strategy would also work!).
2) In year 2040, I don't think that the total hectares in Lake Elmo should be changing. This seems NOT a multiplier issue, but rather an issue that if MORE land cover is taken up by trees...another land cover must DECREASE (like a scale that needs to be balanced). What land uses will you decrease if tree planting intensifies???

```{r, message = F}
#| fig.width = 8,
#| fig.height = 5
#| 
c("tree_planting_on_all_pervious",
  "double",
  "match_la_million_trees_goal")

la_goal <- c(land_use_data$scenario_parameters$scenario_description_2[c(1:2,4)]) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_tree_planting_land_cover(tb = land_use_data,
                                                    detail = FALSE,
                                                    .urban_form_scenario = .x,
                                                    .tree_planting_intervention = "match_la_million_trees_goal",
                                                    .tree_planting_per_capita = 0.26,
                                                    .tree_planting_per_hectare = 247) 
  ) %>%
  bind_rows(.id = "scenario_description_2") %>%
  mutate(intervention = "match_la_million_trees_goal")

double_goal <- c(land_use_data$scenario_parameters$scenario_description_2[c(1:2,4)]) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_tree_planting_land_cover(tb = land_use_data,
                                                    detail = FALSE,
                                                    .urban_form_scenario = .x,
                                                    .tree_planting_intervention = "double",
                                                    .tree_planting_per_capita = 0.26,
                                                    .tree_planting_per_hectare = 247) 
  ) %>%
  bind_rows(.id = "scenario_description_2") %>%
  mutate(intervention = "double")

pervious_goal <- c(land_use_data$scenario_parameters$scenario_description_2[c(1:2,4)]) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_tree_planting_land_cover(tb = land_use_data,
                                                    detail = FALSE,
                                                    .urban_form_scenario = .x,
                                                    .tree_planting_intervention = "tree_planting_on_all_pervious",
                                                    .tree_planting_per_capita = 0.26,
                                                    .tree_planting_per_hectare = 247) 
  ) %>%
  bind_rows(.id = "scenario_description_2") %>%
  mutate(intervention = "tree_planting_on_all_pervious")

la_goal %>%
  bind_rows(double_goal) %>%
  bind_rows(pervious_goal) %>%
  filter(ctu_name == "Lake Elmo") %>%
  pivot_longer(names_to = "landcover", values_to = "hectares", -c(scenario_description_2:year, total_area_hectares, intervention)) %>%
  ggplot(aes(x = intervention, y = hectares, fill = landcover)) + 
  geom_bar(stat = "identity") +
  facet_wrap(year~scenario_description_2) +
  theme_council_open() +
  scale_fill_brewer(palette = "Paired") +
  theme(axis.text.x = element_text(angle = 24)) +
  labs(title = "Lake Elmo") +
  # coord_flip() +
  labs(x = "")


# detail_y <- c(land_use_data$scenario_parameters$scenario_description_2[c(1)]) %>%
#     purrr::set_names() %>%
#     purrr::map(~ghg.sp::calc_tree_planting_land_cover(tb = land_use_data,
#                                                       detail = T,
#                                               .urban_form_scenario = .x,
#                                               .tree_planting_intervention = "match_la_million_trees_goal",
#                                               .tree_planting_per_capita = 0.26,
#                                               .tree_planting_per_hectare = 247) 
#                  ) %>%
#   bind_rows(.id = "scenario_description_2") %>%
#   mutate(intervention = "match_la_million_trees_goal") 
# 
# detail_n <- c(land_use_data$scenario_parameters$scenario_description_2[c(1)]) %>%
#     purrr::set_names() %>%
#     purrr::map(~ghg.sp::calc_tree_planting_land_cover(tb = land_use_data,
#                                                       detail = F,
#                                               .urban_form_scenario = .x,
#                                               .tree_planting_intervention = "match_la_million_trees_goal",
#                                               .tree_planting_per_capita = 0.26,
#                                               .tree_planting_per_hectare = 247) 
#                  ) %>%
#   bind_rows(.id = "scenario_description_2") %>%
#   mutate(intervention = "match_la_million_trees_goal") 
# 
# detail_y; detail_n
```

3) Why doesn't modifying the `.tree_planting_per_capita` function change the output for trees? Here showing for the tree planting on all pervious intervention.

```{r}
#| fig.width = 8,
#| fig.height = 5
pervious_goal_percapita <- seq(0,5, .5) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_tree_planting_land_cover(tb = land_use_data,
                                                    detail = FALSE,
                                                    .urban_form_scenario = "bau",
                                                    .tree_planting_intervention = "tree_planting_on_all_pervious",
                                                    .tree_planting_per_capita = .x,
                                                    .tree_planting_per_hectare = 247) 
  ) %>%
  bind_rows(.id = "scenario_description_2") %>%
  mutate(intervention = "tree_planting_on_all_pervious_bau")

pervious_goal_percapita %>%
  filter(ctu_name == "Lake Elmo") %>%
  pivot_longer(names_to = "landcover", values_to = "hectares", -c(scenario_description_2:year, total_area_hectares, intervention)) %>%
  ggplot(aes(x = scenario_description_2, y = hectares, fill = landcover)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~year) +
  theme_council_open() +
  scale_fill_brewer(palette = "Paired") +
  theme(axis.text.x = element_text(angle = 24)) +
  labs(subtitle = "Lake Elmo",
       x = "", title = "modifying `.tree_planting_per_capita` parameter")
```

4) Additionally, why doesn't modifying the `.tree_planting_per_hectare` function change the output for trees?

```{r}
#| fig.width = 8,
#| fig.height = 5
pervious_goal_perhectare <- seq(0, 500, 100) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_tree_planting_land_cover(tb = land_use_data,
                                                    detail = FALSE,
                                                    .urban_form_scenario = "bau",
                                                    .tree_planting_intervention = "tree_planting_on_all_pervious",
                                                    .tree_planting_per_capita = 0.26,
                                                    .tree_planting_per_hectare = .x) 
  ) %>%
  bind_rows(.id = "scenario_description_2") %>%
  mutate(intervention = "tree_planting_on_all_pervious_bau")

pervious_goal_perhectare %>%
  filter(ctu_name == "Lake Elmo") %>%
  pivot_longer(names_to = "landcover", values_to = "hectares", -c(scenario_description_2:year, total_area_hectares, intervention)) %>%
  ggplot(aes(x = scenario_description_2, y = hectares, fill = landcover)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~year) +
  theme_council_open() +
  scale_fill_brewer(palette = "Paired") +
  theme(axis.text.x = element_text(angle = 24)) +
  labs(subtitle = "Lake Elmo",
       x = "", title = "modifying `.tree_planting_per_hectare` parameter")
```



### Function to Calcuate Parking Lot Reduction Strategy

Function `ghg.sp::calc_parking_lot_land_cover()`.

1) hmmm, not seeing the reduction in parking lots with this function....

```{r}
#| fig.width = 8,
#| fig.height = 8
la_goal <- c(land_use_data$scenario_parameters$scenario_description_2[c(1:2,4)]) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_parking_lot_land_cover(
    tb = land_use_data,
    .urban_form_scenario = .x,
    .parking_lot_reduction_percentage = 0.8,
    .tree_planting_intervention = "match_la_million_trees_goal",
    .tree_planting_per_capita = 0.26,
    .tree_planting_per_hectare = 247,
    detail = FALSE
  )
  ) %>%
  bind_rows(.id = "scenario_description_2") %>%
  mutate(intervention = "match_la_million_trees_goal")

double_goal <- c(land_use_data$scenario_parameters$scenario_description_2[c(1:2,4)]) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_parking_lot_land_cover(
    tb = land_use_data,
    .urban_form_scenario = .x,
    .parking_lot_reduction_percentage = 0.8,
    .tree_planting_intervention = "double",
    .tree_planting_per_capita = 0.26,
    .tree_planting_per_hectare = 247,
    detail = FALSE
  )
  ) %>%
  bind_rows(.id = "scenario_description_2") %>%
  mutate(intervention = "double")


pervious_goal <- c(land_use_data$scenario_parameters$scenario_description_2[c(1:2,4)]) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_parking_lot_land_cover(
    tb = land_use_data,
    .urban_form_scenario = .x,
    .parking_lot_reduction_percentage = 0.8,
    .tree_planting_intervention = "tree_planting_on_all_pervious",
    .tree_planting_per_capita = 0.26,
    .tree_planting_per_hectare = 247,
    detail = FALSE
  )
  ) %>%
  bind_rows(.id = "scenario_description_2") %>%
  mutate(intervention = "tree_planting_on_all_pervious")

la_goal %>%
  bind_rows(double_goal) %>%
  bind_rows(pervious_goal) %>%
  filter(ctu_name == "Lake Elmo") %>%
  pivot_longer(names_to = "landcover", values_to = "hectares", -c(scenario_description_2:year, total_area_hectares, intervention)) %>%
  ggplot(aes(x = intervention, y = hectares, fill = landcover)) + 
  geom_bar(stat = "identity") +
  facet_wrap(year~scenario_description_2) +
  theme_council_open() +
  scale_fill_brewer(palette = "Paired") +
  theme(axis.text.x = element_text(angle = 24)) +
  labs(title = "Lake Elmo") +
  # coord_flip() +
  labs(x = "")




```

### Function to Calculate Conservation Tillage Strategy

*Q's: a couple things seem to be going on here...*

1) Why does agricultural_hectares decrease in year 2040 if the scenario is "compact_dev_with_drs"? Similarly, why does agricultural_hectares increase with "post_covid_sprawl" by year 2040?
2) Why are the other parameters (in the second figure) not associated with a year? 
3) Should something like "reduced_tractor_emissions_relative_to_current_conservation_tillage_mg_co2e_per_year" be set at zero for baseline year (2016), and then the value be multiplied by years (so year X - current year) to translate this into an annual measurement? Basically the question is how are these numbers going to be associated with a year so that emissions (reductions) can be shown? 
4) Because the amount of agricultural land increases sharply with the post_covid_sprawl in year 2040, it seems like some other variables are also behaving strangely. For instance, the "current_conservation_tillage_mg_c" is also much higher for post_covid_sprawl than it is for the bau. Shouldn't baseline (aka "current") be the same for both? This same comment holds true for the strange 2040 numbers in the compact_dev_with_drs (the numbers are so low!).

```{r}
#| fig.width = 8,
#| fig.height = 6
# choices = c(
#         "current_conservation_tillage",
#         "double_conservation_tillage",
#         "maximum_conservation_tillage"
#       )


current_goal <- c(land_use_data$scenario_parameters$scenario_description_2[c(1:2,4)]) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_conservation_tillage(
    tb = land_use_data,
    .urban_form_scenario = .x,
    .conservation_tillage_intervention = "current_conservation_tillage",
    .tree_planting_intervention = "match_la_million_trees_goal",
    .w2w_diesel_emission_factor_kg_co2e_per_gal = 12.50,
    .avoided_emissions_tractor_use_mg_co2e_per_hectare = 0.0102,
    .agricultural_land_carbon_stock_mg_c_per_hectare = 3,
    .maximum_soc_accumation_under_reduced_or_no_till_ag = 1.54,
    .tree_planting_per_capita = 0.26,
    .tree_planting_per_hectare = 247,
    .parking_lot_reduction_percentage = 0.8,
    detail = FALSE
  )
  ) %>%
  bind_rows(.id = "scenario_description_2") %>%
  mutate(intervention = "current_conservation_tillage")

double_goal <- c(land_use_data$scenario_parameters$scenario_description_2[c(1:2,4)]) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_conservation_tillage(
    tb = land_use_data,
    .urban_form_scenario = .x,
    .conservation_tillage_intervention = "double_conservation_tillage",
    .tree_planting_intervention = "match_la_million_trees_goal",
    .w2w_diesel_emission_factor_kg_co2e_per_gal = 12.50,
    .avoided_emissions_tractor_use_mg_co2e_per_hectare = 0.0102,
    .agricultural_land_carbon_stock_mg_c_per_hectare = 3,
    .maximum_soc_accumation_under_reduced_or_no_till_ag = 1.54,
    .tree_planting_per_capita = 0.26,
    .tree_planting_per_hectare = 247,
    .parking_lot_reduction_percentage = 0.8,
    detail = FALSE
  )
  ) %>%
  bind_rows(.id = "scenario_description_2") %>%
  mutate(intervention = "double_conservation_tillage")


max_goal <- c(land_use_data$scenario_parameters$scenario_description_2[c(1:2,4)]) %>%
  purrr::set_names() %>%
  purrr::map(~ghg.sp::calc_conservation_tillage(
    tb = land_use_data,
    .urban_form_scenario = .x,
    .conservation_tillage_intervention = "maximum_conservation_tillage",
    .tree_planting_intervention = "match_la_million_trees_goal",
    .w2w_diesel_emission_factor_kg_co2e_per_gal = 12.50,
    .avoided_emissions_tractor_use_mg_co2e_per_hectare = 0.0102,
    .agricultural_land_carbon_stock_mg_c_per_hectare = 3,
    .maximum_soc_accumation_under_reduced_or_no_till_ag = 1.54,
    .tree_planting_per_capita = 0.26,
    .tree_planting_per_hectare = 247,
    .parking_lot_reduction_percentage = 0.8,
    detail = FALSE
  )
  ) %>%
  bind_rows(.id = "scenario_description_2") %>%
  mutate(intervention = "maximum_conservation_tillage")

current_goal %>%
  bind_rows(double_goal) %>%
  bind_rows(max_goal) %>%
  filter(ctu_name == "Lake Elmo") %>%
  tidyr::pivot_longer(
    names_to = "var",
    values_to = "value",
    cols = -c(ctu_name, co_name, scenario_description_2, intervention)) %>%
  filter(str_detect(var, "_year_")) %>%
  ggplot(aes(x = var, y = value, fill = intervention)) + 
  geom_bar(stat = "identity", position = position_dodge(width = .9)) +
  facet_wrap(~scenario_description_2) +
  theme_bw() +
  scale_fill_brewer(palette = "Paired") +
  theme(axis.text.x = element_text(angle = 24, hjust = 1)) +
  labs(title = "Ag hectares",
       subtitle = "Lake Elmo",
       x = "", y = "hectares")

current_goal %>%
  bind_rows(double_goal) %>%
  bind_rows(max_goal) %>%
  filter(ctu_name == "Lake Elmo") %>%
  tidyr::pivot_longer(
    names_to = "var",
    values_to = "value",
    cols = -c(ctu_name, co_name, scenario_description_2, intervention)) %>%
  filter(!str_detect(var, "_year_")) %>%
  ggplot(aes(x = var, y = value, fill = intervention)) + 
  geom_bar(stat = "identity", position = position_dodge(width = .9)) +
  facet_wrap(~scenario_description_2) +
  theme_bw() +
  scale_fill_brewer(palette = "Paired") +
  theme(axis.text.x = element_text(angle = 24, hjust = 1)) +
  labs(subtitle = "Lake Elmo",
       x = "", y = "")

```


# Additional comments  

- In **$scenario_parameters** can you explain why all scneario_description_1 have exurban_development == 0? If exurban_development doesn't vary between the scenarios, can this parameter simply be removed? Or does a new scenario need to get created that leverages this exurban development?
- PS, you probably want to get rid of the number suffix on the column names; ie "scneario_description_1" becomes "scneario_description". There is also a typo where scneario should be scENario; not sure where all you refer to this though, so I'll let you fix it because you'll know where all to check and update the name to make sure that fixing the typo doesn't break stuff.
- Which year is the current_conservation_tillage_county data pulling? The document says: For example: What percentage of agricultural land was using conservation tillage in the Washington County in 201? Is this 2017 (date on the ag census) or another year?

# STOP REVIEW HERE 


----


```{r}
ghg.sp::calc_carbon_stock_per_ctu(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  .parking_lot_reduction_percentage = 0.8,
  .conservation_tillage_intervention = "current_conservation_tillage",
  .w2w_diesel_emission_factor_kg_co2e_per_gal = 12.50,
  .agricultural_land_carbon_stock_mg_c_per_hectare = 3,
  .maximum_soc_accumation_under_reduced_or_no_till_ag = 1.54,
  .avoided_emissions_tractor_use_mg_co2e_per_hectare = 0.0102,
  detail = FALSE
) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 10) %>%
  kable()
```

