---
title: "Introduction to Building Energy Module"
author: Mauricio Leon, Senior Researcher, Metropolitan Council
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro_to_building_energy_module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Background

The *Metropolitan Council Greenhouse Gas Scenario Planning Tool* estimates business-as-usual greenhouse gas emissions and greenhouse gas mitigation strategies by city/township for the Twin Cities Region. The tool covers greenhouse gas emissions from the transportation, building energy, and land use sectors.

# Building Energy Module

The Building Energy Module of the *Metropolitan Council Greenhouse Gas Scenario Planning Tool* 
consists of a modified version of the *Zero Carbon Calculator for Communities* created by the Sustainable Healthy 
Cities Network (SHCN). The original concept of the tool was provided by SHCN as a Microsoft Excel Spreadsheet submitted to the Metropolitan Council on October of 2021, which 
was later adapted into an R Package. 

The module allows users to explore strategies for decarbonization within the building energy sector.

## Strategies

The GHG mitgation strategies included in the Building Energy Module include:

1. **Metroplitan Council's Planned Land Use**: Multi-family housing effects from the planned land use by the Metropolitan Council. This strategy consists of an estimate of the GHG emissions impact of increasing multifamily
housing in accordance with the Metropolitan Council 2040 Planned Land Use estimates.

2. **More Compact Residential Buildings**: This strategy goes beyond the Metroplitan Council’s 2040 Planned Land Use, by assuming that 50% of all new single-family construction will now become multi-family, with the multi-family homes having smaller square footage. 

3. **Increased Affordable Floor Area**: This strategy assumes that about 50% of the single-family homes will respond to increased energy prices by decreasing their living space. 

4. **New Homes are LEED-Gold**: This strategy calculates the effect of 50% of new homes being constructed to LEED Gold standards.

5. **Existing Homes Retrofits**:This strategy focuses on existing homes and assumes 80% of them are retrofitted to performance-based high energy efficiency standards, which reduce energy use by 33%, which is mid-way between the trends in EUI already seen, and a passive house standard. Most of these cases have on-site energy use for renewables, so they conflate renewables with efficiency. 

# Techical Documentation

All the functions for the Building Energy Module are contained within the `ghg.sp` package.

```{r setup}
library(ghg.sp)
```

## Source Datasets

The `building_data.rda` is a preloaded dataset that includes cleaned datasets to 
run the building energy module of the `ghg.sp` package. 

The R scripts to produce `building_data.rda` can be found on the `/raw-data folder` of the `ghg.sp` package.

The source tables for producing the `building_data.rda` are stored in the Metropolitan Council `Emissions`
database, on the Council's Microsoft SQL Server. 

The following chunk shows the names of the tables contained in `building_data.rda` dataset.

```{r}
names(building_data)
```
As seen above, the `building_data` dataset contains two tables. 
`residential` and `non_residential`.

These two tables contain all the key variables necessary to run the functions of the building
energy module. 

Next, we will go in depth on the two main tables of `building_data.rda`


### Residential Energy

```{r}
names(building_data$residential)
```


The following chunk prints the main key variables for the building energy module for
the city of Lake Elmo

Note that when there is missing data for the `multifamily_average_floor_area_sqft_ctu` the 
package will refer to the county wide average. 

```{r}
building_data$residential %>% 
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  tidyr::pivot_wider(names_from = year, values_from = value) %>%
  dplyr::select(ctu_name, var, `2018`, `2040`) %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.numeric, round, 2)
```

### Non-Residential Energy 

The following chunk prints the key variables for the commercial/industrial strategies 
under the building energy module. 

```{r}
building_data$non_residential %>% 
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  tidyr::pivot_wider(names_from = year, values_from = value) %>%
  dplyr::select(ctu_name, var, `2018`, `2040`) %>%
  dplyr::ungroup() %>%
  dplyr::mutate_if(is.numeric, round, 2)
```

