---
title: "Introduction to the Green Infrastructure & Land Use Module"
author: Mauricio Leon, Senior Researcher, Metropolitan Council
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the Green Infrastructure & Land Use Module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Background

The *Metropolitan Council Greenhouse Gas Scenario Planning Tool* estimates business-as-usual greenhouse gas emissions and greenhouse gas mitigation strategies by city/township for the Twin Cities Region. The tool covers greenhouse gas emissions from the transportation, building energy, and land use sectors.

## Land Use & Green Infrastructure Module

The Land Use & Green Infrastructure Module of the *Metropolitan Council Greenhouse Gas Scenario Planning Tool* 
consists of a modified version of the *Zero Carbon Calculator for Communities* created by the Sustainable Healthy 
Cities Network (SHCN). The original concept of the tool was provided by SHCN as a Microsoft Excel Spreadsheet submitted to the Metropolitan Council on October of 2021, which 
was later adapted into an R Package. 

The module allows users to explore strategies for decarbonization within the land use and greenery of communities.

## Strategies

The GHG mitigation strategies included in the Land Use & Green Infrastructure Module include:


1. **Urban Form (LU.1)**: The impact of changing the urban form (the land use and land cover composition) of the city/township to more compact residential and mixed use development. Reduction of greenhouse gas emissions occurs when there is more land available to greenery rather than impervious surfaces. 
2. **Increase Tree Canopy (LU.2)**: The impact of increasing the tree canopy area of the city/township by a given amount. For example, what is the effect of planting trees on all pervious surfaces available in the city/township. 
3. **Reduced Parking Lot Area (LU3)**: The impact of reducing the parking lot area of the city/township by a given percentage. Reducing parking lot increases the availability of surface area for tree planting. 
4. **Conservation Tillage (LU4)**: The impact of conservation tillage on the emissions of agricultural land of the city/township. Sustainable agriculture practice can have a significant effect on a city /township emissions. 

Some key concept to understand the Land Use Module are the following:

* Land Use
* Land Cover
* Conservation Tillage
* Development Types
* Exurban Development
* Urban Infill Development
* Urban Expansion Development 

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>")
options(rmarkdown.html_vignette.check_title = FALSE)
```

# Techical Walkthrough

All the functions for the Land Use Module are contained within the `ghg.sp` package.
```{r setup}
library(ghg.sp)
library(knitr)
```

## Land Use and Land Cover Data

The package provided `land_use_data.rda` dataset is required for the land use module to run.
To explore the dataset, run the following command:
`ghg.sp::land_use_data`.

The source of these tables is the `Emissions` database of the Metropolitan Council. 

Most of these tables are derived from GIS analysis of the 2016 Generalized Land Use as well as the Land Cover of the Metropolitan Region. 

The following code provides a list of the tables within the `land_use_data` dataset.
```{r}
names(land_use_data) %>%
  kable(col.names = "tables within `land_use_data`")
```

### Table Descriptions
The following is a description of the content of all input tables of the land use module, which are
contained in the `land_use_data` dataset.

**$ctu_county**: 
Table shows the relationship between city/townships and county. 
*For example: What county is Lake Elmo within?*
```{r echo=FALSE}
land_use_data$ctu_county %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 10) %>%
  kable()

# # ee test
# library(tidyverse); library(ghg.sp)
# land_use_data$ctu_county %>%
#   count(ctu_name) %>%
#   filter(n > 1)
# 
# land_use_data$ctu_county %>% 
#   filter(ctu_name == "St. Anthony")
```

**$ctu_forecast**: 
Table shows the demographic forecasts for city/township from Community Profiles.
*For example: What is the forecast population of Lake Elmo in 2040?*
```{r echo=FALSE}
land_use_data$ctu_forecast %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 10)  %>%
  kable()
```

**$scenario_parameters**: 
Table provides key parameters for the different urban form scenarios of the land use module.
These parameters show relative assumed proportions for the the intensity of land by three development types: *exurban development*, *urban expansion*, and *urban infill*. These parameter vary for each of the scenario available to the user. 
*For example, a value of 0.5 will represent half of the given development type for the given scenario. A value of 2 will represent double. *
```{r echo=FALSE}
land_use_data$scenario_parameters %>%
  kable()
```

**$land_cover_percentages_filled**: 
Table shows the percentage of land use by land cover type in the Metropolitan Region. 
*For example: What percent of impervious surfaces can be classified single family residential in the city of Lake Elmo?*
```{r echo=FALSE}
land_use_data$land_cover_percentages_filled %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::filter(land_cover_description_2 == "parking_lot") %>%
  head(., n = 15) %>%
  kable()
```

**$land_composition_ctu**: 
Table provides the percentages of land use composition categorized in to three types of development: *exurban development*, *urban expansion*, and *urban infill*. 
*For example: What percentage or what number of hectares of exurban development land area corresponds to parking lot surfaces within agricultural in the city of Lake Elmo for the forecast year 2040?*
```{r echo=FALSE}
land_use_data$land_composition_ctu %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::filter(year == 2040) %>%
  head(., n = 60) %>%
  kable()
```

**$current_conservation_tillage_county**:
Tables shows the percent of agricultural land that is managed with conservation tillage by county. This is a necessary assumption for the conservation tillage strategy. 
*Data comes from the Census of Agriculture (2017).*
*For example: What percentage of agricultural land was using conservation tillage in the Washington County in 201?*
```{r echo=FALSE}
land_use_data$current_conservation_tillage_county %>%
  head(., n = 10) %>%
  kable()
```

**$land_by_development_type_sum_bau**:
Table shows the forecast total number of hectares by city/township for the business as usual scenario. 
*For example: What is the forecast number of hectare for the city of Lake Elmo in 2040?*
```{r echo=FALSE}
land_use_data$land_by_development_type_sum_bau %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 10) %>%
  kable()
```

**$multifamily_mixed_area**: 
Table show the estimate of the number of hectares of multifamily mixed area by city/township for three development types: *exurban development*, *urban expansion*, and *urban infill*.
*For example: What is the forecast hectares of multifamily or mixed use land use for the year 2040 for the city of Lake Elmo?*
```{r echo=FALSE}
land_use_data$multifamily_mixed_area %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 10) %>%
  kable()
```

## Functions

The impact of different land use and greenery strategies is estimated using a series of functions with arguments provided by the user.

The key principle is of this model is that different land cover types have different carbon stock and sequestration value. By altering the composition of a city/township, we could observe higher or lower overall carbon stocks or sequestration value per year. 

### Functions to Calculate Urban Form Strategy. 

The first function is `ghg.sp::calc_by_development_type()`
For help run the following command `?ghg.sp::calc_by_development_type()`
This function is intermediary and does not provide the final result. 

`ghg.sp::calc_by_development_type()` 
The function returns a table with the land composition of different urban form scenarios of land use by development type.

The tool will  perform a series of calculations in 
response to different values of argument `.urban_form_scenario`.
For example, if the selected input is `"bau"`, or the business as usual, 
the tool will select the table it needs. 

The possible scenarios to select on the `.urban_form_scenario` argument include:
* `"bau"`: the business as usual scenario.
* `"post_covid_sprawl"`: post-covid sprawl scenario.
* `"compact_development_beyond_bau"`: compact development beyond the business as usual scenario.
* `"compact_development_with_drs"`: compact development with dynamic ride sharing.

The following code prints the value of function `ghg.sp::calc_by_development_type()` filtered to show only the city of Lake Elmo. 
```{r}
ghg.sp::calc_land_by_development_type(tb = land_use_data,
                                      .urban_form_scenario = "bau") %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  kable()
```

The second function is `ghg.sp::calc_scen_land_use()`
The following code runs the `ghg.sp::calc_scen_land_use()` for the `"bau"` or business as usual scenario. 
```{r}
ghg.sp::calc_scen_land_use(tb =  land_use_data,
                           .urban_form_scenario = "bau") %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 70) %>%
  kable()
```

The third function is `ghg.sp::calc_land_cover_by_land_use()`
The following code runs the `ghg.sp::calc_land_cover_by_land_use()` for the `"bau"` or business as usual scenario.
```{r}
ghg.sp::calc_land_cover_by_land_use(tb =  land_use_data,
                                    .urban_form_scenario = "bau") %>%
  dplyr::filter(year == 2040) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::filter(land_cover_description_2 == 	"forest") %>%
  head(., n = 25)  %>%
  kable()
```

The fourth function is `ghg.sp::calc_land_cover_by_city()`
This function returns the urban form scenario land composition in hectares without adding additional strategies. This is the last function used to estimate the impact of urban form changes alone.  
```{r}
ghg.sp::calc_land_cover_by_city(tb = land_use_data,
                                .urban_form_scenario = "bau") %>%
  dplyr::filter(year == 2040) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 15) %>%
  kable()
```

### Function to Calculate Tree Planting Strategy

Having calculated the hectares of land cover under the selected urban form scenario, we now will add the change to the land composition  that is cause by implement a tree planting strategy. 

Function `ghg.sp::calc_tree_planting_land_cover()`.
```{r}
ghg.sp::calc_tree_planting_land_cover(
  tb = land_use_data,
  detail = FALSE,
  .urban_form_scenario = "bau",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247
)  %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::filter(year == 2040) %>%
  tidyr::pivot_longer(
    names_to = "var",
    values_to = "value",
    cols = -c("ctu_name", "year")
  ) %>%
  head(., n = 15) %>%
  kable()
```

### Function to Calcuate Parking Lot Reduction Strategy

Having calculate the hectares of land cover under the selected urban form and tree planting scenario, we now will add the change to the land composition that is cause by implementing a reduced parking lot strategy. 

Function `ghg.sp::calc_parking_lot_land_cover()`.
```{r}
ghg.sp::calc_parking_lot_land_cover(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .parking_lot_reduction_percentage = 0.8,
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  detail = FALSE
)  %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::filter(year == 2040) %>%
  tidyr::pivot_longer(
    names_to = "var",
    values_to = "value",
    cols = -c("ctu_name", "year")
  ) %>%
  head(., n = 15) %>%
  kable()
```

### Function to Calculate Conservation Tillage Strategy

```{r}
ghg.sp::calc_conservation_tillage(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .conservation_tillage_intervention = "current_conservation_tillage",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .w2w_diesel_emission_factor_kg_co2e_per_gal = 12.50,
  .avoided_emissions_tractor_use_mg_co2e_per_hectare = 0.0102,
  .agricultural_land_carbon_stock_mg_c_per_hectare = 3,
  .maximum_soc_accumation_under_reduced_or_no_till_ag = 1.54,
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  .parking_lot_reduction_percentage = 0.8,
  detail = FALSE
)  %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  tidyr::pivot_longer(
    names_to = "var",
    values_to = "value",
    cols = -c("ctu_name", "co_name")
  ) %>%
  head(., n = 15) %>%
  kable()
```

---

# STOP REVIEW HERE 


----


```{r}
ghg.sp::calc_carbon_stock_per_ctu(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  .parking_lot_reduction_percentage = 0.8,
  .conservation_tillage_intervention = "current_conservation_tillage",
  .w2w_diesel_emission_factor_kg_co2e_per_gal = 12.50,
  .agricultural_land_carbon_stock_mg_c_per_hectare = 3,
  .maximum_soc_accumation_under_reduced_or_no_till_ag = 1.54,
  .avoided_emissions_tractor_use_mg_co2e_per_hectare = 0.0102,
  detail = FALSE
) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 10) %>%
  kable()
```

