---
title: "Introduction to the Green Infrastructure & Land Use Module"
author: Mauricio Leon, Senior Researcher, Metropolitan Council
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro_to_land_use_module}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Background

The *Metropolitan Council Greenhouse Gas Scenario Planning Tool* estimates business-as-usual greenhouse gas emissions and greenhouse gas mitigation strategies by city/township for the Twin Cities Region. The tool covers greenhouse gas emissions from the transportation, building energy, and land use sectors.

## Land Use & Green Infrastructure Module

The Land Use & Green Infrastructure Module of the *Metropolitan Council Greenhouse Gas Scenario Planning Tool* 
consists of a modified version of the *Zero Carbon Calculator for Communities* created by the Sustainable Healthy 
Cities Network (SHCN). The original concept of the tool was provided by SHCN as a Microsoft Excel Spreadsheet submitted to the Metropolitan Council on October of 2021, which 
was later adapted into an R Package. 

The module allows users to explore strategies for decarbonization within the land use and greenery of communities.

The GHG mitgation strategies included in the Land Use & Green Infrastructure Module include:

1. *Increase Tree Canopy*:
2. *Reduced Parking Lot Area*:
3. *Conservation Tillage*:

## Land Use Data

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ghg.sp)
```

The `land_use_data.rda` data set is required for the land use module to run. 

The source of these tables is the `Emissions` database. 

```{r}
names(land_use_data)
```
## CTU Forecast

```{r}
names(land_use_data$ctu_forecast)
```


The first function is `calc_by_development_type`

It provides a series of scenarios of land use by development type

```{r}
names(
  ghg.sp::calc_land_by_development_type(
    tb = land_use_data,
    .scenario = "bau",
    .luse_scen = "compact_dev_with_drs"
  )
)
```

## The Business as Usual Scenario

How is the Business as Usual Obtained? 

The estimates for the land by development type for the business as usual are based on the ...

```{r}
ghg.sp::calc_land_by_development_type(tb = land_use_data,
                                      .scenario = "bau",
                                      .luse_scen = "compact_dev_with_drs")$bau_total
```

