---
title: "Minneapolis Greenhouse Gas Emissions Scenario Planning"
author: "Metropolitan Council"
format:
  html:
    toc: true
    html-math-method: katex
    css: styles.css
    mainfont: "Helvetica"
    df-print: kable
---

# Minneapolis Greenhouse Gas Emissions Scenario Planning

```{r, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>", echo = FALSE, fig.width = 6
)
```

```{r}
#| echo: false
#| message: false
library(ghg.sp)
library(tidyverse)
library(knitr)
library(kableExtra)
```

```{r}
#| echo: false
#| message: false
options(scipen = 1, digits = 2)


kableformat <- function(data, .caption) {
  data %>%
    dplyr::left_join(., variables, by = "var") %>%
    dplyr::select(c(desc, value)) %>%
    dplyr::rename("Variables" = "desc",
                  "Value" = "value") %>%
    kableExtra::kable(
      caption = .caption,
      format = "html",
      escape = FALSE,
      format.args = list(big.mark = ",", digits = 6)
    ) %>%
    kableExtra::kable_classic(html_font = "Arial",
                              full_width = F,
                              font_size = 12)
}
```

```{r variables}
#| echo: false
#| message: false
variables <-
  data.frame(
    var = c(
      "households",
      "population",
      "total_jobs",
      "jobs",
      "commercial_jobs",
      "industrial_jobs",
      "single_family_units",
      "multifamily_units",
      "single_family_average_floor_area_sqft_ctu",
      "multifamily_average_floor_area_sqft_ctu",
      "single_family_average_floor_area_sqft_county",
      "multifamily_average_floor_area_sqft_county",
      "commercial_workers_county",
      "industrial_workers_county",
      "residential_mwh",
      "residential_ng_therms",
      "residential_elec_emis_t_co2e",
      "residential_ng_emis_t_co2e",
      "residential_kwh_per_floor_area",
      "residential_therms_per_floor_area",
      "residential_mwh_per_households",
      "residential_therms_per_households",
      "total_residential_kwh_forecast",
      "total_residential_therms_forecast",
      "commercial_mwh", 
      "commercial_therms", 
      "industrial_mwh", 
      "industrial_therms", 
      "commercial_mwh_per_worker",
      "industrial_mwh_per_worker",
      "commercial_therm_per_worker",  
      "industrial_therm_per_worker",
      "commercial_elec_emis_t_co2e",  
      "commercial_ng_emis_t_co2e", 
      "industrial_elec_emis_t_co2e",
      "industrial_ng_emis_t_co2e",
      "kg_per_mwh",
      "kg_per_therm",
      "residential_electricity_emissions_kg_co",
      "residential_natural_gas_emissions_kg_co",
      "residential_therms",
      "total_residential_emissions"
    ),
    desc = c(
      "Households",
      "Population",
      "Jobs",
      "Total Jobs",
      "Commercial Jobs",
      "Industrial Jobs",
      "Number of Single Family Units",
      "Number of Multifamily Units",
      "Single Family Average Floor Area (ft<sup>2</sup>)",
      "Multifamily Average Floor Area (ft<sup>2</sup>)",
      "Single Family Average Floor Area for the County (ft<sup>2</sup>)",
      "Multifamily Average Floor Area for the County (ft<sup>2</sup>)",
      "Commercial Jobs",
      "Industrial Jobs",
      "Residential Electricity (MWh)",
      "Residential Natural Gas (Therms)",
      "Residential Electricity Emissions (tonnes CO<sub>2</sub>e)",
      "Residential Natural Gas Emissions (tonnes CO<sub>2</sub>e)",
      "Residential Electricity per Floor Area (KWh/ft<sup>2</sup>)",
      "Residential Therms per Floor Area (therms/ft<sup>2</sup>)",
      "Residential Electricity per Household (KWh/household)",
      "Residential Therms per Floor Household (therms/household)",
      "Residential Electricity - Forecast (MWh)",
      "Residential Natural Gas - Forecast (Therms)",
      "Commercial Electricity (MWh)",
      "Commercial Natural Gas (Therms)",
      "Industrial Electricity (MWh)",
      "Industrial Natural Gas (Therms)",
      "Commercial Electricity per Worker (KWh/worker)",
      "Industrial Electricity per Worker (KWh/worker)",
      "Commercial Natural Gas per Worker (therms/worker)",
      "Industrial Natural Gas per Worker (therms/worker)",
      "Commercial Electricity Emissions (tonnes CO<sub>2</sub>e)",
      "Commercial Natural Gas Emissions (tonnes CO<sub>2</sub>e)",
      "Industrial Electricity Emissions (tonnes CO<sub>2</sub>e)",
      "Industrial Natural Gas Emissions (tonnes CO<sub>2</sub>e)",
      "Emissions Factor for Electricity (kg/MWh)",
      "Emissions Factor for Natural Gas (kg/therm)",
      "Residential Electricity Emissions (kg CO<sub>2</sub>)",
      "Residential Natural Gas Emissions (kg CO<sub>2</sub>)",
      "Residential Natural Gas (therms)",
      "Total Residential Emissions (kg CO<sub>2</sub>)"
    )
  )
```

## Disclaimer 

The purpose of this report is to attempt to estimate the potential of greenhouse gas mitigation strategies at the City scale. The report should be used to inform conversations and for educational purposes. However, the information presented does not represent any formal recommendations or policy decisions.

## Executive Summary

The current report was prepared by the Metropolitan Council for the City of Minneapolis. The report presents a portfolio of strategies that the City of Minneapolis could take to reduce greenhouse gas emissions under a Business as Usual Scenario. The strategies can be categorized into three sectors 1) stationary energy, 2) transportation, and 3) land use and forestry. There might be differences between this report's greenhouse gas emissions estimates and the emissions reported in the Minneapolis Greenhouse Gas Inventory. This is because the report relies in a different methodology to estimate emissions. 

## 1. Introduction

Cities across the Minnesota Metro Region have recognized the urgency of taking action to reduce greenhouse gas emissions and increase community resilience towards climate change.

The Metropolitan Council seeks to expand the information and technical assistance it provides to local governments to support regional and local climate change planning. The Council offers tools and guidance to help get cities on track to meet local, regional, and state goals. This report is intended to support the City of Minneapolis to consider potential strategies for decarbonization.

To alter the input assumptions on this report please contact the Metropolitan Council (mauricio.leon\@metc.state.mn.us).

### 1.1 Greenhouse Gas Emissions

Greenhouse gases (GHG) trap heat in the Earth's atmosphere and contribute to global warming. Carbon dioxide, methane, and nitrous oxide are the most common greenhouse gases.

Human activities have been the primary driver of climate change by releasing greenhouse gases in large amounts. Combustion of fossil fuels (e.g., coal, oil, and gas) releases greenhouse gases, which trap heat in the atmosphere, causing temperatures to rise.

The primary sources of emissions in cities come from building energy use (electricity and natural gas), transportation (gas and diesel), and waste (methane at landfills or combustion for energy generation).

The transportation sector, which includes cars, trucks, ships, and airplanes, is the largest source of carbon dioxide emissions in the United States, followed by the electricity and heat production and industrial sectors. Other sources of greenhouse gas emissions in the United States include agriculture, waste, and industrial processes.

### 1.2 Impact of Climate Change in Minnesota

Climate change is already occurring and impacting Minnesotans. As the Earth's temperature continues to rise, these changes are expected to increase and intensify. Vulnerable populations, infrastructure, and natural systems will be most impacted. The Minnesota Pollution Control Agency highlights the following observations and potential impacts:

-   Temperatures are rising. Nights are warming faster than days.
-   Winter is warming faster than summer.
-   Increase in average number of days above 100°F.
-   Less frequency of extreme cold.
-   Increased freeze/thaw cycles, stressing infrastructure.
-   Increased drought cycles.
-   More intense flooding.
-   Extreme storm events.
-   Impacts on ecosystem.

### 1.3 Scenario Planning

Scenario planning, also known as scenario analysis, is a method to make flexible long-term plans. It is a way of thinking about and preparing for the future by considering possible events and their potential impact on a city or region. Scenario planning involves creating multiple possible future scenarios and then considering how to respond to each one. This method allows organizations to be more flexible and adaptable in uncertainty and change.

The Metropolitan Council has developed a Greenhouse Gas Scenario Planning Tool to help metro cities determine pathways to achieving emissions reductions. The Tool builds upon the DECAL model created by the Sustainable Healthy Cities Network with the support and funding of the Metropolitan Council.

The Tool looks at strategies across multiple sectors and calculates the emissions reductions that can result from successful implementation. These sectors include compact land use, mobility, buildings, waste, and green infrastructure.

The Metropolitan Council will offer the scenario planning tool on an easy-to-use online platform. Cities can use this tool to identify solutions tailored to their communities by testing different scenarios across strategies.

The following report compiles results for the City of Minneapolis based on the Greenhouse Gas Scenario Planning Tool created by the Metropolitan Council and adapted from the collaborative work between the Sustainable Healthy Cities Network and the Metropolitan Council.

## 2. Demographic Assumptions

The Greenhouse Gas Scenario Planning Tool uses the best available demographic assumptions for cities in the Metropolitan Region in order to forecast the emissions in 2040.

#### 2.1 City-level Demographic Assumptions

The households, population, and jobs data is obtained from the Metropolitan Council, which in terms obtains it from the [US Census and American Community Survey](https://www.census.gov/programs-surveys/acs) (ACS) a nationwide survey conducted by the United States Census Bureau to collect data on a variety of social, economic, and demographic characteristics of the American population.

##### Table 1

```{r tab1}
#| echo: false
get_demographic_baseline(tb = building_energy_data)$ctu %>%
  dplyr::ungroup() %>%
  dplyr::filter(ctu_name == "Minneapolis") %>%
  kableformat(., "Basic Demographic Assumptions for City of Minneapolis for 2018")
```

In cases when there are data gaps for city-level data, the Greenhouse Gas Scenario Planning Tool defaults to average county-level assumptions. `Table 2` shows the relevant assumptions for Hennepin County.

##### Table 2

```{r tab2}
#| echo: false
get_demographic_baseline(tb = building_energy_data)$county %>%
  dplyr::ungroup() %>%
  dplyr::filter(co_name == "Hennepin") %>%
  kableformat(., "Basic Demographic Assumptions Hennepin County for 2018")  
```

#### 2.2 City-level Demographic Forecast

The Metropolitan Council generates a forecast of population, households, and employment for all cities in the seven county metropolitan region.

The floor area assumption are an extrapolation of recent historic trends in floor area growth.

##### Table 3

```{r tab3}
#| echo: false
calc_demographic_forecast(tb = building_energy_data)$ctu %>%
  dplyr::ungroup() %>%
  dplyr::filter(ctu_name == "Minneapolis")  %>%
  kableformat(., .caption = "Forecast Demographic Assumptions for City of Minneapolis for 2040")
```

Figure 1 shows the comparisson of population for the City of Minneapolis in 2018 with the 2040 forecast.

##### Figure 1

```{r fig1}
#| echo: false
df <- dplyr::bind_rows(
  get_demographic_baseline(tb = building_energy_data)$ctu %>%
    dplyr::filter(ctu_name == "Minneapolis"),
  calc_demographic_forecast(tb = building_energy_data)$ctu %>%
    dplyr::filter(ctu_name == "Minneapolis")
) %>% dplyr::filter(var == "population")

# make ggplot2 barplot with dataset 'df' and x-axis 'year' and y-axis 'value', with year being a factor
 ggplot(df, aes(x = as.factor(year), y = value)) +
   geom_bar(stat = "identity") +
   theme(panel.background = element_blank())
```

------------------------------------------------------------------------

## 3. Reducing Emissions in the Building Energy Sector

Strategies to reduce emissions in the building energy sector include, increasing efficiency, decarbonizing electricity, electrifying heat, using sustainable construction practices, and changing behaviors.

The emissions from electricity were calculated using an emissions factor of \_\_\_ per MWh of electricity.

### 3.1 Residential Energy Assumptions

The biggest share of residential electricity use is typically for heating and cooling. In the United States, heating and air conditioning account for about half of the average home's energy use. Therefore, for the residential sector, electricity and natural gas used tend to be closely correlated with the square footage of a building.

#### 3.1.1 Residential Energy Baseline

Table 4 shows the basic assumptions about electricity and natural gas use for the baseline year of 2018 for the City of Minneapolis.

##### Table 4

```{r tab4}
#| echo: false
ghg.sp::get_residential_energy_baseline() %>%
  dplyr::ungroup() %>%
  dplyr::filter(ctu_name == "Minneapolis") %>%
  kableformat(., .caption = "Residential Energy Assumption for Minneapolis 2018")
```

#### 3.1.2 Residential Energy Forecast

Table 5 shows the forecast assumption for the residential energy sector.

##### Table 5
```{r tab5}
#| echo: false
ghg.sp::calc_residential_energy_forecast() %>%
  dplyr::ungroup() %>%
  dplyr::filter(ctu_name == "Minneapolis") %>%
  dplyr::mutate(value = dplyr::if_else(var == "total_residential_kwh_forecast", value / 1000, value)) %>%   kableformat(., .caption = "Residential Energy Assumption for Minneapolis 2018")
```

## 3.2 Residential Energy Strategies

### 3.2.1 Reducing Residential Floor Area and Improving Efficiency

#### 3.2.1.a Affordable Floor Area

Cities can control the size of new single-family homes to reduce energy use. 

The growth of the single-family home size is reduced from 15% to 5% to make the new single-family home smaller. We apply this impact to 50% of new single-family homes. Cities can adjust the growth rate of single-family homes' floor area and the percentage of new single-family homes being controlled.

##### Equation 1
$$
S_1sqft_{2040}^{sf} = (units_{2040}^{sf} - units_{2018}^{sf}) \cdot X_1 \cdot (\bar{sqft}_{2040}^{sf}- ( 1 + \frac{\Delta sqft}{year}) \cdot \bar{sqft}_{2018})
$$

##### Table 6

```{r tab6}
#| echo: false
ghg.sp::calc_affordable_floor_area(
  res_tb = building_data$residential,
  .single_family_floor_area_growth_pct = 0.05) %>%
  dplyr::ungroup() %>%
  dplyr::filter(ctu_name == "Minneapolis") %>%
  kableformat(., .caption = "Affordable Floor Area Assumptions for Minneapolis 2018")
```

#### 3.2.1.b New Home Efficiency

Cities can adopt strict building codes requiring new homes to achieve LEED Gold (or similar) standards. We assumed `50%` of new homes (single and multifamily homes) to be LEED-Gold with both electricity and gas energy use intensity reduced `64%`. 

Cities can adjust the percentage of new homes achieving this standard.

Existing home: `80%` reduced EUI by `33%`.

##### Table 7
```{r tab7}
#| echo: false
ghg.sp::calc_floor_area_leed(
  res_tb = building_data$residential,
   .new_homes_leed_gold_pct = 0.5,
   .enviro_factors = enviro_factors) %>%
  dplyr::ungroup() %>%
  dplyr::filter(
    year == 2040,
    var %in% c(
       "multifamily_average_floor_area_sqft_county", 
       "multifamily_units",
       "single_family_units",
       "single_family_average_floor_area_sqft_ctu"
       )) %>%
  dplyr::filter(ctu_name == "Minneapolis") %>%
  kableformat(., .caption = "New Home Efficiency Minneapolis 2040")
``` 

##### Equation 2
$$
S_2sqft_{2040}^{sf} = (units_{2040}^{sf} - units_{2018}^{sf}) \cdot x \cdot (S_1sqft \cdot 0.64 \cdot \bar{sqft}_{2018}^{sf})
$$

#### 3.2.1.c Existing Home Efficiency

Cities can retrofit existing homes to be energy efficient. We assume `80%` of existing homes with energy use intensity reduced by `33%`. 

Cities can adjust the percentage of existing homes to be retrofitted.


##### Table 8

```{r tab8}
#| echo: false
ghg.sp:: calc_floor_area_retrofit(
   res_tb = building_data$residential,
   .existing_home_retrofit_pct = 0.80,
   .existing_home_ultra_retrofit_pct = 0.20,
   .enviro_factors = enviro_factors ) %>%
  dplyr::ungroup() %>%
  dplyr::filter(
    year == 2040,
    var %in% c(
       "multifamily_average_floor_area_sqft_county", 
       "multifamily_units",
       "single_family_units",
       "single_family_average_floor_area_sqft_ctu"
       )) %>%
  dplyr::filter(ctu_name == "Minneapolis") %>%
  kableformat(., .caption = "Existing Home Efficiency Minneapolis 2040")
```

### 3.2.1.d Behavior Change

Efficient messaging, in-home energy display, and smart meters together can reduce household energy use by `11%`. The default is `100%` of homes change behaviors due to these actions. 

Cities can adjust this participation rate accordingly.

##### Equation 4
$$ S_4sqft = X_1 \cdot population_{2040} \cdot S_3sqft/person_{2040}^{residential} \cdot 0.11 $$

##### Table 9

```{r tab9}
#| echo: false
ghg.sp::calc_floor_area_behavior_change(res_tb = building_data$residential,
   .home_behavior_change_pct = 1.00,
   .enviro_factors = enviro_factors) %>%
  dplyr::ungroup() %>%
  dplyr::filter(
    year == 2040,
    var %in% c(
       "multifamily_average_floor_area_sqft_county", 
       "multifamily_units",
       "single_family_units",
       "single_family_average_floor_area_sqft_ctu"
       )) %>%
  dplyr::filter(ctu_name == "Minneapolis") %>%
  kableformat(., .caption = "Behavior Change Minneapolis 2040")
```

### 3.2.2 Decarbonize Residential Energy Use

Due to decarbonizing the grid, the emission factor of electricity decreases from `568` kg<sub>2</sub>e/MWh (2018 Minnesota state level) to `113` kgCO<sub>2</sub>e/MWh. 

This change is fixed by the grid.

$$
S_5kgCO_2e = (population_{2040} \cdot S_4sqft/person_{2040}^{residential} \cdot kWh/sqft_{residential}^{2040}) \cdot [(kg CO_2e/MWh_{2040}^{electricity} \cdot 1 \times 10^{-3} kWh) \cdot X_1]
$$

```{r tab10}
#| echo: false
calc_ghg_residential(
  res_tb = building_energy_bau_data$residential,
  res_tb_bau = building_energy_bau_data$residential,
  .grid_decarbonization_pct = 0.8,
  .enviro_factors = enviro_factors
) %>% 
  dplyr::ungroup() %>%
  dplyr::filter(
    year == 2040) %>%
  dplyr::filter(ctu_name == "Minneapolis") %>%
  kableformat(., .caption = "Decarbonize Residential Energy Use Minneapolis 2040")
```

### 3.2.3 Electrifying Residential Heating

`59%` of homes are using air-source heat pumps and electric water heaters, instead of using gas for heating. 

Cities can adjust the percentage of homes participated in electrifying heating.

```{r tab11}
#| echo: false
ghg.sp::calc_electrify_residential_heating(
  res_tb = calc_ghg_residential(
  res_tb = building_energy_bau_data$residential,
  res_tb_bau = building_energy_bau_data$residential,
  .grid_decarbonization_pct = 0.8,
  .enviro_factors = enviro_factors
   ),
   .additional_electrified_residential_buildings_pct = 0.45,
   .res_natural_gas_for_space_heating_pct = 0.71,
   .res_natural_gas_for_water_heating_pct = 0.24,
   .grid_decarbonization_pct = 0.80,
   .enviro_factors = enviro_factors) %>%
  dplyr::ungroup() %>%
  dplyr::filter(
    year == 2040,
    ctu_name == "Minneapolis",
    scen == "scen") %>%
  kableformat(., .caption = "Electrify Residential Heating Minneapolis 2040")
```

### 3.3 Reducing Non Residential Energy Use

Non-residential energy use corresponds to commercial and industrial buildings. The estimate of non-residential energy use is based on the square footage of commercial and industrial buildings and employment forecasts. 

#### 3.3.1 Non Residential Energy Baseline

Table 6 shows the basic assumptions about non residential electricity and natural gas use for the baseline year of 2018 for the City of Minneapolis.

##### Table 12
```{r tab12}
#| echo = FALSE
ghg.sp::get_non_residential_energy_baseline(
  tb = building_energy_data) %>%
  dplyr::ungroup() %>%
  dplyr::filter(ctu_name == "Minneapolis") %>%
  kableformat(., .caption = "Non Residential Energy Baseline Minneapolis 2018")
```

#### 3.3.2 Non Residential Energy Forecast

`Table 7` shows the forecast assumptions for non residential electricity and natural gas use for the City of Minneapolis.

##### Table 13

```{r tab13}
#| echo = FALSE
ghg.sp::calc_non_residential_energy_forecast(
  tb = building_energy_data) %>%
  dplyr::ungroup() %>%
  dplyr::filter(ctu_name == "Minneapolis") %>%
  kableformat(., .caption = "Non Residential Energy Forecast Minneapolis 2040")
```

#### 3.3.3 Non Residential Energy Strategies

##### Table 14

```{r tab14}
#| echo = FALSE
#| message = FALSE
ghg.sp::calc_electrify_commercial_heating(
  non_res_tb = ghg.sp::calc_ghg_non_residential(
    non_res_tb = building_energy_bau_data$non_residential,
    non_res_tb_bau = building_energy_bau_data$non_residential,
    .industrial_smart_grid_pct = 1,
    .commercial_smart_grid_pct = 1,
    .grid_decarbonization_pct = 1,
    .smart_grid_energy_reduction_pct = 1,
    .enviro_factors = enviro_factors,
    .existing_high_efficiency_buildings_pct = 0.8
  ),
  .grid_decarbonization_pct = 0.8,
  .electrified_buildings_pct = 0.40,
  .non_res_natural_gas_for_water_heating_pct = 0.20,
  .non_res_natural_gas_for_space_heating_pct = 0.69,
  .enviro_factors = enviro_factors
) %>% 
  dplyr::ungroup() %>%
  dplyr::filter(ctu_name == "Minneapolis",
  year == 2040,
  scen == "scen") %>%
  kableformat(., .caption = "Electrify Commercial Heating Minneapolis 2040")
```


## 4. Reducing Emissions in the Transportation Sector

The inputs for the transportation sector baseliene and forecast are:

* Passenger Miles Traveled (PMT) forecasts from Metroplitan Council Regional Transportation Demand Model (RTDM)
* Average vehicle occupancy by mode
* MA3T stock turnover and adoption forecasts by powertrain: SI (spark ignition gasoline) vehicles, CI (combustion emissions diesel) vehicles, hybrid-electric vehicles (HEVs), plug-in hybrid-electric vehicles (PHEV), and battery-electric


### 4.1 Transportation Baseline

```{r}
ghg.sp::transportation_index$modes
```

#### 4.1.1 Passenger Transportation

##### Table 

```{r tabx}
#| echo: false
ghg.sp::transportation_data$passenger %>%
  dplyr::filter(ctu == "Minneapolis",
                var == "PMT",
                year == 2018) %>% 
  dplyr::left_join(., 
                ghg.sp::transportation_index$modes,
                by = c("mode" = "mode_abbrev")) %>%
  dplyr::select(mode_description_1, value) %>%
  dplyr::rename("var" = "mode_description_1") %>%
  dplyr::rename("Variables" = "var",
                  "Value" = "value") %>%
  kableExtra::kable(
      caption = "Passenger Miles Traveled by Mode Minneapolis 2018",
      format = "html",
      escape = FALSE,
      format.args = list(big.mark = ",", digits = 6)
    ) %>%
    kableExtra::kable_classic(html_font = "Arial",
                              full_width = F,
                              font_size = 12)
```

##### Table

#### 4.1.2 Freight Transportation

```{r}
ghg.sp::transportation_data$freight %>%
  dplyr::filter(ctu == "Minneapolis",
                var == "TMT",
                year == 2018) %>% 
  dplyr::left_join(., 
                ghg.sp::transportation_index$modes,
                by = c("mode" = "mode_abbrev")) %>%
  dplyr::select(mode_description_1, value)  %>%
  dplyr::rename("var" = "mode_description_1") %>%
  dplyr::rename("Variables" = "var",
                  "Value" = "value") %>%
  kableExtra::kable(
      caption = "Freight Miles Traveled by Mode Minneapolis 2018",
      format = "html",
      escape = FALSE,
      format.args = list(big.mark = ",", digits = 6)
    ) %>%
    kableExtra::kable_classic(html_font = "Arial",
                              full_width = F,
                              font_size = 12)
```

### 4.2 Transportation Forecast

#### 4.2.1 Passenger Transportation Forecast

##### Table
```{r tabx}
#| echo: false
ghg.sp::transportation_data$passenger %>%
  dplyr::filter(ctu == "Minneapolis",
                var == "PMT",
                year == 2040) %>% 
  dplyr::left_join(., 
                ghg.sp::transportation_index$modes,
                by = c("mode" = "mode_abbrev")) %>%
  dplyr::select(mode_description_1, value) %>%
  dplyr::rename("var" = "mode_description_1") %>%
  dplyr::rename("Variables" = "var",
                  "Value" = "value") %>%
  kableExtra::kable(
      caption = "Passenger Miles Traveled by Mode Minneapolis 2040 Forecast",
      format = "html",
      escape = FALSE,
      format.args = list(big.mark = ",", digits = 6)
    ) %>%
    kableExtra::kable_classic(html_font = "Arial",
                              full_width = F,
                              font_size = 12)
```

#### 4.3 Transportation Strategies

#### 4.3.1 Land Use Strategies

Increasing density decreases travel distances and encourages the use of transit, cycling, and walking. We estimate a `0.4%` decrease in auto VMT for a `10%` increase in population density by 2040, a `7%` increase in walking/cycling VMT for a `10%` increase in population density by 2040, and a `7%` increase in transit VMT for a `10%` increase in population density by 2040.



```{r}

```



## 5. Reducing Emission from Land Use and Forestry

The Metropolitan Council has developed guides for each of the strategies listed at right. Links will be added when the guides are made avail

### 5.1 Land Use Baseline

```{r}
development_types <- ghg.sp::calc_land_by_development_type(tb = land_use_data,
   .urban_form_scenario = "bau") %>%
  dplyr::filter(ctu_name == "Minneapolis") 
```


