---
title: "transportation_scenario"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{transportation_scenario}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ghg.sp)
library(tidyverse)
ggplot2::theme_set( councilR::council_theme(use_showtext = TRUE))
st_paul_pass <- transportation_data$passenger %>% 
                                  filter(ctu %in% c("St. Paul",
                                                    "All"))
st_paul_freight <- transportation_data$freight %>% 
                                  filter(ctu %in% c("St. Paul",
                                                    "All"))
```

## Business as usual

```{r}
bau_summary <- scenario_results(pass_tb = st_paul_pass,
                                freight_tb = st_paul_freight,
                                .scenario = "BAU",
                                .electric_scenario = "ER",
                                .aeo_scenario = "REF")

bau_pldv <-  bau_summary$passenger_all %>% filter(mode == "PLDV") 

```

## Interventions  

### Land use  

```{r}
mitigation_lu <-  scenario_results(
  pass_tb = st_paul_pass,
  freight_tb = st_paul_freight,
  .scenario = "MIT-LU",
  .electric_scenario = "ER",
  .aeo_scenario = "REF", 
  .pop_dens_pct_change = 0.05,
  .emp_dens_pct_change = 0.05,
  .land_use_diversity_pct_change = 0.05,
  .intersection_design_pct_change = 0.05,
  .job_access_pct_change = 0.05,
  .transit_dist_pct_change = -0.05,
  .comb_5d_impact_pct_change = 0.02
) %>% 
  suppressMessages()
```


### Vehicle electrification  

- Increase private EV sales to 20% HEV, 20% PHEV, and 60% BEV of all vehicle sales by 2050. 
- Phase out biodiesel buses from fleet by 2035 and HEV bus sales by 2050.
- Introduce dynamic ride-sharing (DRS) in 2025 to replace 90% of motorized PMT by 2050. 
- Introduce autonomous vehicles in 2025 to replace 10% of motorized PMT by 2050.


```{r, message=FALSE}
mitigation_ev <-  scenario_results(
  pass_tb = st_paul_pass,
  freight_tb = st_paul_freight,
  .scenario = "MIT-EV",
  .electric_scenario = "ER",
  .aeo_scenario = "REF", 
  .bev_pct_sales = 0.6,
  .hev_pct_sales = 0.175,
  .phev_pct_sales = 0.175
  # .hev_pct_sales =  0.2,
  # .phev_pct_sales = 0.2
  # .av_pct = 0.05,
  # .av_fuel_type = "BEV"
) %>%
  suppressMessages()


mit_ev_pldv <-  mitigation_ev$passenger_all %>% filter(mode == "PLDV") 

```

### Transit improvements   

- Increase average vehicle occupancy by 10% by 2050
- Increase transit ridership by 10% by 2050
- Increase daily fraction of telework by 10% by 2050.

```{r}
mitigation_trans <-  scenario_results(
  pass_tb = st_paul_pass,
  freight_tb = st_paul_freight,
  .scenario = "MIT-TRANSIT",
  .electric_scenario = "ER",
  .aeo_scenario = "REF", 
  .transit_avo = 0.47 * 1.1,
  .transit_rider_pct = 0.10,
  .telework_pct = 0.10
) %>% 
  suppressMessages()
```


### Economic instruments  

- Introduce $0.05 per mile VMT fee in 2025
- Introduce a $0.10 per mile congestion fee in 2025
- Introduce a $0.05 per mile pay-as-you-drive insurance fee in 2025
- Double the gas tax (currently $0.47 per gallon)
- Double parking price (currently $0.22 per hour)


```{r}
mitigation_econ <-  scenario_results(
  pass_tb = st_paul_pass,
  freight_tb = st_paul_freight,
  .scenario = "MIT-ECON",
  .electric_scenario = "ER",
  .aeo_scenario = "REF", 
  .vmt_fee = 0.05,
  .cong_price = 0.10,
  .payd_fee = 0.05,
  .gas_tax = 0.47 * 2,
  .parking_price = 0.22 * 2
) %>% 
suppressMessages()
```


## Compare  

```{r}
all_scen_passenger_av <- purrr::map_dfr(
  list(bau_summary,
       mitigation_econ,
       mitigation_ev,
       mitigation_trans,
       mitigation_lu),
  function(x){
    x$passenger_all %>% 
      filter(mode %in% c("AV", "PLDV")) %>% 
      group_by(year, scenario) %>% 
      summarize(dir_ghg = sum(dir_ghg, na.rm = T))
  }
)

ggplot(all_scen_passenger_av,
       aes(x = year,
           y = dir_ghg,
           color = scenario,
           group = scenario)) +
  geom_point() +
  geom_line() +
  labs(title = "All Passenger Vehicle Emissions")
```

```{r}
all_scen_passenger_vmt <- purrr::map_dfr(
  list(bau_summary,
       mitigation_econ,
       mitigation_ev,
       mitigation_trans,
       mitigation_lu),
  function(x){
    x$passenger_all %>% 
      filter(mode %in% c("PLDV")) %>% 
      group_by(year, scenario) %>% 
      summarize(vmt = sum(vmt, na.rm = T))
  }
)

ggplot(all_scen_passenger_vmt,
       aes(x = year,
           y = vmt,
           color = scenario,
           group = scenario)) +
  # facet_wrap(~class) +
  geom_point() +
  geom_line(alpha = 0.5,
            size = 1.1) +
  labs(title = "All Passenger Vehicle Miles Traveled")
```




```{r}
pass_bau <- bau_summary$passenger$BU_BRT$dir_ghg %>% 
  group_by(year, scenario) %>% 
  summarize(dir_ghg = sum(dir_ghg, na.rm = T))


pass_ev <- mitigation_ev$passenger$BU_BRT$dir_ghg %>% 
  group_by(year, scenario) %>% 
  summarize(dir_ghg = sum(dir_ghg))

ggplot(bind_rows(pass_ev, pass_bau),
       aes(x = year,
           y = dir_ghg,
           color = scenario,
           group = scenario)) +
  geom_point() +
  geom_line() +
  labs(title = "All Bus and BRT Emissions")
```



