---
title: "transportation_scenario"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{transportation_scenario}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(ghg.sp)
library(ggplot2)
library(dplyr)
library(stringr)
library(scales)
```

```{r setup}

ggplot2::theme_set(councilR::council_theme(use_showtext = TRUE))
st_paul_pass <- transportation_data$passenger %>% 
  filter(ctu %in% c("St. Paul",
                    "All"),
         !year %in% c("2045",
                      "2050"))
st_paul_freight <- transportation_data$freight %>% 
  filter(ctu %in% c("St. Paul",
                    "All"),
         !year %in% c("2045",
                      "2050"))
```

## Business as usual

```{r}
bau_summary <- run_scenario(pass_tb = st_paul_pass,
                            freight_tb = st_paul_freight,
                            .scenario = "BAU",
                            .electric_scenario = "ER",
                            .aeo_scenario = "REF")

bau_pldv <-  bau_summary$passenger_all %>% filter((mode %in% c("AV", "PLDV",
                                                               "DRS"))) 

```

## Interventions  

### Land use  

```{r}
mitigation_lu <-  run_scenario(
  pass_tb = st_paul_pass,
  freight_tb = st_paul_freight,
  .scenario = "strategy_land_use",
  .electric_scenario = "ER",
  .aeo_scenario = "REF", 
  .pop_dens_pct_change = 0.05,
  .emp_dens_pct_change = 0.05,
  .land_use_diversity_pct_change = 0.05,
  .intersection_design_pct_change = 0.05,
  .job_access_pct_change = 0.05,
  .transit_dist_pct_change = -0.05,
  .comb_5d_impact_pct_change = 0.02
) %>% 
  suppressMessages()
```

### Transit improvements   

- Increase average vehicle occupancy by 10% by 2040
- Increase transit ridership by 10% by 2040
- Phase out biodiesel fuel buses from fleet by 2035 and hybrid electric bus sales by 2040. Fleet will be entirely electric  


```{r}
mitigation_trans <-  run_scenario(
  pass_tb = st_paul_pass,
  freight_tb = st_paul_freight,
  .scenario = "strategy_improve_transit",
  .electric_scenario = "ER",
  .aeo_scenario = "REF", 
  .transit_avo = 0.10,
  .transit_rider_pct = 0.10,
) %>% 
  suppressMessages()
```



### Vehicle electrification  

- Increase private EV sales to 20% HEV, 20% PHEV, and 60% BEV of all vehicle sales by 2040. 
- Phase out biodiesel buses from fleet by 2035 and HEV bus sales by 2040.
- Introduce dynamic ride-sharing (DRS) in 2025 to replace 10% of motorized PMT by 2040. 
- Introduce autonomous vehicles in 2025 to replace 10% of motorized PMT by 2040.


```{r, message=FALSE}
mitigation_ev <-  run_scenario(
  pass_tb = st_paul_pass,
  freight_tb = st_paul_freight,
  .scenario = "strategy_veh_electric",
  .electric_scenario = "ER",
  .aeo_scenario = "REF", 
  .bev_pct_sales = 0.6,
  .hev_pct_sales =  0.2,
  .phev_pct_sales = 0.2,
  .av_pct = 0.10,
  .av_fuel_type = "HEV",
  .drs_pct = 0.10,
  .drs_fuel_type = "HEV"
) %>%
  suppressMessages()


mit_ev_pldv <-  mitigation_ev$passenger_all %>%
  filter(mode %in% c("AV", "PLDV", "DRS")) 

```

### Road price  

```{r}
mitigation_road_price<-  run_scenario(
  pass_tb = st_paul_pass,
  freight_tb = st_paul_freight,
  .scenario = "strategy_road_price",
  .electric_scenario = "ER",
  .aeo_scenario = "REF", 
  .parking_price = 44,
  .freight_parking_price = 2,
  .vmt_fee = 0.05
  
)  %>% 
  suppressMessages()

```


### Telework  

```{r}
mitigation_telework <-  run_scenario(
  pass_tb = st_paul_pass,
  freight_tb = st_paul_freight,
  .scenario = "strategy_telework",
  .electric_scenario = "ER",
  .aeo_scenario = "REF", 
  .telework_pct = 0.10
)  %>% 
  suppressMessages()
```



## Compare  


### Passenger  

```{r}
all_scen_passenger_av <- purrr::map_dfr(
  list(bau_summary,
       mitigation_ev,
       mitigation_trans,
       mitigation_lu,
       mitigation_telework,
       mitigation_road_price
  ),
  function(x){
    x$passenger_all %>% 
      filter(mode %in% c("AV", "PLDV", "DRS" ,"BU", "BRT")) %>%
      group_by(year, scenario) %>% 
      summarize(dir_ghg = sum(dir_ghg, na.rm = T))
  }
)

ggplot(all_scen_passenger_av,
       aes(x = year,
           y = dir_ghg,
           color = scenario,
           group = scenario)) +
  geom_point() +
  geom_line(alpha = 0.5,
            size = 1) +
  labs(title = "All Passenger Vehicle Emissions")
```

```{r}
all_scen_passenger_vmt <- purrr::map_dfr(
  list(bau_summary,
       mitigation_ev,
       mitigation_trans,
       mitigation_lu,
       mitigation_telework,
       mitigation_road_price
  ),
  function(x){
    x$passenger_all %>% 
      filter(mode %in% c("PLDV", "AV", "DRS")) %>% 
      group_by(year, scenario) %>% 
      summarize(vmt = sum(vmt, na.rm = T))
  }
)

ggplot(all_scen_passenger_vmt,
       aes(x = year,
           y = vmt/1000000,
           color = scenario,
           group = scenario)) +
  # facet_wrap(~class) +
  geom_point() +
  geom_line(alpha = 0.5,
            size = 1.1) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "All Passenger Vehicle Miles Traveled")
```


### Freight  


```{r}
all_scen_freight_em <- purrr::map_dfr(
  list(bau_summary,
       mitigation_ev,
       mitigation_trans,
       mitigation_lu,
       mitigation_telework,
       mitigation_road_price
  ),
  function(x){
    x$freight_all %>% 
      group_by(year, scenario) %>% 
      summarize(dir_ghg = sum(dir_ghg, na.rm = T))
  }
)

ggplot(all_scen_freight_em,
       aes(x = year,
           y = dir_ghg,
           color = scenario,
           group = scenario)) +
  # facet_wrap(~class) +
  geom_point() +
  geom_line(alpha = 0.5,
            size = 1.1) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "All Freight Emissions")
```
## App data output   

```{r}
app_out <- purrr::map_dfr(
  list(bau_summary,
       mitigation_ev,
       mitigation_trans,
       mitigation_lu,
       mitigation_telework,
       mitigation_road_price
  ),
  function(x){
    x$app_data
  }
) %>% 
  ungroup() %>% 
  rowwise() %>% 
  mutate(mode = case_when(
    mode %in% c("PLDV", "AV", "DRS") ~ "personal vehicles",
    mode %in% c("WALK", "BIKE") ~ "walk and bike",
    mode %in% c("BU", "BRT", "BS", "RU", "RI") ~ "public transit",
    mode %in% c("CUT", "SUT") ~ "trucks",
    mode %in% c("MM", "AIR", "WAT", "FR") ~ "other freight"),
    submodule = paste0(submodule, " - ", mode)) %>% 
  group_by(ctu, year, scenario, module, submodule, name) %>% 
  summarize(value = sum(value, na.rm = T), .groups = "keep") %>% 
  ungroup() %>% 
    mutate(type = ifelse(scenario == "BAU", "bau_forecast",
                       scenario),
         ctu = "Hollywood",
         ctu_id = "99999",
         metric = "emissions_tonnes_co2e") %>% 
  select(ctu_id, ctu, year, module,
         submodule, type, source = name, metric, value)

write.csv(app_out, "../vignettes/app_sample.csv", row.names = FALSE)
```

