### Urban Form

In the coming decades, the Twin Cities region is expected to experience significant growth, with its population projected to increase from just under 3.2 million residents to over 3.7 million by 2040. This growth will bring about substantial changes in land use and land cover, particularly in the north and west suburbs. 

**What are Biogenic Emissions?** 

*Biogenic emissions* arise from natural processes, predominantly the decomposition of organic matter. Urban expansion can have a significant
impact on these emissions, as the expansion itself can release long-stored carbon, and the presence of vegetation and soil can sequester carbon dioxide. However, the duration of carbon sequestration in biomass and the level at which it becomes saturated remain unclear.

To assess the impact of urban vegetation on carbon stock changes and sequestration, the method outlined in [@Milnar_Ramaswami_2020] was used
to calculate these values for the study period of 2018-2040.

::: about_strategy
***About this strategy*** Our assumption is that cities can control (or at least influence) the urban expansion of their city into forested or green areas to minimize loss of carbon stock and carbon sequestration. This reference scenario shows the expected carbon sequestration and carbon stock values under the business-as-usual scenario versus  a more compact development scenario.  
:::

**GHG Impact Calculation**

$$CO_2sqlc = (\text{Sequestration} + \text{Land Cover Change}) \times 44/12$$
*Where:*

$\text{Sequestration} = \sum{(Ai \times SRi)}$\
$\text{Land Cover Change} = \sum{(Ait-1 × CDi ) - (Ait \times CDi )]}$

*And:*

- $\text{Sequestration} = \text{annual net sequestration from vegetation for a given city}$
- $\text{Land Cover Change = annual organic carbon stock change from land conversion for a given city, averaged over the study period}$
- $44/12 = \text{factor to convert Mg C to Mg } CO_2e$
- $Ai = \text{Area of vegetation of cover type i (ha)}$
- $SRi = \text{Annual net sequestration rate for cover type i. (Mg C ha-1 yr-1)}$
- $CDi = \text{Carbon stock density factor for vegetation in cover type i (Mg C ha-1)}$
- $i = 1 ,2, 3, ...n$ $\text{land cover types}$
- $t = \text{year for which biogenic emissions were calculated and compared to Scope 1+2 emissions for each city}$

```{r include=FALSE}
urban_form_bau <- calc_scen_land_use(
  .selected_ctu = params$ctu_selection,
  tb = land_use_data,
  .urban_form_scenario = "bau"
) %>%
  filter(ctu_name == params$ctu_selection) %>%
  group_by(development_type) %>%
  summarise(scenario_hectares = sum(scenario_hectares))


urban_form_scen <- calc_scen_land_use(
  .selected_ctu = params$ctu_selection,
  tb = land_use_data,
  .urban_form_scenario = "compact_dev_beyond_bau"
) %>%
  filter(ctu_name == params$ctu_selection) %>%
  group_by(development_type) %>%
  summarise(scenario_hectares = sum(scenario_hectares))
```


```{r fig05_2, include=FALSE}
fig05_1_bau_2018 <- luse_bau_2018 %>%
  ungroup() %>%
  rename(var = detail) %>%
  left_join(.,
            variables,
            by = "var")  %>%
  select(desc, value) %>%
  rename(Variables = desc,
         "2018 Baseline" = value)

fig05_1_bau_2040 <- luse_bau_2040 %>%
  ungroup() %>%
  rename(var = detail) %>%
  left_join(.,
            variables,
            by = "var")  %>%
  select(desc, value) %>%
  rename(Variables = desc,
         "2040 BAU" = value)

fig05_1_scen_2040 <-
  run_scenario_land_use(
    tb = land_use_data,
    .selected_ctu = params$ctu_selection,
    .urban_form_scenario = "compact_dev_beyond_bau",
    .conservation_tillage_intervention = "current_conservation_tillage",
    .tree_planting_intervention = "match_la_million_trees_goal",
    .tree_planting_per_capita = 0.26,
    .tree_planting_per_hectare = 247,
    .parking_lot_reduction_percentage = 0,
    detail = FALSE
  )  %>%
  filter(year == 2040)  %>%
  ungroup() %>%
  rename(var = detail) %>%
  left_join(.,
            variables,
            by = "var")  %>%
  select(desc, value) %>%
  rename(Variables = desc,
         "2040 Scenario" = value)

fig05_1_data <-
  full_join(
  full_join(
    fig05_1_bau_2018,
    fig05_1_bau_2040,
    by = "Variables"
  ),
  fig05_1_scen_2040,
  by = "Variables")

fig05_1_table <- fig05_1_data %>%
  tableformat(.caption = paste0("Land Use for ", params$ctu_selection))
```

```{r}
fig05_1_table
```

For the City of Minneapolis, there does not appear to be a significant difference in carbon sequestration and stocks between the two scenarios presented. 
