---
title: "City/township and county instersection"
author: Ellen
output: rmarkdown::html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = F
)

library(knitr)
library(councilR)
library(dplyr)
library(ggplot2)
library(tidyr)
library(sf)

`%not_in%` <- Negate(`%in%`)
st_erase <- function(x, y) st_difference(x, st_union(st_combine(y)))
```


# Introduction

The goal is to create a table which contains each CTU and county pair for the 7-county metro region. Each ctu-county pair will have information about the percent of the population (2020 census) and percent of land area (water is excluded) contained in each geography. 

1. Use 2020 census to get population percents for each ctu-county pair. 

```{r population}
temp <- tempfile()
download.file("https://resources.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_metc/society_census2020population/xlsx_society_census2020population.zip", destfile = temp)
pop <- readxl::read_xlsx(unzip(temp, "Census2020PopulationMCD.xlsx")) %>%
  filter(
    TCFLAG == 1,
    GEOG_LEVEL == "COCTU"
  ) %>%
  mutate(
    GEONAME = stringr::str_replace(GEONAME, "township", "Twp."),
    GEONAME = stringr::str_remove(GEONAME, " city")
  ) %>%
  transmute(
    ctu_name = GEONAME,
    co_code = COUNTY,
    pop = POPTOTAL
  ) %>%
  group_by(ctu_name) %>%
  mutate(
    n_counties = n(),
    total_pop = sum(pop),
    pct_population = round(pop / total_pop, 4)
  ) %>%
  mutate(ctu_name = case_when(
    ctu_name == "Credit River Twp." ~ "Credit River",
    ctu_name == "Fort Snelling UT" ~ "Fort Snelling (unorg.)",
    TRUE ~ ctu_name
  )) %>%
  select(-total_pop, -pop) %>%
  arrange((ctu_name))

fs::file_delete("Census2020PopulationMCD.xlsx")
```

2.  Get CTU boundary files and remove water. 

```{r}
## get water file
temp <- tempfile()
download.file("ftp://ftp.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_metc/water_lakes_rivers/gpkg_water_lakes_rivers.zip", destfile = temp)

river_lake_all <- sf::read_sf(unzip(temp, "water_lakes_rivers.gpkg")) %>%
  sf::st_transform(26915)

river_lake_buffer <- river_lake_all %>%
  filter(NAME_DNR %in% c("Mississippi", "Minnesota", "St. Croix")) # %>% #these rivers are boundaries

fs::file_delete("water_lakes_rivers.gpkg")

# get ctu boundary, remove water
temp <- tempfile()
temp2 <- tempfile()
download.file(
  "https://resources.gisdata.mn.gov/pub/gdrs/data/pub/us_mn_state_metc/bdry_metro_counties_and_ctus/shp_bdry_metro_counties_and_ctus.zip",
  destfile = temp
)
unzip(zipfile = temp, exdir = temp2) # list.files(temp2)
ctu_boundary <- sf::read_sf(paste0(temp2, pattern = "/CTUs.shp")) %>%
  transmute(
    ctu_id = CTU_ID,
    ctu_name = CTU_NAME
  ) %>%
  sf::st_transform(26915) %>%
  st_erase(river_lake_all %>%
    st_union() %>%
    st_buffer(0))
```

3. Intersect CTU boundaries (less water) with county boundaries. Get the percent LAND AREA of each ctu-county pair.

```{r}
## get county boundaries
metc_counties <- tigris::counties(state = "MN") %>%
  filter(NAME %in% c("Anoka", "Carver", "Dakota", "Hennepin", "Ramsey", "Scott", "Washington")) %>%
  transmute(
    co_code = COUNTYFP,
    co_name = NAME
  ) %>%
  sf::st_transform(26915)

## intersection
land_area <- metc_counties %>%
  sf::st_intersection(ctu_boundary) %>%
  mutate(sq_m = as.numeric(sf::st_area(.))) %>%
  st_drop_geometry() %>%
  group_by(ctu_id, ctu_name) %>%
  mutate(
    n_counties = n(),
    ctu_land_area = sum(sq_m),
    pct_land_area = round(sq_m / ctu_land_area, 4)
  ) %>%
  select(co_code, co_name, ctu_id, pct_land_area)
```

4. Merge population and land area percents. Export file. 

```{r}
ctu_county <- st_drop_geometry(ctu_boundary) %>% # take the full ctu information (with ctu_ids)
  full_join(pop) %>% # add population percents
  left_join(land_area) %>% # add in the percent land area
  mutate(pct_land_area = if_else(n_counties == 1, 1, pct_land_area)) %>% # clean up errors due to slivers
  select(-ctu_name, -co_name) %>%
  mutate(
    ctu_id = as.integer(ctu_id),
    co_code = as.integer(co_code)
  )


ctu_county %>%
  group_by(ctu_id) %>%
  mutate(
    check_land_area = sum(pct_land_area),
    check_pop = sum(pct_population)
  ) %>%
  filter(check_pop != 1 | check_land_area != 1)

usethis::use_data(ctu_county, overwrite = TRUE)
```

