#### Smart Grid Electrification

Smart grid and technologies are projected to reduce industrial electricity use by 11% .

::: about_strategy
***About this strategy*** Our assumption is that by adopting technologies such as smart grid, `100%` of industrial buildings in the city can reduce their electricity use by `11%`.
:::

**GHG Impact Calculation**

$$ \text{electricity } kgCO2e_{2040}^{industrial} = x * jobs^{industrial}_{2040} * MWh/jobs^{industrial}_{2040} * 11\% * kgCO_2e/MWh$$
*Where:*

$x$ = % of industries in the smart grid (assumed to be 100%)
$jobs^{industrial}_{2040}$ = Number of industrial workers in 2040
$MWh/jobs^{industrial}_{2040}$ = Projected industrial electricity use per worker, MWh/worker, for 2040
$kgCO2e/MWh$ = Electricity emission factor in 2040 in kilograms of CO2e/MWh
$11%$ = % reduction in energy use due to smart grid

The total reduction of greenhouse gas emissions in 2040 (E) can be calculated as:

This equation represents the expected reduction of greenhouse gas emissions in the industrial sector due to the implementation of a smart grid in 2040.

```{r fig03b2, include=FALSE}
# 2018
fig03b2_data <- full_join(non_res_bau_2018,
                          non_res_bau_2040,
                          by = "Variables") %>% full_join(
                            .,
                            # 2040 Scenario
                            run_scenario_building(
                              run_residential = FALSE,
                              run_non_residential = TRUE,
                              .selected_ctu = params$ctu_selection,
                              non_res_tb = building_data$non_residential,
                              non_res_tb_bau = building_data$non_residential,
                              .electrified_buildings_pct = .5,
                              .non_res_natural_gas_for_water_heating_pct = 0.20,
                              .non_res_natural_gas_for_space_heating_pct = 0.69,
                              .commercial_smart_grid_pct = 1,
                              .industrial_smart_grid_pct = 1,
                              .smart_grid_energy_reduction_pct = 1,
                              .grid_decarbonization_pct = 0,
                              .existing_high_efficiency_buildings_pct = 0,
                              .enviro_factors = enviro_factors
                            ) %>%
                              ungroup() %>%
                              dplyr::filter(
                                year == 2040,
                                scen == "scen",
                                var %in% c(
                                  "commercial_mwh",
                                  "industrial_mwh",
                                  "commercial_therms",
                                  "industrial_therms",
                                  "total_industrial_commercial_emissions"
                                )
                              ) %>%
                              add_variable_names(., "2040 Scenario"),
                            by = "Variables"
                          )

fig03b2_table <- fig03b2_data %>%
  # add table format
  tableformat(.,
              .caption = paste("Retrofit Commercial Buildings",
                               params$ctu_selection,
                               "2040"))
```

```{r show_fig03b2, echo=FALSE}
fig03b2_table
```

```{r echo=FALSE}
chartformat(fig03b2_data)
```

