#### Smart Grid Electrification

<!-- Smart grid and technologies are projected to reduce industrial electricity use by 11% . -->
Achieving a net-zero grid may require transitioning to a smart grid. Smart grid technology enhances communication between an energy provider and energy users so the grid can adapt as demand changes ([Smart Grid.Gov](https://www.smartgrid.gov/the_smart_grid/smart_grid.html)). The smart grid can not only manage the load but also reduce energy use by approximately 11% as shown in PNNL's "[The Smart Grid: An Estimation of the Energy and CO2 Benefits](https://www.osti.gov/biblio/971445)." This reduction will be applied to all electricity consumers connected to the grid. 

::: about_strategy
***About this strategy*** We model the assumption that by adopting technologies such as smart grid, `100%` of industrial buildings in the city can reduce their electricity use by `11%`.
:::

<!-- **GHG Impact Calculation for Smart Grid** -->

<!-- $$ \text{new electricity } kgCO_2e_{2040}^{industrial} = x * \left(jobs^{industrial}_{2040} * \frac{MWh}{jobs^{industrial}_{2040}} * 11\% \right) * \frac{kgCO_2e}{MWh}$$ -->
<!-- *Where:* -->

<!-- -  $x\text{, where } 0 \le x \le 1$: The percentage of industries in the smart grid (assumed to be 100%) -->
<!-- -  $jobs^{industrial}_{2040}$: The projected number of industrial workers in 2040 -->
<!-- -  $MWh/jobs^{industrial}_{2040}$ = The projected industrial electricity use per worker in 2040 -->
<!-- -  $11%$ = The percentage reduction in energy use due to smart grid, fixed at 11% -->
<!-- -  $kgCO2e/MWh$ = The projected electricity emission factor in 2040 -->

```{r fig03b2, include=FALSE}
# 2018
fig03b2_data <- full_join(non_res_bau_2018,
  non_res_bau_2040,
  by = "Variables"
) %>% full_join(
  .,
  # 2040 Scenario
  run_scenario_building(
    run_residential = FALSE,
    run_non_residential = TRUE,
    .selected_ctu = params$ctu_selection,
    non_res_tb = building_data$non_residential,
    non_res_tb_bau = building_data$non_residential,
    .electrified_buildings_pct = .5,
    .non_res_natural_gas_for_water_heating_pct = 0.20,
    .non_res_natural_gas_for_space_heating_pct = 0.69,
    .commercial_smart_grid_pct = 1,
    .industrial_smart_grid_pct = 1,
    .smart_grid_energy_reduction_pct = 1,
    .grid_decarbonization_pct = 0,
    .existing_high_efficiency_buildings_pct = 0,
    .enviro_factors = enviro_factors
  ) %>%
    ungroup() %>%
    dplyr::filter(
      year == 2040,
      scen == "scen",
      var %in% c(
        "commercial_mwh",
        "industrial_mwh",
        "commercial_therms",
        "industrial_therms",
        "total_industrial_commercial_emissions"
      )
    ) %>%
    add_variable_names(., "2040 Scenario"),
  by = "Variables"
)

fig03b2_table <- fig03b2_data %>%
  # add table format
  tableformat(.,
    .caption = paste(
      "Smart Grid Efficiency",
      params$ctu_selection,
      "2040"
    )
  )
```

```{r show_fig03b2, echo=FALSE}
fig03b2_table
```

```{r echo=FALSE}
chartformat(fig03b2_data)
```
