#### Decarbonize the Electricity Grid

This strategy demonstrates the impact of grid decarbonization.

::: about_strategy
***About this strategy*** Our assumption is that cities can control (or at least influence)... 
:::

**GHG Impact Calculation**

$$\text{new electricity }kgCO2e_{2040}^{nonresidential} = (jobs_{2040}^{commercial} * MWh_{2040}^{commercial} + jobs_{2040}^{industrial} * MWh_{2040}^{industrial}) * \left(\frac{kgCO_2e/MWh_{2040}}{1000}\right) * x$$
*Where:*

-  $x\text{, where } 0 \le x \le 1$: = The percentage reduction in grid emission factor from 2018 (assumed to be 80%)
-  $jobs_{2040}^{commercial}$: The projected number of commercial jobs in 2040
-  $MWh/jobs_{2040}^{commercial}$: The projected commercial electricity use per worker in 2040
-  $jobs_{2040}^{industrial}$: The projected number of industrial jobs in 2040
-  $MWh/jobs_{2040}^{industrial}$: The projected number industrial electricity use per worker in 2040
-  $kgCO_2e/MWh_{2040}$: The projected electricity emission factor in 2040

Note: The values are multiplied by 1000 in the original formula to retain the units in kilowatt hours.

```{r fig03b4, include=FALSE}
fig03b4_data <- full_join(non_res_bau_2018,
                          non_res_bau_2040,
                          by = "Variables") %>% full_join(
                             .,
                             run_scenario_building(
                                run_residential = FALSE,
                                run_non_residential = TRUE,
                                .selected_ctu = params$ctu_selection,
                                non_res_tb = building_data$non_residential,
                                non_res_tb_bau = building_data$non_residential,
                                .renewable_ng_nonres = FALSE,
                                .electrified_buildings_pct = 0,
                                .non_res_natural_gas_for_water_heating_pct = 0.20,
                                .non_res_natural_gas_for_space_heating_pct = 0.69,
                                .commercial_smart_grid_pct = 1,
                                .industrial_smart_grid_pct = 1,
                                .smart_grid_energy_reduction_pct = 1,
                                .grid_decarbonization_pct = 0,
                                .existing_high_efficiency_buildings_pct = 0,
                                .enviro_factors = enviro_factors
                             ) %>%
                                ungroup() %>%
                                dplyr::filter(
                                   year == 2040,
                                   scen == "scen",
                                   var %in% c(
                                      "commercial_mwh",
                                      "industrial_mwh",
                                      "commercial_therms",
                                      "industrial_therms",
                                      "total_industrial_commercial_emissions"
                                   )
                                ) %>%
                                add_variable_names(., "2040 Scenario"),
                             by = "Variables"
                          )

fig03b4_table <- fig03b4_data %>%
  # add table format
  tableformat(.,
              .caption = paste("Decarbonise Grid",
                               params$ctu_selection,
                               "2040"))
```

```{r show_fig03b4, echo=FALSE}
fig03b4_table
```

```{r echo=FALSE}
chartformat(fig03b4_data)
```

