### Land Use
```{r}
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F
)
```

Transportation and land use policies rely on each other. Single-use zoning and low-density land use patterns lead to a reliance on automobiles, increased vehicle miles traveled, and decreased efficacy of shared travel. 

However, altering land use policy is complex and only sometimes leads to changes in land use, as construction takes time to catch up with zoning. The scenario planning model uses the "five D's" (population and employment density, diversity of land uses, street design, distance to transit, and destination accessibility) as the land use factors that affect transportation.

The Metropolitan Council [Transportation Policy Plan](https://metrocouncil.org/Transportation/Planning-2/Key-Transportation-Planning-Documents/Transportation-Policy-Plan.aspx) discusses the need for communities to consider the connections between land use and transportation as the region changes in the coming years. The plan mentions that it is not feasible to expand the highway system in a sustainable manner, that there is increased private investment in already developed parts of the region along transit lines, and that higher-density development and redevelopment occur in already-developed areas. That new growth is happening in suburban edge and emerging suburban edge communities, which stresses the regional highway system. The plan suggests that planners incorporate multimodal travel options, such as transit, walking, and biking, into land use and design to address issues like congestion, climate change, equity, and livability.

<!-- TODO equations, citations, more  -->

::: about_strategy
***About this strategy*** Increasing density decreases travel distances and encourages transit, cycling, and walking. For this example, we estimate a `0.4%` decrease in auto vehicle miles traveled (VMT), a `7%` increase in walking/cycling VMT, and a `7%` increase in transit VMT based on a `10%` increase in population density by 2040.
:::

```{r tab-land-use-effects}
#| echo: false
#| message: false
ghg.sp::elast_5d %>%
  dplyr::filter(year == "2050") %>%
  mutate(type = stringr::str_to_title(type)) %>%
  select(
    `Mode` = type,
    `Population density` = population_density,
    `Employment density` = employment_density,
    `Land use diversity` = diversity,
    `Intersection design` = design,
    `Job accessability by transit` = job_access,
    `Distance to transit` = distance
  ) %>%
  tidyr::pivot_longer(
    names_to = "Land Use Adjustment",
    values_to = "Effect on miles traveled by 2050",
    cols = 2:7
  ) %>%
  mutate(
    `Multiplier relative to BAU` =
      ifelse(`Land Use Adjustment` == "Distance to transit", -0.1, 0.1)
  ) %>%
  select(
    Mode, `Land Use Adjustment`, `Multiplier relative to BAU`,
    `Effect on miles traveled by 2050`
  ) %>%
  rowwise() %>%
  mutate(across(c(3, 4), function(x) {
    ifelse(x != "NA",
           paste0(ifelse(x >= 0, "+", ""), x * 100, "%"),
           x
    )
  })) %>%
  gt::gt(
    id = "land-use-effects",
    caption = "Land use adjustment values",
    rowname_col = "`Land Use Adjustment`",
    groupname_col = "Mode"
  ) %>%
  gt::opt_row_striping() %>%
  gt::tab_options(row_group.font.weight = "bold")
```

<!-- #### Elasticities  -->


<!-- Land use elasticities are taken from a meta-analysis by Ewing and Cervero [@ewingTravelBuiltEnvironment2010] for each of automobile, walking and cycling, and transit trips. A more recent meta-analysis provides additional values for the automobile mode [@stevensDoesCompactDevelopment2016]. These values are used to give an elasticity range. In addition to the below values, an overall elasticity of -0.22 is used for the combined effect of the 5Ds to give the user the option to consider a combination of land use changes. -->

<!-- ```{r tab-land-use-elast} -->
<!-- #| echo: false -->
<!-- #| warning: false -->
<!-- tibble::tibble( -->
<!--   Mode = c( -->
<!--     "Automobile", -->
<!--     "Walking and cycling", -->
<!--     "Transit" -->
<!--   ), -->
<!--   `Population elasticity` = c( -->
<!--     "-0.04 (-0.04 to -0.22)", -->
<!--     "0.07", -->
<!--     "0.07" -->
<!--   ), -->
<!--   `Employment elasticity` = c( -->
<!--     "0 (0.00 to -0.07)", -->
<!--     "0.04", -->
<!--     "0.01" -->
<!--   ), -->
<!--   `Diversity (Land use entropy)` = c( -->
<!--     "-0.09 (-0.02 to -0.10)", -->
<!--     "0.15", -->
<!--     "0.12" -->
<!--   ), -->
<!--   `Design (% 4-way intersections` = c( -->
<!--     "-0.12 (-0.06 to-0.12)", -->
<!--     "-0.06", -->
<!--     "0.12" -->
<!--   ), -->
<!--   `Destination accessibility` = c( -->
<!--     "-0.20", -->
<!--     "0.15", -->
<!--     "" -->
<!--   ), -->
<!--   `Distance to transit` = c( -->
<!--     "-0.05 (-0.05 to -0.08)", -->
<!--     "0.15", -->
<!--     "0.29" -->
<!--   ), -->
<!--   `Composite Effect of 5Ds` = c( -->
<!--     "-0.22", -->
<!--     "0.33", -->
<!--     "0.62" -->
<!--   ) -->
<!-- ) %>% -->
<!--   gt::gt( -->
<!--     id = "land-use-elast", -->
<!--     caption = "Land Use Elasticity" -->
<!--   ) %>% -->
<!--   gt::tab_footnote(footnote = "Measured as distance to central business district or jobs reachable within a given travel time", locations = gt::cells_body(6, 1)) %>% -->
<!--   gt::tab_footnote("Measured as jobs located within 1 mile of the home", gt::cells_body(6, 2)) %>% -->
<!--   gt::tab_footnote("Not provided in the Ewing and Cervero study", gt::cells_body(6, 3)) %>% -->
<!--   gt::tab_source_note("Numbers in parentheses indicate the range of elasticities found in the literature") %>% -->
<!--   gt::opt_row_striping() %>% -->
<!--   gt::tab_options(row_group.font.weight = "bold") -->
<!-- ``` -->

#### Density of population and employment

Population and employment are calculated as the total population (jobs) in the CTU divided by the area of *developed* land. This definition allows for the simple comparison of alternative land development scenarios. Developed land includes both impervious surfaces and greenery such as grass and trees. These cover types are explored under the green infrastructure scenarios.

<!-- Distinct maximums are placed on the potential reductions in VMT for each mode. If a value higher than the maximum is entered, the maximum value will be used.  LM: I'm not sure what import this has on a static report?-->

<!-- ```{r tab-maximum-reduction-by-mode} -->
<!-- #| echo: false -->
<!-- #| warning: false -->
<!-- #|  -->
<!-- tibble( -->
<!--   Mode = c("Automobile", -->
<!--            "Walking and cycling", -->
<!--            "Transit"), -->
<!--   `Maximum Reduction in VMT`  = c(enviro_factors$MAX_5D_DR + 1, -->
<!--                                   enviro_factors$MAX_5D_ACT, -->
<!--                                   enviro_factors$MAX_5D_TRANS) -->
<!-- ) %>% -->
<!--   gt::gt( -->
<!--     id = "land-use-elast", -->
<!--     caption = "Land Use Elasticity" -->
<!--   ) %>%  -->
<!--   gt::fmt_percent(columns = 2, decimals = 0) -->

<!-- ``` -->


##### Population

A negative value here indicates a *decrease* in population density, wehre a positive one indicates an *increase* in population density.  

```{r fig-population-density}
#| echo: false
#| message: false

# vmt_telework()
density_ctu %>%
  filter(
    type == "direct",
    year %in% c("2018", "2040")
  ) %>%
  group_by(ctu, year, type, param) %>%
  summarize(
    emissions = sum(emissions, na.rm = T)
  ) %>%
  mutate(level = paste0(as.numeric(param) * 100, "%")) %>%
  ggplot(aes(x = year, y = emissions, color = level, group = level)) +
  geom_line() +
  facet_wrap(~type,
             scales = "free",
             labeller = as_labeller(c(
               direct = str_wrap("Direct emissions", 10),
               embodied = str_wrap("Embodied emissions", 10)
             ))
  ) +
  scale_color_manual(
    values = RColorBrewer::brewer.pal(7, "Blues")[2:7],
    breaks = c("-100%", "-50%", "0%", "50%", "100%")
  ) +
  labs(
    x = "",
    y = "",
    color = paste0(
      params$ctu_selection,
      "\npopulation density\nadjustment (%)"
    )
  ) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(
    panel.background = element_rect(fill = "white"),
    # panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 10)
  )
```

```{r tab-population-density-scenario}
#| echo: false
#| message: false

bau_ctu %>%
  ungroup() %>%
  left_join(transportation_index$modes[2:3], by = c("mode" = "mode_abbrev")) %>%
  filter(
    year %in% c("2018", "2040"),
    type == "direct"
  ) %>%
  pivot_wider(
    names_from = year,
    values_from = emissions
  ) %>%
  select(
    `Mode` = mode_description_1,
    `2018 Baseline` = `2018`,
    `2040 Business-as-Usual` = `2040`
  ) %>%
  left_join(density_ctu %>%
              ungroup() %>%
              left_join(transportation_index$modes[2:3], by = c("mode" = "mode_abbrev")) %>%
              filter(
                year %in% c("2040"),
                type == "direct",
                scenario == "dens50"
              ) %>%
              pivot_wider(
                names_from = year,
                values_from = emissions
              ) %>%
              select(
                `Mode` = mode_description_1,
                `2040 Scenario` = `2040`
              )) %>%
  tableformat(.,
              .caption = paste("Emissions in thousands of tonnes CO2e - 50% Increase in Population Density", params$ctu_selection)
  )
```

##### Employment  

```{r fig-employment-density}
#| echo: false
#| message: false

# vmt_telework()
emp_density_ctu %>%
  filter(
    type == "direct",
    year %in% c("2018", "2040")
  ) %>%
  group_by(ctu, year, type, param) %>%
  summarize(
    emissions = sum(emissions, na.rm = T)
  ) %>%
  mutate(level = paste0(as.numeric(param) * 100, "%")) %>%
  ggplot(aes(x = year, y = emissions, color = level, group = level)) +
  geom_line() +
  facet_wrap(~type,
             scales = "free",
             labeller = as_labeller(c(
               direct = str_wrap("Direct emissions", 10),
               embodied = str_wrap("Embodied emissions", 10)
             ))
  ) +
  scale_color_manual(
    values = RColorBrewer::brewer.pal(7, "Blues")[2:7],
    breaks = c("-100%", "-50%", "0%", "50%", "100%")
  ) +
  labs(
    x = "",
    y = "",
    color = paste0(
      params$ctu_selection,
      "\nemployment density\nadjustment (%)"
    )
  ) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(
    panel.background = element_rect(fill = "white"),
    # panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 10)
  )
```

```{r tab-population-employment-scenario}
#| echo: false
#| message: false

bau_ctu %>%
  ungroup() %>%
  left_join(transportation_index$modes[2:3], by = c("mode" = "mode_abbrev")) %>%
  filter(
    year %in% c("2018", "2040"),
    type == "direct"
  ) %>%
  pivot_wider(
    names_from = year,
    values_from = emissions
  ) %>%
  select(
    `Mode` = mode_description_1,
    `2018 Baseline` = `2018`,
    `2040 Business-as-Usual` = `2040`
  ) %>%
  left_join(emp_density_ctu %>%
              ungroup() %>%
              left_join(transportation_index$modes[2:3], by = c("mode" = "mode_abbrev")) %>%
              filter(
                year %in% c("2040"),
                type == "direct",
                scenario == "empdens50"
              ) %>%
              pivot_wider(
                names_from = year,
                values_from = emissions
              ) %>%
              select(
                `Mode` = mode_description_1,
                `2040 Scenario` = `2040`
              )) %>%
  tableformat(.,
              .caption = paste("Emissions in thousands of tonnes CO2e - 50% Increase in Employment Density", params$ctu_selection)
  )
```




#### Precedents 

The Minneapolis 2040 plan includes a series of changes to zoning to encourage multi-unit dwellings in the city [@cityofminneapolisdepartmentofcommunityplanningandeconomicdevelopmentMinneapolis2040City2019]. The plan calls for additional density along public transit corridors and provides an allowance for up to three dwellings on a single lot in the most outlying neighborhoods. 

<!-- There are several caveats to interpreting this policy change as a precedent for other municipalities in MSP. Minneapolis is the largest city in the region and its high level of existing development means it has minimal land available for new developments. The higher price of land in Minneapolis, relative to other municipalities in MSP, provides an additional incentive for upzoning. It must also be borne in mind that this is an ongoing change and its contribution towards realizing increased density remains unclear. LM: removing for the Mpls specific report-->

A study by MAPC in Massachusetts used a combination of the VisionEval transportation model and the UrbanSim land use model to examine land use and pricing scenarios [@gatelyImpactsLandUse2021] They found a 5.2% difference in VMT between smart growth and sprawl scenarios, equivalent to 2.3 million vehicle-miles traveled per day in that region.

