### Telework

The most recent studies that also include energy associated with working from home find a effective reduction of 2.5-2.8% in household CO2 emissions per day. [@Kim_Choo_Mokhtarian_2015]

A scenario factor (SF) can be introduced to the generic GHG impact equation as an adjustment based on the effect of telework on both transportation and building energy.

At a first approximation, telework eliminates commute VMT. However, the net effect on VMT and household GHG production is less clear. Individuals who work from home tend to make additional non-commute trips during their additional discretionary time [@Mokhtarian_2009] Other household members also can use the now available vehicle for their commute (potentially switching away from transit) or for other trips.

Finally, working from home introduces additional requirements on household energy consumption: more daytime heating and, potentially, the need for a larger dwelling to provide space for a home office. The most recent studies that also include energy associated with working from home find an effective reduction of 2.5-2.8% in household CO2 emissions per telework day. [@fuEnvironmentalPolicyImplications2012a]

An important secondary question regarding telework is its market potential. It is estimated that about 35 percent of jobs in the United States can be performed from home full-time [@Dingel_Neiman_2020].  Prior to the COVID-19 pandemic, time use data for the United States indicates that less than a quarter of workers worked from home, and most spent less than half their working hours at home. The portion of jobs that can be done from home varies by metropolitan area. In MSP, approximately 38 to 52 percent of jobs can be done fully through telework [@Dingel_Neiman_2020]. 

```{r}
#| echo: false
#| message: false
# help(vmt_telework, package = ghg.sp)
```

::: about_strategy
We estimate a 2.75% decrease in personal vehicle miles traveled for each person-day of telework.
:::

```{r fig-telework}
#| echo: false
#| message: false

#| fig.cap = "Comparing the direct and embodied greenhouse gas emissions between years 2018 and 2040 for different percentages of telworking.",
#| fig.align = 'center',
#| fig.width = 8, fig.height = 3


# vmt_telework()
telework_ctu %>%
  filter(year %in% c("2018", "2040")) %>%
  group_by(year, ctu, param, type) %>%
  summarize(emissions = sum(emissions)) %>%
  mutate(level = paste0(as.numeric(param) * 100, "%")) %>%
  ggplot(aes(x = year, y = emissions, color = level, group = level)) +
  geom_line() +
  facet_wrap(~type,
    scales = "free",
    labeller = as_labeller(c(
      direct = str_wrap("Direct emissions", 10),
      embodied = str_wrap("Embodied emissions", 10)
    ))
  ) +
  scale_color_manual(
    values = RColorBrewer::brewer.pal(7, "Blues")[2:7],
    breaks = c("20%", "40%", "60%", "80%", "100%")
  ) +
  labs(x = "", y = "", color = paste0(params$ctu_selection, "\nincrease in\nteleworking (%)")) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(
    panel.background = element_rect(fill = "white"),
    # panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 10)
  )
```

```{r tab-bev-comp}
bau_ctu %>%
  ungroup() %>%
  left_join(transportation_index$modes[2:3], by = c("mode" = "mode_abbrev")) %>%
  filter(
    year %in% c("2018", "2040"),
    type == "direct"
  ) %>%
  pivot_wider(
    names_from = year,
    values_from = emissions
  ) %>%
  select(
    `Mode` = mode_description_1,
    `2018 Baseline` = `2018`,
    `2040 Business-as-Usual` = `2040`
  ) %>%
  left_join(telework_ctu %>%
    ungroup() %>%
    left_join(transportation_index$modes[2:3], by = c("mode" = "mode_abbrev")) %>%
    filter(
      year %in% c("2040"),
      type == "direct",
      scenario == "t60"
    ) %>%
    pivot_wider(
      names_from = year,
      values_from = emissions
    ) %>%
    select(
      `Mode` = mode_description_1,
      `2040 Scenario` = `2040`
    )) %>%
  tableformat(.,
    .caption = paste("Emissions in tonnes CO2e - 60% increase in people teleworking", params$ctu_selection)
  )
```
