### Road Pricing
```{r}
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F
)
```

<!-- TODO links/citations, explain differences between each pricing option and that the tool doesn't allow all three, add equation -->
Road pricing is likely to be an important component of transportation policy going forward. It provides a clear signal to travelers about the cost of driving and contains its own revenue stream. Finally, road pricing provides an alternative revenue stream to the gas tax as vehicle efficiency increase and EV adoption grows.   

The scenario planning model includes three alternatives for road pricing. The first is a vehicle miles traveled (VMT) fee that charges for each mile of travel. The second is a pay-as-you-drive insurance plan. Finally, the model includes a representation of congestion pricing. Ideally, congestion price should vary between the time of day and road link based on the current congestion level, but this complexity lies beyond the scope of this model.

Instead, an elevated VMT fee is applied to a portion of travel, approximating the percent of congested VMT. This method is similar to a the Carbon Free Boston (CFB) study [@porterTransportationTechnicalReport2019].  Credit-based congestion pricing is recommended as a revenue-neutral strategy to address equity in road pricing. It provides credits to travelers who shift their travel model or departure time.


```{r tab-road-price-effects}
readr::read_csv("data/road_price_effects.csv", show_col_types = FALSE) %>%
  rowwise() %>%
  mutate(across(c(3, 4), function(x) {
    ifelse(x != "NA",
      paste0(ifelse(x >= 0, "+", ""), x * 100, "%"),
      x
    )
  })) %>%
  unique() %>%
  gt::gt(
    id = "road-price-effects",
    caption = "Road pricing values",
    rowname_col = "`Road pricing adjustment`",
    groupname_col = "Mode"
  ) %>%
  gt::opt_row_striping() %>%
  gt::tab_options(row_group.font.weight = "bold")
```


```{r tab-road-price-adj-values}
readr::read_csv("data/road_adj_values.csv",
  progress = F,
  show_col_types = F
) %>%
  gt::gt(
    id = "road-price-adj-values",
    caption = "Road pricing adjustment values",
    rowname_col = "`Road pricing adjustment`"
  ) %>%
  gt::opt_row_striping() %>%
  gt::tab_footnote("Congestion pricing typically higher than baseline vehicle miles traveled fee", locations = gt::cells_body(1, 4))
```

<!-- LM: moved to appendix -->
<!-- #### Vehicle miles traveled (VMT) fee  {#vmt-fee} -->

<!-- A study by MAPC [@gatelyImpactsLandUse2021] in Massachusetts  uses a combination of the [VisionEval](https://visioneval.org/) transportation model and UrbanSim land use model to examine land use and pricing scenarios. The introduction of a 25 cent per mile VMT fee is anticipated to lower VMT by 15%. They find that even a tripling of the gas tax from 24 cents per gallon to 75 cents per gallon would only slow VMT growth by 1.5%. The study results indicate the Commonwealth region cannot rely solely on technology-based solutions such as fuel efficiency and electrification to meet ambitious GHG reduction targets. -->


<!-- However, these results are highly dependent on the choice of elasticity and marginal cost. It is also important to note the variation in elasticity obtained in different jurisdictions. -->

<!-- <!-- In one study, the long-run gasoline price elasticity in the United States is found to be -0.23 compared with -1.35 in Europe. This variation in the elasticity is attributable to the historically much higher fuel taxes in Europe leading to the adoption of more fuel-efficient vehicles and more robust transit networks.  --> -->

<!-- Recently, Spiller et al. [@Spiller_Stephens_Timmins_Smith_2014] estimated the price elasticity of VMT at -0.511, which is on the higher end of the range found in the literature. Their higher estimate is partially due to the consideration of multi-vehicle households switching to more fuel-efficient vehicles in response to increases in fuel price. In the case of a VMT fee, this option would not be available to these households. -->


<!-- #### Congestion pricing   -->

<!-- In the CFB analysis [@porterTransportationTechnicalReport2019], they assume that 29% of urban VMT is congested as an input to the determination of congesting pricing. Their estimate is in line with the findings of TAM for the Boston Metropolitan Area, which gives an estimate of 29.5%.  -->

<!-- However, this estimate is for peak VMT, which the TAM team define as 6 AM to 10 AM and 3 PM to 7 PM.  The same TAM study estimates peak VMT congestion in MSP as 24.6%. We assume there is no congestion during non-peak periods and estimate the percent of VMT during the peak period, as a proportion of total VMT for MSP, using weighted TBI 2019 results. -->

<!-- PMT is converted to VMT using the Council's 2019 Travel Behavior Inventory average vehicle occupancy.  -->

<!-- We calculate a secondary set of congestion estimates using TBI data for each CTU. It is assumed that most congestion occurs in the peak periods on weekdays. TBI trips are selected that occur wholly within the same CTU and average travel times estimated for each of peak and off-peak periods. This method suggests an approximately 13% increase in average travel times during peak periods relative to off-peak periods.  -->


<!-- #### Other per-mile costs   -->

<!-- Gas taxes are set at the state and federal level. Current gas tax in Minnesota is \$0.469 per gallon for gasoline, and \$0.529 per gallon for diesel fuel [@CommercePolicyChapter2017].   -->

<!-- Pay-as-you-drive insurance requires technology -->


```{r tab-road-policy}
#| echo: false
#| message: false
# vmt_road_policy()
```


```{r fig-roadprice}
#| echo: false
#| message: false
#| fig.cap:  "Comparing the direct and embodied greenhouse gas emissions between years 2018 and 2040 for different levels of road pricing."
#| fig.align : 'center'
#| fig.width : 8
#| fig.height : 3

roadprice_ctu %>%
  filter(year %in% c("2018", "2040")) %>%
  group_by(year, ctu, param, type) %>%
  summarize(emissions = sum(emissions)) %>%
  filter(
    type == "direct",
    year %in% c("2018", "2040")
  ) %>%
  mutate(level = paste0("$", as.numeric(param))) %>%
  ggplot(aes(x = year, y = emissions, color = level, group = level)) +
  geom_line() +
  facet_wrap(~type,
    scales = "free",
    labeller = as_labeller(c(
      direct = str_wrap("Direct emissions", 10),
      embodied = str_wrap("Embodied emissions", 10)
    ))
  ) +
  scale_color_manual(
    values = RColorBrewer::brewer.pal(7, "Blues")[2:7]
    # ,breaks = c("20%", "40%", "60%", "80%", "100%")
  ) +
  labs(x = "", y = "", color = paste0(params$ctu_selection, "\ncongestion price\n($ per mile)")) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(
    panel.background = element_rect(fill = "white"),
    # panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 10)
  )
```

```{r tab-rp-comp}
bau_ctu %>%
  ungroup() %>%
  left_join(transportation_index$modes[2:3], by = c("mode" = "mode_abbrev")) %>%
  filter(
    year %in% c("2018", "2040"),
    type == "direct"
  ) %>%
  pivot_wider(
    names_from = year,
    values_from = emissions
  ) %>%
  select(
    `Mode` = mode_description_1,
    `2018 Baseline` = `2018`,
    `2040 Business-as-Usual` = `2040`
  ) %>%
  left_join(roadprice_ctu %>%
    ungroup() %>%
    left_join(transportation_index$modes[2:3], by = c("mode" = "mode_abbrev")) %>%
    filter(
      year %in% c("2040"),
      type == "direct",
      scenario == "rp100"
    ) %>%
    pivot_wider(
      names_from = year,
      values_from = emissions
    ) %>%
    select(
      `Mode` = mode_description_1,
      `2040 Scenario` = `2040`
    )) %>%
  tableformat(.,
    .caption = paste("Emissions in tonnes CO2e - $1 per mile congestion price", params$ctu_selection)
  )
```


#### Precedents  

Minnesota examined mileage-based user fees (MBUF) through a pilot program focused on both technical feasibility and traveler attitudes [@rephloConnectedVehiclesSafety2013].  The study included 500 residents of Wright County, which is adjacent and to the west of the 7-county jurisdiction of Met Council. The MBUF rate was set at \$0.03 per mile for travel during peak periods or within an area bounded by I-494 and I-694 and \$0.01 per mile for all other travel. The pilot test generated \$40,000 of simulated revenue from an average monthly fee of \$20 per traveler. The cost was designed to be roughly commensurate with the Minnesota fuel tax. 

Only 17% of participants felt the MBUF was higher than they anticipated. Preference for an MBUF rather than a fuel tax was attributed to it providing a more equitable distribution of transportation costs. Those opposed to the change cited the clearer tax (i.e., monthly bill) versus the current system of incorporating the tax into the price of fuel at the pump. 

Data privacy was not a significant concern among participants, although there is some question as to whether self-selection bias may be a factor since participation in the study was voluntary. The pilot study also took advantage of the GPS technology used to track VMT to provide safety notifications to participants. These notifications included alerts when the driver exceeded the posted speed limit or entered a highway work zone. Many participants preferred that these elements be removed, and the focus placed on MBUF. 

While the project was motivated by a need to find alternative revenue streams due to increased vehicle fuel efficiency, the use of a simulated price means that there was no significant reduction in VMT among participants.  


