### Parking Policy  
```{r, echo=FALSE}
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F
)
```

Parking is an important component of the transportation system. A lack of parking can result in additional VMT from vehicles searching for a parking spot (up to 30% of vehicles in downtown traffic [@shoup_cruising_2006]). Surface lots and parkades take up valuable space that could be put to more useful purposes. 

The most common strategies for controlling parking demand are increasing its price and restricting the time that parking is available. These strategies have been reviewed in detail and price elasticities are available for the price of parking on VMT (see Appendix).

<!-- TODO check equation -->

<!-- The long-run elasticity of VMT to parking cost is estimated to be in the range of -0.18 to -0.45, with a mean of 0.07 based on a meta-analysis of values reported in the literature [@lehner_price_2019]. -->

<!-- $$adj_{park} = 1 + \frac{price_{y}}{price_{y = 2018}} \times elast_{park_{m}} $$ -->

<!-- Where -->

<!-- $price_y$ is the price of parking on dollars per hour in year $y$\ -->
<!-- $price_{y = 2018}$ is the price of parking in dollars per hour in 2018\ -->
<!-- $elast_{park}$ is the elasticity or effect size of parking price increase on travel mode $m$. -->


```{r tab-parking-price-effects}
#| echo: false
#| message: false
ghg.sp::elast %>%
  dplyr::filter(year == "2050") %>%
  select(year, contains("park")) %>%
  tidyr::pivot_longer(
    # names_to = "Increase in parking price",
    values_to = "Effect on miles traveled by 2050",
    cols = 2:4
  ) %>%
  mutate(`Mode` = case_when(
    name == "park_elast" ~ "Automobile",
    name == "park_active" ~ "Walking and cycling",
    name == "park_transit" ~ "Transit"
  )) %>%
  mutate(
    `Increase in parking price relative to baseline` = 0.1,
    "Increase in parking price" = name
  ) %>%
  select(
    Mode,
    `Increase in parking price relative to baseline`,
    `Effect on miles traveled by 2050`
  ) %>%
  rowwise() %>%
  mutate(across(c(2:3), function(x) {
    ifelse(x != "NA",
           paste0(ifelse(x >= 0, "+", ""), x * 100, "%"),
           x
    )
  })) %>%
  gt::gt(
    id = "parking-price-effects",
    caption = "Parking price adjustment values"
  ) %>%
  gt::tab_source_note(gt::md("TRACE 1999")) %>%
  gt::opt_row_striping() %>%
  gt::tab_options(row_group.font.weight = "bold")
```


```{r fig-parkingprice}
#| echo: false
#| message: false
#| fig.cap: "Comparing the direct and embodied greenhouse gas emissions between years 2018 and 2040 for levels of parking pricing."
#| fig.align: "center"
#| fig.width: 8
#| fig.height: 3

parking_ctu %>%
  filter(
    type == "direct",
    year %in% c("2018", "2040")
  ) %>%
  group_by(year, ctu, param, type) %>%
  summarize(emissions = sum(emissions)) %>%
  mutate(level = paste0("$", as.numeric(param))) %>%
  ggplot(aes(x = year, y = emissions, color = level, group = level)) +
  geom_line() +
  facet_wrap(~type,
             scales = "free",
             labeller = as_labeller(c(
               direct = str_wrap("Direct emissions", 10),
               embodied = str_wrap("Embodied emissions", 10)
             ))
  ) +
  scale_color_manual(
    values = RColorBrewer::brewer.pal(7, "Blues")[2:7]
    # ,breaks = c("20%", "40%", "60%", "80%", "100%")
  ) +
  labs(
    x = "", y = "",
    color = paste0(
      params$ctu_selection,
      "\nparking price\n($ per hour)"
    )
  ) +
  scale_y_continuous(labels = scales::label_comma()) +
  theme(
    panel.background = element_rect(fill = "white"),
    # panel.grid = element_blank(),
    axis.title = element_text(size = 14),
    strip.background = element_rect(fill = "white"),
    strip.text = element_text(size = 10)
  )
```


```{r tab-pp-comp}
bau_ctu %>%
  ungroup() %>%
  left_join(transportation_index$modes[2:3], by = c("mode" = "mode_abbrev")) %>%
  filter(
    year %in% c("2018", "2040"),
    type == "direct"
  ) %>%
  pivot_wider(
    names_from = year,
    values_from = emissions
  ) %>%
  select(
    `Mode` = mode_description_1,
    `2018 Baseline` = `2018`,
    `2040 Business-as-Usual` = `2040`
  ) %>%
  left_join(parking_ctu %>%
              ungroup() %>%
              left_join(transportation_index$modes[2:3], by = c("mode" = "mode_abbrev")) %>%
              filter(
                year %in% c("2040"),
                type == "direct",
                scenario == "pk100"
              ) %>%
              pivot_wider(
                names_from = year,
                values_from = emissions
              ) %>%
              select(
                `Mode` = mode_description_1,
                `2040 Scenario` = `2040`
              )) %>%
  tableformat(.,
              .caption = paste("Emissions in thousands of tonnes CO2e - $50 parking per hour", params$ctu_selection)
  )
```
