---
title: "ZTRAX Building Area & CTU Residential Electricity Analysis"
author: "Metropolitan Council"
date: "10/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load required libraries
```{r message=FALSE}
library(data.table)
library(tidyverse)
library(readxl)
library(sf)
library(odbc)
library(DBI)
library(dplyr)
library(tibble)
library(tidyr)
# library(xts)
library(ggplot2)
library(fst)
# library(forecast)
```

test system info to decide path of ZTRAX dataset (stored in N drive)
```{r}
sys.info <- Sys.info()

if (sys.info["sysname"][1] == "Windows"){
  path <- "N:/CommDev/Research/ZTRAX/originals/current_transaction_assessor/20191009/27/ZAsmt/"
} else{
  path <- "/Volumes/shared/CommDev/Research/ZTRAX/originals/current_transaction_assessor/20191009/27/ZAsmt/"
}
```


```{r}
sys.info <- Sys.info()

if (sys.info["sysname"][1] == "Windows"){
  path_layout <- "N:/CommDev/Research/ZTRAX/originals/current_transaction_assessor/20191009/Layout.xlsx"
} else{
  path_layout <- "/Volumes/shared/CommDev/Research/ZTRAX/originals/current_transaction_assessor/20191009/Layout.xlsx"
}
```

load ZTRAX ZAsmt layout 
```{r}
ZAsmt_layout <- readxl::read_excel(path_layout, sheet = "ZAsmt")
ZAsmt_layout <- data.table::as.data.table(ZAsmt_layout)
```
test system info to decide path of ZTRAX layout table (stored in N drive)



```{r}
bldg_areas <- fst::read.fst("/Volumes/shared/CommDev/Research/ZTRAX/fst_data/20201012/ZAsmt/metro_BuildingAreas.fst")
bldg <- fst::read.fst("/Volumes/shared/CommDev/Research/ZTRAX/fst_data/20201012/ZAsmt/metro_Building.fst")
main <- fst::read.fst("/Volumes/shared/CommDev/Research/ZTRAX/fst_data/20201012/ZAsmt/metro_Main.fst")
```

```{r}
main <- as.data.table(main)
geo_info <- main[,.(RowID,
                    PropertyCity,
                    County)]

bldg_area_geo <- data.table::merge.data.table(bldg, geo_info,by = "RowID")

bldg_all <- data.table::merge.data.table(bldg_area_geo, bldg_areas, by = "RowID") %>% as.data.table()

```

```{r}
landuse_dict <- read_xlsx("/Volumes/shared/CommDev/Research/ZTRAX/originals/ZTRAXDataDictionary/LandUse.xlsx") %>% as.data.table()


bldg_all <- data.table::merge.data.table(bldg_all, landuse_dict,by.x = "PropertyLandUseStndCode",
                                         by.y = "StndCode") %>% as.data.table()

```


```{r}
bldg_all[,.(PropertyCity,
            "PropertyLandUse Description",
            Classification,
            PropertyStateLandUseCode,
            PropertyLandUseStndCode,
            PropertyCountyLandUseCode,
            PropertyCountyLandUseDescription)] %>% unique()
```


```{r}
bldg_all[Classification %in% c("Residential",
                               "Residential Income - Multi-Family"),]




[,.(
  mean_sqft = mean(Building_2),
  median_sqft = median(Building_2),
  count = n(),
  sd_sqft = sd(Building_2),
  largest_value_sqft = max(Building_2),
  smallest_value_sqft = min(Building_2),
  average_year_built = round(mean(YearBuilt))), by = .("PropertyCity",
                                                       "County",
                                                       "PropertyCountyLandUseCode",
                                                       "PropertyLandUseStndCode") ]
```


```{r}
ctu_ztrax_summary <-
  intersect_ztrax_ctu %>%
  dplyr::select(c(
    CTU_NAME,
    Prop
    CTU_ID,
    County,
    PropertyLa,
    YearBuilt,
    Building_1,
    Building_2
  )) %>%
  dplyr::filter(Building_2 > 100) %>%
  dplyr::group_by(CTU_NAME, CTU_ID, Building_1, PropertyLa) %>%
  dplyr::summarise(
    mean_sqft = mean(Building_2),
    median_sqft = median(Building_2),
    count = n(),
    sd_sqft = sd(Building_2),
    largest_value_sqft = max(Building_2),
    smallest_value_sqft = min(Building_2),
    average_year_built = round(mean(YearBuilt))
  ) %>%
  dplyr::filter(count > 30) %>%
  dplyr::left_join(.,
                   ztrax_landuse,
                   by = c("PropertyLa" = "StndCode")) %>%
  dplyr::mutate(property_land_use = proper(`PropertyLandUse Description`)) %>%
  dplyr::filter(Classification %in% c("Residential Income - Multi-Family", "Residential")) %>%
  dplyr::rename(
    type_of_sqft_measurment = Building_1,
    ctu_name = CTU_NAME,
    ctu_id = CTU_ID,
    designation = Designation
  ) %>%
  dplyr::mutate(type_of_sqft_measurment = case_when(
    type_of_sqft_measurment == "BAL" ~ "Living Building Area",
    type_of_sqft_measurment == "BAH" ~ "Heated Building Area"
  )) %>%
  dplyr::select(
    ctu_name,
    ctu_id,
    property_land_use,
    designation,
    average_year_built,
    type_of_sqft_measurment,
    mean_sqft,
    median_sqft,
    count,
    sd_sqft,
    smallest_value_sqft,
    largest_value_sqft
  ) 
```



function to open ZTRAX tables with their proper headers
```{r}
OpenZTRAXTable <- 
  function(table_name){
    fread(
      paste0(path, paste0(table_name, ".txt")),
      header = FALSE,
      sep = "|",
      stringsAsFactors = FALSE,
      quote="", 
      col.names = ZAsmt_layout[TableName == paste0("ut", table_name)][,FieldName]
    )
  }
```

---

# Prepare ZTRAX data 

this step consist on fetching the ZTRAX data from where it is store in the N drive
the folder containing this data require special permissions to read/write, contact IS to obtain permission 
for this analysis 

## Load ZTRAX Tables

load Main table from ZTRAX
```{r}
Main <- OpenZTRAXTable("Main")
```

filter Main table to only seven-county metro region
this helps to reduce the volume of the data
```{r}
Main <- Main[County %in% c("ANOKA", "CARVER", "DAKOTA", "HENNEPIN", "Ramsey", "SCOTT", "WASHINGTON")]
```

load "Building" table from ZTRAX
```{r}
Building <- OpenZTRAXTable("Building")
```

load "BuildingAreas" table from ZTRAX
```{r}
BuildingAreas <-OpenZTRAXTable("BuildingAreas")
```

## Filter Raw Data Tables to Fields of Interest

filter "Main" table, which contains longitude and latitude information
```{r}
filtered_Main <- Main[,c('RowID', 'County', 'PropertyAddressLatitude', 'PropertyAddressLongitude')]
```

filter "Building" table, which contains property land use and year built information
```{r}
filtered_Building <- Building[,c('RowID', 'PropertyLandUseStndCode', 'YearBuilt')]
```

filter "BuildingAreas" tables which contains square footage and building area measurement type information
```{r}
filtered_BuildingAreas <- BuildingAreas[, c('RowID','BuildingAreaStndCode','BuildingAreaSqFt')]
filtered_BuildingAreas <- BuildingAreas[BuildingAreaStndCode %in% c("BAL","BAH")]
```

## Join Data Tables

once the ZTRAX tables are cleaned, join them by "RowID"
```{r}
metro_ztrax <-
  filtered_Main %>%
  dplyr::left_join(.,
                   filtered_Building,
                   by = "RowID") %>%
  dplyr::left_join(.,
                   filtered_BuildingAreas,
                   by = "RowID") %>%
  na.omit() 
```

define path to save cleaned and joined data from ZTRAX
in this case I am saving it to my personal folder on the N drive 
```{r}
if (sys.info["sysname"][1] == "Windows"){
  path3 <- "N:/CommDev/Research/Research/Mauricio L/metro_ztrax.csv"
} else{
  path3 <- "/Volumes/shared/CommDev/Research/Research/Mauricio L/metro_ztrax.csv"
}

utils::write.csv(metro_ztrax, path3, row.names = FALSE)
```

remove tables, since they are not longer needed
```{r}
base::rm(Main)
base::rm(Building)
base::rm(BuildingAreas)
base::rm(filtered_Main)
base::rm(filtered_Building)
base::rm(filtered_BuildingAreas)
```

---

# perform analysis of ZTRAX building area

I did the geocoding and intersect analysis on ArcMap, the CRS has to be set up to 4326 for the latitude and longitude to be correct 
```{r}
intersect_ztrax_ctu <- read_csv("N:/CommDev/Research/Research/Mauricio L/data/intersect_ztrax_ctu.csv")
proper=function(x) paste0(toupper(substr(x, 1, 1)), tolower(substring(x, 2)))
ztrax_landuse <- read_csv("N:/CommDev/Research/Research/Mauricio L/data/ztrax_landuse.csv")
```

obtain the building area summary statistics (mean, median, standard deviation) for the CTUs in order to use in regression model
```{r}
ctu_ztrax_summary <-
  intersect_ztrax_ctu %>%
  dplyr::select(c(
    CTU_NAME,
    CTU_ID,
    County,
    PropertyLa,
    YearBuilt,
    Building_1,
    Building_2
  )) %>%
  dplyr::filter(Building_2 > 100) %>%
  dplyr::group_by(CTU_NAME, CTU_ID, Building_1, PropertyLa) %>%
  dplyr::summarise(
    mean_sqft = mean(Building_2),
    median_sqft = median(Building_2),
    count = n(),
    sd_sqft = sd(Building_2),
    largest_value_sqft = max(Building_2),
    smallest_value_sqft = min(Building_2),
    average_year_built = round(mean(YearBuilt))
  ) %>%
  dplyr::filter(count > 30) %>%
  dplyr::left_join(.,
                   ztrax_landuse,
                   by = c("PropertyLa" = "StndCode")) %>%
  dplyr::mutate(property_land_use = proper(`PropertyLandUse Description`)) %>%
  dplyr::filter(Classification %in% c("Residential Income - Multi-Family", "Residential")) %>%
  dplyr::rename(
    type_of_sqft_measurment = Building_1,
    ctu_name = CTU_NAME,
    ctu_id = CTU_ID,
    designation = Designation
  ) %>%
  dplyr::mutate(type_of_sqft_measurment = case_when(
    type_of_sqft_measurment == "BAL" ~ "Living Building Area",
    type_of_sqft_measurment == "BAH" ~ "Heated Building Area"
  )) %>%
  dplyr::select(
    ctu_name,
    ctu_id,
    property_land_use,
    designation,
    average_year_built,
    type_of_sqft_measurment,
    mean_sqft,
    median_sqft,
    count,
    sd_sqft,
    smallest_value_sqft,
    largest_value_sqft
  ) 

```

obtain the building area summary statistics (mean, median, standard deviation) for the CTUs in order to use in regression model
```{r counties ztrax building area summary (county)}
county_ztrax_summary <-
  intersect_ztrax_ctu %>%
  dplyr::select(c(County,
                  PropertyLa,
                  YearBuilt,
                  Building_1,
                  Building_2)) %>%
  dplyr::filter(Building_2 > 100) %>%
  dplyr::group_by(County, Building_1, PropertyLa) %>%
  dplyr::summarise(
    mean_sqft = mean(Building_2),
    median_sqft = median(Building_2),
    count = n(),
    sd_sqft = sd(Building_2),
    largest_value_sqft = max(Building_2),
    smallest_value_sqft = min(Building_2),
    average_year_built = round(mean(YearBuilt))
  ) %>%
  dplyr::filter(count > 30) %>%
  dplyr::mutate(County = proper(County)) %>%
  dplyr::left_join(.,
                   ztrax_landuse,
                   by = c("PropertyLa" = "StndCode")) %>%
  dplyr::mutate(property_land_use = proper(`PropertyLandUse Description`)) %>%
  dplyr::filter(Classification %in% c("Residential Income - Multi-Family", "Residential")) %>%
  dplyr::rename(
    co_name = County,
    type_of_sqft_measurment = Building_1,
    designation = Designation
  ) %>%
  dplyr::mutate(
    type_of_sqft_measurment = case_when(
      type_of_sqft_measurment == "BAL" ~ "Living Building Area",
      type_of_sqft_measurment == "BAH" ~ "Heated Building Area"
    )
  ) %>%
  dplyr::select(
    co_name,
    property_land_use,
    designation,
    average_year_built,
    type_of_sqft_measurment,
    mean_sqft,
    median_sqft,
    count,
    sd_sqft,
    smallest_value_sqft,
    largest_value_sqft
  ) 
```

save the tables
```{r}
# save county ztrax building area summary
write.csv(
  county_ztrax_summary,
  "N:/CommDev/Research/Research/Mauricio L/data/county_ztrax_summary.csv",
  row.names = FALSE
)

# save ctus ztrax building area summary
write.csv(
  ctu_ztrax_summary,
  "N:/CommDev/Research/Research/Mauricio L/data/ctu_ztrax_summary.csv",
  row.names = FALSE
)

ctu_ztrax_summary
```

estimate percentage of housing stock that is single family detached, express as percentage
```{r}
housing_stock_ctu_percentages <- 
  housing_stock_ctu %>%
  dplyr::mutate(
    total_housing_units = multifamily_in_5_or_more_units_bldng +
      single_family_detached +
      townhouse +
      manufactured_homes +
      duplex_triplex_or_quadplex +
      other +
      adu
  ) %>%
  dplyr::filter(year==2017) %>%
  dplyr::group_by(ctu_name) %>%
  dplyr::summarise(
    total_housing_units = sum(total_housing_units),
    multifamily_in_5_or_more_units_bldng  = sum(multifamily_in_5_or_more_units_bldng),
    single_family_detached  = sum(single_family_detached),
    townhouse  = sum(townhouse),
    manufactured_homes  = sum(manufactured_homes),
    duplex_triplex_or_quadplex  = sum(duplex_triplex_or_quadplex),
    other = sum(other),
    adu = sum(adu)
  ) %>%
  dplyr::mutate(
    single_family_detached_perc_housing_stock = single_family_detached / total_housing_units,
    townhouse_perc_housing_stock = townhouse / total_housing_units,
    manufactured_homes_perc_housing_stock = manufactured_homes/ total_housing_units,
    duplex_triplex_or_quadplex_perc_housing_stock = duplex_triplex_or_quadplex/ total_housing_units, 
    other_perc_housing_stock = other/ total_housing_units,
    adu_perc_housing_stock = adu / total_housing_units
  ) 
```

estiamte percentage of all residential land use  that is single family 
```{r}
land_use_historical_ctu_sf_mf_percentage <-
  land_use_historical %>%
  dplyr::filter(year == 2016,
                type == "residential") %>%
  dplyr::group_by(ctu_name, land_use_description) %>%
  dplyr::summarise(acres = sum(acres), .groups = "drop") %>%
  mutate(land_use_description = str_replace_all(land_use_description, " ", "_")) %>%
  mutate(land_use_description = str_replace(land_use_description, "/", "_")) %>%
  mutate(land_use_description = tolower(land_use_description)) %>%
  tidyr::pivot_wider(names_from = land_use_description,
                     values_from = acres,
                     values_fil = 0) %>%
  mutate(total = rowSums(across(c(2:7)))) %>%
  group_by(ctu_name) %>%
  mutate(multifamily_percentage = (mixed_use_residential + multifamily /
                                     total) / total) %>%
  mutate(
    singlefamily_percentage = (
      seasonal_vacation + single_family_detached + single_family_attached + manufactured_housing_park
    ) / total
  )
```


```{r}
residential_electricity_model <-
  rii_mmbtu_electricity_sector %>%
  filter(year == 2017) %>%
  select(ctu_name, residential_mmbtu) %>%
  inner_join(
    .,
    ctu_population %>%
      filter(year == 2017) %>%
      select(ctu_name, population) %>%
      group_by(ctu_name) %>%
      summarise(population = sum(population)),
    by = "ctu_name"
  ) %>%
  inner_join(
    .,
    ctu_ztrax_summary %>%
      select(ctu_name, property_land_use, mean_sqft, average_year_built),
    by = "ctu_name"
  ) %>%
  filter(property_land_use == "Single family residential") %>%
  mutate(residential_mmbtu_per_person = residential_mmbtu / population) %>%
  filter(residential_mmbtu_per_person < 50)  %>%
  inner_join(
    .,
    housing_stock_ctu_percentages %>%
      select(ctu_name, single_family_detached_perc_housing_stock) %>%
      group_by(ctu_name) %>%
      summarise(
        single_family_detached_perc_housing_stock = sum(single_family_detached_perc_housing_stock)
      ),
    by="ctu_name"
  ) %>%
  inner_join(
    .,
    land_use_historical_ctu_sf_mf_percentage %>%
      select(ctu_name, multifamily_percentage, singlefamily_percentage) %>%
      rename(multifamily_percentage_landuse = multifamily_percentage,
             singlefamily_percentage_landuse = singlefamily_percentage),
    by = "ctu_name"
  ) %>%
  inner_join(
    .,
    census_facts %>%
      filter(subject_area_desc == "Mean Household Income") %>%
      filter(year == 2017) %>%
      select(ctu_name, measure) %>%
      rename(mean_household_income = measure)
    ,
    by = "ctu_name"
  )
```

```{r}
regression_model <- lm(residential_mmbtu_per_person ~ 
                         mean_sqft +
                         single_family_detached_perc_housing_stock +
                         I(single_family_detached_perc_housing_stock^2) +
                         singlefamily_percentage_landuse +
                         I(singlefamily_percentage_landuse^2) +
                         mean_household_income, 
                       data = residential_electricity_model)
summary(regression_model)
```

```{r}
plot(residential_electricity_per_ho2$residential_mmbtu_per_person, (residential_electricity_per_ho2$single_family_detached)^2)
plot(residential_electricity_per_ho2$residential_mmbtu_per_person, (residential_electricity_per_ho2$mean_sqft))
```

```{r}
plot(residential_electricity_per_ho3$residential_mmbtu_per_person, residential_electricity_per_ho3$singlefamily_percentage)
ggplot(residential_electricity_per_ho3, aes(x=residential_mmbtu_per_person, y=singlefamily_percentage))+geom_point()+geom_smooth(method=lm)
ggplot(residential_electricity_model, aes(residential_mmbtu_per_person, mean_sqft)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1)

p1 <- ggplot(residential_electricity_model, aes(mean_sqft, residential_mmbtu_per_person)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x, size = 1)

p1 + xlab("Average Building Area (sqft)") + ylab("Per Capita Residential Electricity (MMBtu/person)")



```


```{r}
p2 <- ggplot(residential_electricity_model, aes(single_family_detached_perc_housing_stock, residential_mmbtu_per_person)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1)

p2 + xlab("single family detached percentage (of total housing stock)") + ylab("per capita residential electricity (MMBtu/person)")
```


```{r}
p3 <- ggplot(residential_electricity_model, aes(singlefamily_percentage_landuse, residential_mmbtu_per_person)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1)

p3 + xlab("single family percentage (of land use)") + ylab("per capita residential electricity (MMBtu/person)")
```


```{r}
p4 <- ggplot(residential_electricity_model, aes(mean_household_income, residential_mmbtu_per_person)) +
  geom_point() +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), size = 1)

p4 + xlab("mean household income") + ylab("per capita residential electricity (MMBtu/person)")
```
