---
title: "Land Use Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{land_use_examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ghg.sp)
library(ggplot2)
```

1. How do different land use interventions impact the number of hectares dedicated to different land uses in 2040? (e.g., consider a stacked bar chart with interventions on the x axis, number of hectares on the y axis, colors indicating different land uses, and faceted by different communities)

## EXAMPLE 1

Assume bussiness as usual scenario with the following interventions
- 80% parking lot reduction
- tree planting on all pervious surfaces
- the `tree planting per capita` and `per hectare` are not used when you use the "tree_planting_on_all_previous" intervention. These arguments are only used when .tree_planting_intervention = "match_la_million_trees_goal".
```{r}
parkinglot <- calc_parking_lot_land_cover(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .parking_lot_reduction_percentage = 0.8,
  .tree_planting_intervention = "tree_planting_on_all_pervious",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  detail = FALSE
) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::select(-c(total_area_hectares)) %>%
  tidyr::pivot_longer(names_to = "var", cols = 3:last_col(), values_to = "value")
```

```{r}
ggplot(parkinglot, aes(fill = var, y = value, x = year)) +
  xlab("year") +
  ylab("hectares") +
  geom_bar(position = "stack", stat = "identity")
```

Tree planting on all pervious surfaces is a wacky assumption, but it serves to show the basic idea behind the land use strategies. 

## EXAMPLE 2: A more plausible scenario.

Assume bussiness as usual scenario with the following interventions
- 0% parking lot reduction
- matches la tree planting goal 
- the `tree planting per capita` and `per hectare` are not used when you define the "tree_planting_on_all_previous" intervention as an argumnet. These arguments are only used when .tree_planting_intervention = "match_la_million_trees_goal".
```{r}
parkinglot2 <- calc_parking_lot_land_cover(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .parking_lot_reduction_percentage = 0,
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  detail = FALSE
) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::select(-c(total_area_hectares)) %>%
  tidyr::pivot_longer(names_to = "var", cols = 3:last_col(), values_to = "value")
```

```{r}
ggplot(parkinglot2, aes(fill = var, y = value, x = year)) +
  xlab("year") +
  ylab("hectares") +
  geom_bar(position = "stack", stat = "identity")
```
