---
title: "Land Use Module - Outputs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{land_use_module_outputs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ghg.sp)
```

## Functions

The impact of different land use and greenery strategies is estimated using a series of functions with arguments provided by the user.

The key principle is of this model is that different land cover types have different carbon stock and sequestration value. By altering the composition of a city/township, we could observe higher or lower overall carbon stocks or sequestration value per year. 

### Functions to Calculate Urban Form Strategy. 

The first function is `ghg.sp::calc_by_development_type()`
For help run the following command `?ghg.sp::calc_by_development_type()`
This function is intermediary and does not provide the final result. 

`ghg.sp::calc_by_development_type()` 
The function returns a table with the land composition of different urban form scenarios of land use by development type.

The tool will  perform a series of calculations in 
response to different values of argument `.urban_form_scenario`.
For example, if the selected input is `"bau"`, or the business as usual, 
the tool will select the table it needs. 

The possible scenarios to select on the `.urban_form_scenario` argument include:
* `"bau"`: the business as usual scenario.
* `"post_covid_sprawl"`: post-covid sprawl scenario.
* `"compact_development_beyond_bau"`: compact development beyond the business as usual scenario.
* `"compact_development_with_drs"`: compact development with dynamic ride sharing.

The following code prints the value of function `ghg.sp::calc_by_development_type()` filtered to show only the city of Lake Elmo. 
```{r}
ghg.sp::calc_land_by_development_type(tb = land_use_data,
                                      .urban_form_scenario = "bau") %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

The second function is `ghg.sp::calc_scen_land_use()`
The following code runs the `ghg.sp::calc_scen_land_use()` for the `"bau"` or business as usual scenario. 
```{r}
ghg.sp::calc_scen_land_use(tb =  land_use_data,
                           .urban_form_scenario = "bau") %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 70)
```

The third function is `ghg.sp::calc_land_cover_by_land_use()`
The following code runs the `ghg.sp::calc_land_cover_by_land_use()` for the `"bau"` or business as usual scenario.
```{r}
ghg.sp::calc_land_cover_by_land_use(tb =  land_use_data,
                                    .urban_form_scenario = "bau") %>%
  dplyr::filter(year == 2040) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::filter(land_cover_type == 	"forest") %>%
  head(., n = 25)
```

The fourth function is `ghg.sp::calc_land_cover_by_city()`
This function returns the urban form scenario land composition in hectares without adding additional strategies. This is the last function used to estimate the impact of urban form changes alone.  
```{r}
# ghg.sp::calc_land_cover_by_city(tb = land_use_data,
#                                 .urban_form_scenario = "bau") %>%
#   dplyr::filter(year == 2040) %>%
#   dplyr::filter(ctu_name == "Lake Elmo") %>%
#   head(., n = 15)
```

### Function to Calculate Tree Planting Strategy

Having calculated the hectares of land cover under the selected urban form scenario, we now will add the change to the land composition  that is cause by implement a tree planting strategy. 

Function `ghg.sp::calc_tree_planting_land_cover()`.
```{r}
ghg.sp::calc_tree_planting_land_cover(
  tb = land_use_data,
  detail = FALSE,
  .urban_form_scenario = "bau",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247
)  %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  dplyr::filter(year == 2040) %>%
  tidyr::pivot_longer(
    names_to = "var",
    values_to = "value",
    cols = -c("ctu_name", "year")
  ) %>%
  head(., n = 15)
```

### Function to Calcuate Parking Lot Reduction Strategy

Having calculate the hectares of land cover under the selected urban form and tree planting scenario, we now will add the change to the land composition that is cause by implementing a reduced parking lot strategy. 

Function `ghg.sp::calc_parking_lot_land_cover()`.
```{r}
ghg.sp::calc_parking_lot_land_cover(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .parking_lot_reduction_percentage = 0.8,
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  detail = FALSE
) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  tidyr::pivot_longer(
    names_to = "var",
    values_to = "value",
    cols = -c("ctu_name", "year")
  )

```

### Function to Calculate Conservation Tillage Strategy  

Calculates the impact of conservation tillage scenarios on carbon stocks by city/township


```{r}
ghg.sp::calc_conservation_tillage(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .conservation_tillage_intervention = "maximum_conservation_tillage",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  .parking_lot_reduction_percentage = 0.8,
  detail = FALSE
)  %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  tidyr::pivot_longer(
    names_to = "var",
    values_to = "value",
    cols = -c("ctu_name", "co_name")
  ) %>%
  head(., n = 15)
```

### Function to Calculate Carbon Stock per City/Township
```{r}
ghg.sp::calc_carbon_stock_per_ctu(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  .parking_lot_reduction_percentage = 0.8,
  .conservation_tillage_intervention = "current_conservation_tillage",
  detail = FALSE
) %>%
  dplyr::filter(ctu_name == "Lake Elmo")
```

### Function to Calculate Carbon Sequestration per City/Township
```{r}
ghg.sp::calc_carbon_sequestration_per_ctu(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  .parking_lot_reduction_percentage = 0.80,
  detail = FALSE
) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 10)
```

### Main function to Run Scenario 
```{r}
ghg.sp::run_scenario_land_use(
  tb = land_use_data,
  .urban_form_scenario = "bau",
  .conservation_tillage_intervention = "current_conservation_tillage",
  .tree_planting_intervention = "match_la_million_trees_goal",
  .tree_planting_per_capita = 0.26,
  .tree_planting_per_hectare = 247,
  .parking_lot_reduction_percentage = 0.8,
  detail = FALSE
) %>%
  dplyr::filter(ctu_name == "Lake Elmo") %>%
  head(., n = 10) 
```
