---
title: "Transportation - Land Use"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{trans_land_use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  fig.width = 8
)

requireNamespace("ggplot2", quietly = TRUE)
requireNamespace("scales", quietly = TRUE)
requireNamespace("councilR", quietly = TRUE)
requireNamespace("wesanderson", quietly = TRUE)
requireNamespace("gt", quietly = TRUE)
library(ghg.sp)
library(ggplot2)
library(dplyr)
library(stringr)
library(scales)
library(wesanderson)
library(councilR)
library(tibble)
library(gt)
library(emo)

ggplot2::theme_set(
  if(testthat:::on_ci() == TRUE){
    theme_minimal()
  } else {
    councilR::theme_council(use_showtext = T,
                            use_manual_font_sizes = T)
  }
)

```

Goal - we need to have the existing values for every CTU for each land use

## Population and Employment density


```{r "pop-emp-dens", echo=F}
pop_density <- tibble::tribble(
  ~type,   ~CTU,   ~`2018`,   ~`2020`,   ~`2030`,   ~`2040`,
  "Population", "Lake Elmo",  1.67,  1.82,  2.58,  2.94,
  "Population",  "Shoreview",  4.43,   4.4,  4.39,  4.41,
  "Population",  "St. Paul", 10.02, 10.06,  10.6, 11.18,
  "Employment",  "Lake Elmo", 0.47, 0.48, 0.48,  0.5,
  "Employment", "Shoreview", 2.03, 2.14, 2.28, 2.39,
  "Employment", "St. Paul",  6.1, 6.47, 6.86, 7.24
)

pop_density %>% 
  gt::gt(
    caption = "Density: persons/jobs per urban acre",
    rowname_col = "CTU",
    groupname_col = "type"
  ) %>% 
  # gt::fmt_number(columns = 2:9, sep_mark = ",",
  #                drop_trailing_zeros = T) %>% 
  gt::opt_row_striping()
```

### Housing/population     

```{r eval = FALSE, echo = FALSE}
import_from_emissions("metro_demographic.vw_ctu_population") %>% 
  filter(year == "2016") %>% 
  left_join(
    import_from_emissions("metro_land.vw_ctu_land_use_hectares") %>% 
      dplyr::left_join(  import_from_emissions("metro_land.land_use_2016_types")) %>% 
      dplyr::mutate(is_developed =case_when(description_2 %in% c(
        "agricultural",
        "open_water",
        "park_recreational_or_preserve",
        "undeveloped"                          
      ) ~ "undeveloped_sq_mi",
      TRUE ~ "developed_sq_mi")) %>% 
      group_by(ctu_name, is_developed) %>% 
      summarize(total_area_sq_mi = sum(hectares) * 2.471) %>% 
      tidyr::pivot_wider(names_from = is_developed,
                         values_from = total_area_sq_mi)) %>% 
  mutate(pop_density_developed = population/developed_sq_mi) %>%
  arrange(ctu_name) %>% View
```

### Employment    

## Land use diversity  

Land use diversity is calculated using an entropy measure and the forecasted land uses. The “mixed-use” categories are assumed to contain two land uses for the purpose of calculating the entropy term.

$$S = \sum_i LU_i \times \ln(LU_i)$$

where $S$ entropy, $LU$ is the proportion of land devoted to land use type $i$. Entropy is anticipated to increase in the more outlying CTUs, such as Lake Elmo and Shoreview, than in the central city of St. Paul.

```{r tab-entropy, echo = FALSE}
tibble::tribble(
  ~CTU, ~`2015`, ~`2020`, ~`2025`, ~`2030`, ~`2035`, ~`2040`, ~`2045`, ~`2050`,
  "Lake Elmo",    0.44,    0.46,    0.48,     0.5,    0.51,    0.53,    0.55,    0.57,
  "Shoreview",    0.41,    0.42,    0.43,    0.44,    0.46,    0.48,     0.5,    0.53,
  "St. Paul",    0.62,    0.61,    0.59,    0.58,    0.59,    0.61,    0.63,    0.64
) %>% 
  gt::gt(
    caption = "Current and forecasted land used entropy by CTU",
    rowname_col = "CTU"
  ) %>% 
  gt::opt_row_striping()

```

## Intersection design  

OpenStreetMap street network data were used to calculate the distribution of intersection configurations in each CTU (OpenStreetMap Contributors, 2017). Service roads, parking lots, and several other features that are coded as streets in the data were removed from the network.

Percentage of intersections by number of intersecting paths.  

```{r intersection-dens, echo=F}
tibble::tribble(
  ~CTU, ~`3-way`, ~`4-way`, ~`Other`,
  "Lake Elmo",    "75.9%",    "21.8%",     "2.2%",
  "Shoreview",    "78.1%",    "20.6%",     "1.3%",
  "St. Paul",    "42.9%",    "55.6%",     "1.5%"
) %>% 
  gt::gt(
    caption = "Intersection density",
    rowname_col = "CTU"
  ) %>% 
  gt::opt_row_striping()
```

SHCN did provide us with a complete dataset for all CTUs, but did NOT provide reproducible code.  


1. Figure out how intersections were identified in OSM and how they quantified the number of ways. Key columns here would be `intersection_id`, `ctu_name`, `n_ways`. `n_ways` should be a factor, with levels `2-way`, `3-way` and `4+  way`
1. `dplyr::group_by()` CTU and calculate distribution by each number of ways. 


Ashley Asmus provided her code from another project [here](./data-raw/trans_intersection_design.R)

## Average minimum distance to transit stops   

SHCN did not provide complete data for average minimum distance to transit stops and provided limited methods documentation. To recreate these values, here is a suggested workflow.  


1. Pull point level residential building data  
- Jason pulled data from OpenStreetMaps (OSM) and filtered to use only residential buildings (exactly which key values is unknown). We could replicate this using `{osmdata}`, but we don't have as  much familiarity and confidence in the OSM building data
- **Best option** Another option would be to use the metro [AddressPoints](https://gisdata.mn.gov/dataset/us-mn-state-metrogis-loc-address-points) data. These include point location and official address for all occupiable units and any other official addresses. The `RESIDENCE` field could be used to filter. The data is also vetted by each county, which ensures some additional data quality.  
- Yet another option is to go down to the parcel level. The data here will be more detailed, but much more difficult to use and reliably identify residential building types.   
1. Pull transit stops point data (from GTFS, GISLibrary, etc).
1. For each residence point, find the nearest transit stop using `sf::st_nearest()` and index the residence - stop pair.  
1. For each residence-stop pair, calculate the distance between each residence-stop pair using `sf::st_distance()`.
1. `dplyr::group_by()` CTU, and calculate the mean, median, mode, etc. for each residence-stop pair distance.  

Linear distance should be fine for an analysis of this magnitude. Network distance would be better, but because we are calculating a mean at the CTU level, the difference will likely be relatively small.  

In similar analyses previously, I've found that using vector and parallel processing with `{furrr}` is the most efficient method. Basically, use `sf::st_nearest()` on a single residence point and the entire stops dataset. Then, only calculate the distance with `sf::st_distance()` for the residence-stop pair (not the distance between all residences and all stops). 

```{r "transit-stop-distance", echo=F}
tibble::tribble(
  ~CTU,  ~Min,   ~Max,  ~Mean, ~Median, ~"Std Dev",   ~Q1,    ~Q3,
  "Lake Elmo", 185.9, 4956.2, 2269.1,  1039.2,     1956.9, 457.9, 4742.7,
  "Shoreview", 145.2,   7818, 1032.9,   704.9,     1098.4, 533.1,  992.9,
  "St. Paul",  23.7, 2764.7,  619.2,   551.8,      367.7, 332.5,    866
) %>% 
  gt::gt(
    rowname_col = "CTU",
    caption = "Summary of distance to nearest transit stop in feet"
  ) %>% 
  gt::opt_row_striping() %>% 
  gt::fmt_number(columns = 2:8, sep_mark = ",",
                 drop_trailing_zeros = T) 


```

## Job accessability via transit  

Its not clear what this is measuring exactly. There was some note about number of jobs within a 1 mile radius of the home, but this won't work well for less dense CTUs. Number of jobs you can reach within 1 hour by transit?

